---
title: "Sensitivity analyses"
author: "Esin Ickin"
date: "2/23/2023"
output: html_document
---

This code is an example on how to parameterize vital-rate functions using coefficients and covariate inputs to then build a matrix population model and compute sensitivities.

## 1) Vital rates

The example is from a paper on gray mouse lemurs: Ozgul et al. (in press)
In the paper, we parameterized five vital rates as a function of population density, rainfall, and temperature.

The vital rates are defined as functions, and we need the coefficients from vital rate models (here GLMs). We would appreciate a similar input on your part. Alternatively, it possible to have GLM models as an object, and then just use the "predict" function, instead of building out own functions. 

```{r}


```

## 2) Covariates

There are the input data for the vital rate functions. Here, we have a time series of data. This is the best format because it allows us to calculate not only means and variances, but also covariances, the latter being important to calculate scaled sensitivities. 

However, it is also possible to send us just the values we are interested in.

```{r}

```

## 3) Population model

Here, we use the vital rate function to construct an annual population model that can give us the population growth rate (lambda).

In the following example, the MPM is constructed with mean covariate values.
In the perturbations, the covariate values are changed.

```{r}

```

## 4) Scaled sensitivity analyses 

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117)

Note that this is a step that Esin will implement in her MS thesis. With the information given in 1-3, we should be able to run these analyses.


## 5) Sensitivity analyses at equilibrium dynamics 

Here, we perform "classic" sensitivity analyses (see Paniw et al. 2019; DOI: 10.1126/science.aau5905)

1. We look for combinations of covariate values where lambda approaches 1


2. We perturb each covariate for each vital rate in turn, while maintaining the other covariates at values that would ensure stable population dynamics. A perturbation entails a 10 % increase in a covariate.   


3. Plot the sensitivities 

Note that when perturbing all covariates in all vital rate models, the resulting changes in lambda are much less pronounced then when looking at single vital rate models.  

