---
title: "Sensitivity analyses"
author: "Esin Ickin, Chloé Nater"
date: "4/13/2023"
output: html_document
---

This code is an implementation of Esin's template for Nater et al.'s study on arctic foxes.
The data files needed for running the code below, as well as the original files from which the code below is taken, have been deposited in a folder "Files_ArcticFox" within the repository. Questions about details of the study and/or code in sections 1-3 below can be sent to chloe.nater@nina.no.

## 1) Vital rates

The following estimates and code are from a papers on Arctic foxes (Alopex lagopus) from Svalbard: https://doi.org/10.1002/ecs2.3546

The study simultaneously estimated vital rates as functions of three environmental covariates (sea ice extent, availability of reindeer carcasses, and goose population size) and population dynamics using an Integrated Population Model (conveniently also abbreviated "IPM").

Like most integrated analyses, the model has been implemented in a Bayesian framework, meaning that a) all parameter estimates are stored as posterior distributions and b) all vital rate prediction functions are custom-made.

The file containing the entire posterior distributions for all parameters is quite large (about 140 000 KB) and hence not suitable for uploading to GitHub. It can, however, be downloaded as part of the papers supplementary material. 
For convenience, I have therefore included a smaller file with posterior summaries (median, 95\% credible interval, mean, standard deviation) in the folder "Files_ArcticFox". 
In case you want to re-run everything, this is the code used for summarizing the posterior distributions: 

```{r, eval = FALSE}
library(coda)

load("Files_ArcticFox/ArcticFox_IPM_PosteriorSamples.RData")
out.data <- reshape2::melt(as.matrix(MCMCsamples))

out.sum <- out.data %>%
  dplyr::group_by(Var2) %>%
  dplyr::summarise(Median = median(value),
                   lower_90CI = quantile(value, probs = 0.05),
                   upper_90CI = quantile(value, probs = 0.95),
                   Mean = mean(value),
                   SD = sd(value)) %>%
  dplyr::rename(Parameter = Var2)

saveRDS(out.sum, file = "Files_ArcticFox/ArcticFoxIPM_PosteriorSummary.rds")
```

You will find an explanation of all variable names in the model code ("ArcticFox_IPM.R") which I have added to "Files_ArcticFox". 

Below is code used for making vital rate predictions based on enviornmental covariates. The vital rates that get predictions are pregnancy rate (Psi), litter size in utero (rho), survival from birth to emergence from den (s0), and natural mortality (mO). The final vital rate, harvest mortality (mH), was modelled for two time periods but not as a function of covariates (see implementation of matrix model below).

The prediction function below lets one specify which vital rate to make predictions for. It currently works on posterior samples, so you'll need to download those from the article supplementary to make it work. Note that the function also makes predictions over a plausible range of z-standardized covariate values, not just for a single value. 
So that is not quite what you need yet, but it contains everything you need to write additional functions that calculate vital rates for a specific combination of covariate values (either using posterior samples, or a posterior summary such as the median). 

The function inputs are a matrix of posterior samples (`MCMC.mat`), a string called `effect` that specifies the covariate to make predictions for, and an integer `RepAgeClass` which determines which age class the vital rate predictions should apply to. Age class 1 represents young of the year, age classes  2, 3, 4, and 5 represents adult foxes aged 1, 2, 3, and 4+ years. 


```{r, eval = FALSE}
library(coda)

## Making vital rate predictions depending on (standardized) env. covariates

# Load posterior samples
load("Files_ArcticFox/ArcticFox_IPM_PosteriorSamples.RData")

# Re-structure into matrix
MCMC.mat <- as.matrix(MCMCsamples)

# Set prediction range
SDcov <- seq(-2.5, 2.5, length.out = 100)

cov.predict <- function(MCMC.mat, effect, RepAgeClass){
  
  # Prepare data frames
  Psi.pred <- rho.pred <- mO.pred <- S0.pred <- data.frame(SDcov = SDcov, median = rep(NA, length(SDcov)), lCI = rep(NA, length(SDcov)), uCI = rep(NA, length(SDcov)), Covariate = effect)
  
  Psi.pred$VitalRate <- 'Pregnancy rate'
  rho.pred$VitalRate <- 'Fetus number'
  mO.pred$VitalRate <- 'Natural mortality'
  S0.pred$VitalRate <- 'Denning survival'
  
  # Set effect sizes  
  if(effect == 'RdCarcass'){
  	beta.Psi <- MCMC.mat[,"betaRC.Psi"]
  	beta.rho <- MCMC.mat[,"betaRC.rho"]
  	beta.mO <- MCMC.mat[,"betaRC.mO"]
  	beta.m0 <- MCMC.mat[,"betaRC.m0"]
  }

  if(effect == 'SeaIce'){
  	beta.Psi <- MCMC.mat[,"betaSI.Psi"]
  	beta.rho <- MCMC.mat[,"betaSI.rho"]
  	beta.mO <- MCMC.mat[,"betaSI.mO"]
  	beta.m0 <- MCMC.mat[,"betaSI.m0"]
  }
  
  if(effect == 'Goose'){
  	beta.Psi <- 0
  	beta.rho <- 0
  	beta.mO <- MCMC.mat[,"betaG.mO"]
  	beta.m0 <- 0
  }
  
  
  # Make predictions 
  for(i in 1:length(SDcov)){
  	
  	# Pregnacy rate
    eta <- plogis(MCMC.mat[,"par.b"]*(MCMC.mat[,"par.c"] - RepAgeClass) + beta.Psi*SDcov[i])
    Psi <- MCMC.mat[,"par.a"]*eta
    
    Psi.pred[i,2] <- quantile(Psi, 0.5)
    Psi.pred[i,3] <- quantile(Psi, 0.05)
    Psi.pred[i,4] <- quantile(Psi, 0.95)
    
    # Placental scars
    rho <- exp(log(MCMC.mat[,'mean.rho']) + MCMC.mat[,'a.eff1']*RepAgeClass + beta.rho*SDcov[i])
    
    rho.pred[i,2] <- quantile(rho, 0.5)
    rho.pred[i,3] <- quantile(rho, 0.05)
    rho.pred[i,4] <- quantile(rho, 0.95)
    
    # Natural mortality
    if(effect == 'Goose'){
    	mO <- exp(log(MCMC.mat[,'Mu.mO[2]']) + beta.mO*SDcov[i])
    }else{
    	mO <- exp(log(MCMC.mat[,'Mu.mO[1]']) + beta.mO*SDcov[i])
    }
    
    mO.pred[i,2] <- quantile(mO, 0.5)
    mO.pred[i,3] <- quantile(mO, 0.05)
    mO.pred[i,4] <- quantile(mO, 0.95)
    
    # Denning survival
    m0 <- exp(log(-log(MCMC.mat[,'S0'])) + beta.m0*SDcov[i])
    S0 <- exp(-m0)
    
    S0.pred[i,2] <- quantile(S0, 0.5)
    S0.pred[i,3] <- quantile(S0, 0.05)
    S0.pred[i,4] <- quantile(S0, 0.95)
    
  }
  
  results <- rbind(Psi.pred, rho.pred, mO.pred, S0.pred)
  
  return(results)
}


CarcassEff <- cov.predict(MCMC.mat, 'RdCarcass', 3)
SeaIceEff <- cov.predict(MCMC.mat, 'SeaIce', 3)
GooseEff <- cov.predict(MCMC.mat, 'Goose', 3)

```

The dataframes returned by the function can then be used to plot vital rates as a function of covariates: 

```{r, eval = FALSE}
library(ggplot2)
library(viridis)

ggplot(CarcassEff, aes(x = SDcov, y = median)) + geom_line(aes(color = VitalRate)) + geom_ribbon(aes(ymin = lCI, ymax = uCI, fill = VitalRate), color = NA, alpha = 0.5) + ylab('Estimate') + xlab('Covariate value') + theme_bw() + theme(panel.grid = element_blank(), legend.position = 'none', plot.title = element_text(hjust = 0.5)) + scale_color_manual(values = c(viridis(7)[6], viridis(7)[2], viridis(7)[1], viridis(7)[3])) + scale_fill_manual(values = c(viridis(7)[6], viridis(7)[2], viridis(7)[1], viridis(7)[3])) + facet_wrap(~VitalRate, scales = 'free_y') + ggtitle('Effects of reindeer carcasses on vital rates')

ggplot(SeaIceEff, aes(x = SDcov, y = median)) + geom_line(aes(color = VitalRate)) + geom_ribbon(aes(ymin = lCI, ymax = uCI, fill = VitalRate), color = NA, alpha = 0.5) + ylab('Estimate') + xlab('Covariate value') + theme_bw() + theme(panel.grid = element_blank(), legend.position = 'none', plot.title = element_text(hjust = 0.5)) + scale_color_manual(values = c(viridis(7)[6], viridis(7)[2], viridis(7)[1], viridis(7)[3])) + scale_fill_manual(values = c(viridis(7)[6], viridis(7)[2], viridis(7)[1], viridis(7)[3])) + facet_wrap(~VitalRate, scales = 'free_y') + ggtitle('Effects of sea ice on vital rates')

ggplot(GooseEff, aes(x = SDcov, y = median)) + geom_line(aes(color = VitalRate)) + geom_ribbon(aes(ymin = lCI, ymax = uCI, fill = VitalRate), color = NA, alpha = 0.5) + ylab('Estimate') + xlab('Covariate value') + theme_bw() + theme(panel.grid = element_blank(), legend.position = 'none', plot.title = element_text(hjust = 0.5)) + scale_color_manual(values = c('grey95', 'grey95', viridis(7)[5], 'grey95')) + scale_fill_manual(values = c('grey95', 'grey95', viridis(7)[5], 'grey95')) + facet_wrap(~VitalRate, scales = 'free_y') + ggtitle('Effects of goose reproduction on vital rates')

```

## 2) Covariates

For estimating covariate effects on vital rates, covariates were de-trended and z-standardized. The de-trending was necessary because there are strong directional changes in most environmental variables on Svalbard due to climate change and without de-trending, a "general trend" would be the one signal overwhelming everything else. 

Nonetheless, the input covariates for prediction functions will all have means of 0 and standard deviations of 1. 
In case you are interested in mapping this back onto the actual covariate values, I have added a dataframe with the de-trended time series of covariates (original and standardized). 


```{r}
envData <- readRDS("Files_ArcticFox/ArcticFoxIPM_EnvCov.rds")
envData
```

## 3) Population model

While often not explicitly written out as matrices, the cores of integrated population models are typically stage-structured population models that can be formulated as an MPMs. This is also the case for this model, and the matrix formulation can be found in the main article.  
Note, however, that this is an open population model, meaning it also includes an immigration component that gets added in addition to the matrix multiplication. 

The code below is adapted from Appendix 3 of the study, where the posterior medians from the integrated analysis are used to parameterize a stochastic MPM (which is then used to assess whether the estimates provide realistic projections of population dynamics through time).


```{r}
library(ggplot2)
library(viridis)

#-------------------------------------------#
# GENERATE AND STORE YEAR-SPECIFIC MATRICES #
#-------------------------------------------#

## Load dataframe of posterior summaries
post.sum <- readRDS("Files_ArcticFox/ArcticFoxIPM_PosteriorSummary.rds")

## Convert into vector of medians
post.median <- post.sum$Median
names(post.median) <- post.sum$Parameter

## Set maximum time-step in data
Tmax <- 23

## Prepare an array to store projection matrices
mat.Y <- array(0, dim = c(5, 5, Tmax-1))

## Prepare a matrix to store initial population sizes
N0.Y <- matrix(NA, nrow = 5, ncol = Tmax-1)

## Prepare a vector to store immigrant numbers
Imm.Y <- rep(NA, Tmax-1)

## Extract year-specific matrices and population sizes based on posterior means
for(t in 1:(Tmax-1)){
	
	# Extract vital rates
	Sj <- exp(-(post.median[paste("mH[2, ", t, "]", sep = "")] + post.median[paste("mO[2, ", t, "]", sep = "")]))
  	Sa <- exp(-(post.median[paste("mH[1, ", t, "]", sep = "")] + post.median[paste("mO[1, ", t, "]", sep = "")]))
  
  	Psi2 <- post.median[paste("Psi[2, ", t+1, "]", sep = "")]
  	Psi3 <- post.median[paste("Psi[3, ", t+1, "]", sep = "")]
  	Psi4 <- post.median[paste("Psi[4, ", t+1, "]", sep = "")]
  	Psi5 <- post.median[paste("Psi[5, ", t+1, "]", sep = "")]
  	
  	rho2 <- post.median[paste("rho[2, ", t+1, "]", sep = "")]
  	rho3 <- post.median[paste("rho[3, ", t+1, "]", sep = "")]
  	rho4 <- post.median[paste("rho[4, ", t+1, "]", sep = "")]
  	rho5 <- post.median[paste("rho[5, ", t+1, "]", sep = "")]
  	
  	S0 <- exp(-post.median[paste("m0t[", t+1, "]", sep = "")])
	
	
	# Write projection matrix
	mat.Y[1, 1, t] <- Sj*0.5*Psi2*rho2*S0
	mat.Y[1, 2, t] <- Sa*0.5*Psi3*rho3*S0
	mat.Y[1, 3, t] <- Sa*0.5*Psi4*rho4*S0
	mat.Y[1, 4, t] <- Sa*0.5*Psi5*rho5*S0
	mat.Y[1, 5, t] <- Sa*0.5*Psi5*rho5*S0
	
	mat.Y[2, 1, t] <- Sj
	
	mat.Y[3, 2, t] <- Sa
	
	mat.Y[4, 3, t] <- Sa
	
	mat.Y[5, 4, t] <- Sa
	mat.Y[5, 5, t] <- Sa

	
	# Extract starting population size
	N0.Y[1, t] <- post.median[paste("N[1, ", t, "]", sep = "")]
	N0.Y[2, t] <- post.median[paste("N[2, ", t, "]", sep = "")]
	N0.Y[3, t] <- post.median[paste("N[3, ", t, "]", sep = "")]
	N0.Y[4, t] <- post.median[paste("N[4, ", t, "]", sep = "")]
	N0.Y[5, t] <- post.median[paste("N[5, ", t, "]", sep = "")]
	
	# Extract immigrant numbers
	Imm.Y[t] <- post.median[paste("Imm[", t+1, "]", sep = "")]
}


#--------------------------------------------------------------------------#
# FUNCTION FOR STOCHASTIC PROJECTION USING A PRE-DEFINED SEQUENCE OF YEARS #
#--------------------------------------------------------------------------#

stoch.Proj = function(YearSeq){
	
	# Determine projection length
	Tmax <- length(YearSeq)
	
	# Make population vector
	N <- matrix(NA, nrow = 5, ncol = Tmax+1)
	
	# Set starting population size
	N[,1] <- N0.Y[,YearSeq[1]]
	
	# Project population
	for(t in 1:Tmax){
		
		N[,t+1] <- (mat.Y[,,YearSeq[t]]%*%N[,t]) + c(Imm.Y[YearSeq[t]], 0, 0, 0, 0)
	}
	
	return(N)	
}

#--------------------------------------------------------------------------#
# FUNCTION TO RUN A SERIES OF PROJECTIONS AND RETURN RESULTS AS DATA FRAME #
#--------------------------------------------------------------------------#

stoch.Sims = function(i, data.Tmax, sim.Tmax){ # i = simulation number
	
	# Make a random sequence of years
	YearSeq <- sample(c(1:(data.Tmax-1)), sim.Tmax, replace = T)
	
	# Population projection
	N <- stoch.Proj(YearSeq)
	
	# Arrange projection results in a data frame
	output <- data.frame(
	SimNo = i, 
	SimYear = c(1:(length(YearSeq)+1)), 
	Ntot = colSums(N),
	N1 = N[1,],
	N2 = N[2,],
	N3 = N[3,],
	N4 = N[4,],
	N5 = N[5,],
	p1 = N[1,]/colSums(N),
	p2 = N[2,]/colSums(N),
	p3 = N[3,]/colSums(N),
	p4 = N[4,]/colSums(N),
	p5 = N[5,]/colSums(N)
	)
	
	# Return results
	return(output)
}


#------------------------------------------#
# RUNNING SIMULATIONS AND PLOTTING RESULTS #
#------------------------------------------#

## Set number of replicates to simulate
SimNoMax <- 10

## Set number of years to simulate for
SimYearMax <- 50

## Running simulations
sim.results <- do.call("rbind", sapply(1:SimNoMax, FUN = function(i) stoch.Sims(i, data.Tmax = 23, sim.Tmax = SimYearMax), simplify = FALSE))

## Plot simulations over time

# Total population size
ggplot(sim.results, aes(x = SimYear, y = Ntot, group = SimNo)) + geom_line(color = magma(8)[4], alpha = 0.5, linewidth = 1) + geom_hline(aes(yintercept = 0), linetype = 'dotted', linewidth = 0.3) + ylab('Population Size') + xlab('Simulation Year') + theme_bw() + theme(panel.grid = element_blank())

```

## 4) Scaled sensitivity analyses 

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117)

Note that this is a step that Esin will implement in her MS thesis. With the information given in 1-3, we should be able to run these analyses.


## 5) Sensitivity analyses at equilibrium dynamics 

Here, we perform "classic" sensitivity analyses (see Paniw et al. 2019; DOI: 10.1126/science.aau5905)

1. We look for combinations of covariate values where lambda approaches 1


2. We perturb each covariate for each vital rate in turn, while maintaining the other covariates at values that would ensure stable population dynamics. A perturbation entails a 10 % increase in a covariate.   


3. Plot the sensitivities 

Note that when perturbing all covariates in all vital rate models, the resulting changes in lambda are much less pronounced then when looking at single vital rate models.  

