---
title: "Sensitivity analyses"
author: "Esin Ickin, Chloé Nater"
date: "4/11/2023"
output: html_document
---

This code is an implementation of Esin's template for Nater et al.'s study on striped mice.
The data files needed for running the code below, as well as the original files from which the code below is taken, have been deposited in a folder "Files_StripedMouse" within the repository. Questions about details of the study and/or code in sections 1-3 below can be sent to chloe.nater@nina.no.


## 1) Vital rates

The following estimates and code are from two papers on African striped mice (Rhabdomys pumilio):

- Nater et al. 2016; https://doi.org/10.1111/oik.03384
- Nater et al. 2018; https://doi.org/10.1111/1365-2656.12888


The first paper established the vital rate models. The second paper ran updated versions of the vital rate models, incorporated them into a stage-structured matrix population model, and performed a variety of analyses of the MPM. The striped mouse has a short life cycle, and both vital rate predictions and the MPM have a monthly time interval. 

The vital rates are defined as sets of functions, one each to calculate the predicted vital rate on the link scale (logit or log) and one to back-transform to the actual probabilities/rates. 
The functions use z-standardized environmental covariates as input and require the fitted model objects to be loaded into the workspace. 

Note that custom prediction functions are only provided for the vital rates originating from the mark-recapture analysis (survival S and transition/maturation probabilities Psi). For the remaining vital rates (breeding probability Bp, litter probability Lp, and litter size LS), R's internal `predict()` function is used. 

```{r}
library(lme4)
library(ggplot2)
library(popbio)
library(plyr)
library(gridExtra)

# Load workspace containing fitted models
load('Files_StripedMouse/170503_allmodels_final.RData')

#################################
#### CMR PREDICTION FORMULAS ####
#################################

# SURVIVAL

# Provides the y value (linear predictor) of S, where logit(S) = y and y = beta1*x1 + beta2*x2
S.y = function(stage,temp,food,dens_F){
  
  X <- c(1, as.numeric(stage=='I'), as.numeric(stage=='P'), temp, food, dens_F,
         (stage=='I')*temp, (stage=='P')*temp,
         (stage=='I')*food, (stage=='P')*food,
         (stage=='I')*dens_F, (stage=='P')*dens_F,
         temp*food, food*dens_F,
         (stage=='I')*temp*food, (stage=='P')*temp*food,
         (stage=='I')*food*dens_F, (stage=='P')*food*dens_F)
  
  X %*% est$estimate[1:18]
}

# Transforms the y value with the correct link function
S.predict = function(stage,temp,food,dens_F){
  return(plogis(S.y(stage,temp,food,dens_F)))
}

# MATURATION
Psi.y = function(stage,temp,food,dens_F){
  X <- c(as.numeric(stage=='I'), as.numeric(stage=='P'),
         temp, food, dens_F,
         food*dens_F, temp*dens_F, temp*food)
  X %*% est$estimate[57:64]
}

Psi.predict = function(stage,temp,food,dens_F){
  return(1/(1+2*exp(-Psi.y(stage,temp,food,dens_F))))    
}
```


## 2) Covariates

There are the input data for the vital rate functions, and the time-series used in the study is provided as a dataframe named "env" in the workspace. This contains both raw and z-standardized versions of covariates temperature (mean.temp, temp) and food availability (life_cover, food). The density covariate is only provided in the standardized form (dens_all, dens_F). Additionally, the dataframe contains an unused NDVI variable, and season and Period variables that are only relevant for plotting and/or calculation of recapture probabilities.


```{r}
######################################
#### Preparing Environmental Data ####
######################################

# Extract scaling parameters for population density
m_dens <- 19.22219
sd_dens <- 8.199321

# Set "baseline" area
area.fix <- 19343.53/10000 # 1.9434353

# Add "lagged" environmental covariates (t-1)
env$temp_t <- NA
env$food_t <- NA
env$dens_F_t <- NA

for(i in 2:nrow(env)){
  env$temp_t[i] <- env$temp[i-1]
  env$food_t[i] <- env$food[i-1]
  env$dens_F_t[i] <- env$dens_F[i-1]
}
```

## 3) Population model

Below are functions for assembling the monthly projection matrices. It's a set of two nested functions, where the first one returns a matrix based on specified covariate values for temperature, food, and density, while the second one calls upon the first, but allows for incorporation of the density feedback (relevant for projections):

```{r}
my.matrix = function(temp, temp_t, food, food_t, dens_F, dens_F_t){
  
  Si <- S.predict('I',temp,food,dens_F)
  Sp <- S.predict('P',temp,food,dens_F)
  Sb <- S.predict('B',temp,food,dens_F)
  
  PsiIB <- Psi.predict('I',temp,food,dens_F)
  PsiPB <- Psi.predict('P',temp,food,dens_F)
  
  newd <- data.frame(food_t = food_t, dens_F_t = dens_F_t, temp_t = temp_t) 
  
  Bp <- unname(predict(bpmodF, newdata = newd, re.form=NA, type='response'))
  
  Lp <- predict(lpmodF, type = "response", newdata = newd)
  LS <- predict(lsmod_Wlag, type = "response", newdata = newd)
  #LS <- 2.65
  
  A <- matrix(data = c(Si*PsiIB*Lp*LS, Si*(1-PsiIB), Si*PsiIB,
                       Sp*PsiPB*Lp*LS, Sp*(1-PsiPB), Sp*PsiPB,
                       Sb*Bp*Lp*LS,            0,       Sb),nrow=3)
  return(A)
}

timedens.matrix = function(t,dens){
  
  A <- my.matrix(env$temp[t],env$temp_t[t],
                 env$food[t],env$food_t[t],
                 dens[1],dens[2])
  return(A)
}

```

As part of the study, we also set up a function for calculating both asymptotic and transient population growth rates. As this may be relevant, I also add this here. 
Transient growth rates require information on population size/structure, so this is read in as additional data. 

```{r}
############################################
#### Calculating 'Real' Population Size ####
############################################

# Reading in population counts
pop.count <- read.csv('Files_StripedMouse/170110_allfemales_stage.csv')
pop.count <- subset(pop.count, !(session%in%c(6,8,16,17,18,38,39)))

pop.count$total <- pop.count$breeders + pop.count$philopatrics + pop.count$immatures + pop.count$unknown

# Correcting observed numbers with recapture probabilities
pop.count <- merge(pop.count, recap, by = 'session', all.x = TRUE)

pop.count$breeders_corr <- pop.count$breeders / pop.count$pB
pop.count$philopatrics_corr <- pop.count$philopatrics / pop.count$pP
pop.count$immatures_corr <- pop.count$immatures / pop.count$pP

pop.count$total_corr <- pop.count$breeders_corr + pop.count$philopatrics_corr + pop.count$immatures_corr

# Reading in data on area size over time
data <- read.csv('Files_StripedMouse/161222_soc_fac_ad.csv')
area.t <- data$area_h_corr

```

With that in place, we can then calculate different types of population growth rates:

```{r}
##########################################
#### CALCULATE DIFFERENT GROWTH RATES ####
##########################################

growth.rates = function(t){
  
  # Obtain the initial population vector
  N <- c(pop.count$immatures_corr[t],pop.count$philopatrics_corr[t],pop.count$breeders_corr[t])
  
  Nvec <- matrix(NA, 12000, 3)
  Nvec[1,] = c(pop.count$immatures_corr[t-1],pop.count$philopatrics_corr[t-1],pop.count$breeders_corr[t-1])/area.t[t-1]
  Nvec[2,] = c(pop.count$immatures_corr[t],pop.count$philopatrics_corr[t],pop.count$breeders_corr[t])/area.t[t]
  
  # Calculate the matrix
  A <- my.matrix(env$temp[t],env$temp_t[t],
               env$food[t],env$food_t[t],
               env$dens_F[t],env$dens_F_t[t])
  
  # Calculate 1-timestep transient growth rate
  T.lambda <- sum(A%*%N)/sum(N)
  
  # Calculate asymptotic growth rate (dominant right eigenvalue)
  A.lambda <- lambda(A)
  
  # Calculate pseudo-asymptotic growth rate (accounting for density feedback)
  #for(x in 2:11999){
  #  
  #  # Calculate corrected sum of philopatrics and breeders
  #  total.no <- Nvec[x,2] + Nvec[x,3]
  #  total.no_t <- Nvec[x-1,2] + Nvec[x-1,3]
  #  
  #  # Scale density
  #  scale.dens <- (total.no - m_dens) / sd_dens
  #  scale.dens_t <- (total.no_t - m_dens) / sd_dens
  #  
  #  dens <- c(scale.dens, scale.dens_t)
  #  
  #  # 4) Feed density into new matrix and project
  #  Nvec[x+1,] = timedens.matrix(t,dens) %*% Nvec[x,]
  #  
  #}
  
  #psA.lambda <- sum(Nvec[12000,])/sum(N)
  psA.lambda <- NA
  
  return(data.frame(time = t, A.lambda = A.lambda, T.lambda = T.lambda, psA.lambda = psA.lambda))
}

# Re-adding missing sessions in pop.count
sessions <- data.frame(session = c(1:117))
pop.count <- merge(pop.count, sessions, by = 'session', all.y = TRUE)

GR <- do.call("rbind", sapply(2:nrow(env), FUN = function(t) growth.rates(t), simplify = FALSE))
```

This gives us the resulting estimates in a dataframe:
```{r}
head(GR)
```

## 4) Scaled sensitivity analyses 

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117)

Note that this is a step that Esin will implement in her MS thesis. With the information given in 1-3, we should be able to run these analyses.


## 5) Sensitivity analyses at equilibrium dynamics 

Here, we perform "classic" sensitivity analyses (see Paniw et al. 2019; DOI: 10.1126/science.aau5905)

1. We look for combinations of covariate values where lambda approaches 1


2. We perturb each covariate for each vital rate in turn, while maintaining the other covariates at values that would ensure stable population dynamics. A perturbation entails a 10 % increase in a covariate.   


3. Plot the sensitivities 

Note that when perturbing all covariates in all vital rate models, the resulting changes in lambda are much less pronounced then when looking at single vital rate models.  

