---
title: "OPIM_IPM_Resampling"
author: "Esin Ickin"
date: "2023-10-10"
output: pdf_document
---

Code to build an Integral Projection Model (IPM) for Opuntia imbricata and conduct sensitivity analyses.

IPM based on script from Aldo Compagnoni. Accompanying article https://doi.org/10.1002/ecm.1228
2 vital rate models are different. These are based on analysis from Sanne Evers. Accompanying article https://doi.org/10.1111/gcb.15519

Edited by: Esin Ickin (14.09.2023)
Sensitivity analyses in step 7.

In this code we're trying to estimate the uncertainties of the sensitivities. For that we resample 100 times from the posterior distributions. Therefore I had to get the posterior distributions of the parameters and these are now in a list of lists (cholla).

## 0) Prepare session
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Library/CloudStorage/OneDrive-UniversitätZürichUZH/Master Thesis/pert_analyses/OpuntiaImbricata")
```

```{r}
library(tidyverse)
library(lme4)

rm(list=ls())

# load model
s_mod <- readRDS("results/OPIM/OPIM_s_sliding2.rds")[[3]]$BestModel
fp_mod <- readRDS("results/gcb/OPIM_fp_month_result.rds")[[4]]$BestModel

# years climate dataframe
s_df <- readRDS("results/OPIM/OPIM_s_sliding2.rds")[[3]]$BestModelData %>%
  rename(tmin_n1_n9 = climate) %>% select(year, tavg_1_n5, tmin_n1_n9) %>% distinct
fp_df <- readRDS("results/gcb/OPIM_fp_month_result.rds")[[4]]$BestModelData %>%
  rename(tavg_17_12 = climate) %>% select(year, tavg_17_12) %>% distinct

clim_df <- left_join(s_df, fp_df)

#Useful functions
invlogit<-function(x){exp(x)/(1+exp(x))}
vol<-function(h,w,p){(1/3)*pi*h*(w/2)*(p/2)}
```


## 1) Vital rates
```{r}
### Load vital rate model results ########################################
#"Variance-correlation matrix
vCor = read.csv("data/OPIM/ScriptsShare/Results/vitalRates/opuntia_correlation_mean.csv")
vCor = as.matrix(vCor)

#Coefficients
# coefs1 are the mean parameters
# coefs1=read.csv("data/OPIM/ScriptsShare/Results/vitalRates/opuntia_coefficient_mean.csv")
# coefs are the posterior distributions
#coefs=read.csv("Results/vitalRates/cholla_coef_ljk_(nbin2015).csv")
coefs=read.csv("data/OPIM/ScriptsShare/Results/vitalRates/opuntia_coefficient_posterior.csv")

#Posterior(s) of correlations
corPost=read.csv("data/OPIM/ScriptsShare/Results/vitalRates/opuntia_correlation_posterior.csv")

#Posterior of coefficients  
coefPost=read.csv("data/OPIM/ScriptsShare/Results/vitalRates/opuntia_coefficient_posterior.csv")

#Time-varying model parameters
fit.grow=coefs[,c("mu.int.G","beta.G","sigma.int.G","sigma")] # Growth (random year)
fit.surv=coefs[,c("mu.int.S","beta.S","sigma.int.S")]         # Survival (random year)
fit.flower=coefs[,c("mu.int.F","beta.F","sigma.int.F")]       # Probability of flowering
fit.buds=coefs[,c("mu.int.B","beta.B","sigma.int.B")]         # Size-dependent fertility

### Vital rates estimated 'from scratch' #########################################

##A. Dispersal of seeds from maternal plant and survival until germination time
## (from fruit removal experiments)-----
fruit.dat<-read.csv("data/OPIM/opuntia_FruitSurvival.csv",T)
names(fruit.dat)

fruitsurv<-glm((Fr.on.grnd.not.chewed/Fr.on.plant~1),weights=Fr.on.plant,family="binomial",data=fruit.dat)
invlogit(coef(fruitsurv)[1])


##B. Germination----------------------
germ.dat<-read.csv("data/OPIM/opuntia_Germination.csv",T)
names(germ.dat)

## germination of 1-yo seeds
germ1yo<-glm(Seedlings04/Input~1,weights=Input,family="binomial",data=germ.dat)
invlogit(coef(germ1yo)[1])

## germination of 2-yo seeds
germ2yo<-glm(Seedlings05/(Input-Seedlings04)~1,weights=(Input-Seedlings04),family="binomial",data=germ.dat)
invlogit(coef(germ2yo)[1])


##C. Survival from germination to census------------------------
precensus.dat<-read.csv("data/OPIM/opuntia_PrecensusSurvival.csv",T)
names(precensus.dat)

precensus.surv<-glm(survive0405~1,family="binomial",data=precensus.dat)
invlogit(coef(precensus.surv)[1])


##D. Size distribution of recruits------------------------------
demog.dat<-read.csv("data/OPIM/opuntia_20042015.csv",T)

### MODIFICATIONS TO APPLY "vol" function ##########################################################
#I change two "demog.dat$Perp_t" values that are == 0 to 0.2.
#Justified because these are actually small individuals (see below).
demog.dat[demog.dat$Perp_t==0,c("Height_t","Width_t","Perp_t")]
demog.dat$Perp_t[demog.dat$Perp_t==0]=0.2
#I also change "zeros" in year t1
demog.dat[c(394,3117,3258),c("Height_t1","Width_t1","Perp_t1")]
demog.dat$Perp_t1[c(3117,3258)]=0.2
#LIKELY MISTAKE: "Width_t1" is zero, but not in the previous years
#We therefore exclude this individual from analysis
demog.dat[394,]
demog.dat[demog.dat$TagID=="LJ12" & demog.dat$Plot=="T1",]
demog.dat=demog.dat[-394,]

#Data to calculate recruit size
recruits.dat <- subset(demog.dat,Recruit==1 & Year_t==2005)
names(recruits.dat)

kidsize.mean<-with(recruits.dat,
		mean(log(vol(h=Height_t,w=Width_t,p=Perp_t))))
kidsize.sd<-with(recruits.dat,
		sd(log(vol(h=Height_t,w=Width_t,p=Perp_t))))


#E. Calculate sizes for all individuals---------------------------
demog.dat$sizet<-vol(h=demog.dat$Height_t,w=demog.dat$Width_t,p=demog.dat$Perp_t)
demog.dat$sizet1<-vol(h=demog.dat$Height_t1,w=demog.dat$Width_t1,p=demog.dat$Perp_t1)


#F. Number of seeds per bud (from Johanna Ohm's data)---------------------------
seednum <- 139.4815 #Data comes from Ohm and Miller, published in Ecology in 2014 
  

###3.Collect IPM parameters#############################################################
cholla <- list()
cholla[[1]] <- fit.grow[,c("mu.int.G")]					## growth intercept
cholla[[2]] <- fit.grow[,c("beta.G")]					  ## growth slope
cholla[[3]] <- NA #fixef(fit.grow)[3]					  ## growth quadratic term
cholla[[4]] <- NA #resid.model$coef[1]					## intercept of size-dependent variance
cholla[[5]] <- NA #resid.model$coef[2]					## slope of size-dependent variance
cholla[[6]] <- fixef(s_mod)[1]					        ## survival intercept
cholla[[7]] <- fixef(s_mod)[2]					        ## survival slope
cholla[[8]]<- fixef(fp_mod)[1]				          ## pr_flower intercept
cholla[[9]]<- fixef(fp_mod)[2]				          ## pr_flower slope
cholla[[10]]<- fit.buds[,c("mu.int.B")]					## nfruits intercept
cholla[[11]]<- fit.buds[,c("beta.B")]					  ## nfruits slope
cholla[[12]]<- seednum						             ## seeds per fruit
cholla[[13]]<- invlogit(coef(fruitsurv)[1])^2 		## This is 6mo fruit survival, squared to estimate full year. But I think this way overestimates seed predation (I'm equating dispersal with predation).
											        ## Experiment with setting it to 1, i.e. all seeds get to the seed bank (note that germination rates account for seed predation)
cholla[[14]]<-invlogit(coef(germ1yo)[1])			## 1-year-old germination
cholla[[15]]<-invlogit(coef(germ2yo)[1])			## 2-year-old germination
cholla[[16]]<-invlogit(coef(precensus.surv)[1])	## pre-census recruit survival
cholla[[17]]<-kidsize.mean						## mean recruit size
cholla[[18]]<-kidsize.sd						## SD of recruit size
cholla[[19]]<-min(log(demog.dat$sizet),na.rm=T)	## minsize (floor)
cholla[[20]]<-max(log(demog.dat$sizet1),na.rm=T) ## maxsize (ceiling)
cholla[[21]]<-200								## matrix size (results seem v stable by 200)

## RANDOM EFFECTS
cholla[[22]] = fit.grow[,"sigma.int.G"]     ## sd of random year effects for growth 
cholla[[23]] = fit.surv[,"sigma.int.S"]     ## sd of random year effects for survival
cholla[[24]] = fit.flower[,"sigma.int.F"]   ## sd of random year effects for flowering
cholla[[25]] = fit.buds[,"sigma.int.B"]     ## sd of random year effects for nbuds

#Should I include "fixed effect SD" or the SD of residuals?
cholla[[26]] = fit.grow[,"sigma"]           #Only growth has residual variance


## CLIMATE EFFECTS
cholla[[27]] <- fixef(s_mod)[3]                 ## survival climate effect Tavg from 1 to -5
cholla[[28]] <- fixef(s_mod)[4]                 ## survival climate effect Tmin from -1 to -9
cholla[[29]] <- fixef(fp_mod)[3]                ## flower p climate effect Tavg from 17 to 12

#Run simulations, including "correlation on" sims. with "direct estimation of variance"
#diag(vCor)=cholla[22:25]^2
varCovar=function(corMat,varVec){
  out = corMat * (varVec %*% t(varVec))
  return(out)
}
#varVec=cholla[c(22:25)]
varVec=lapply(cholla[22:25], function(inner_list) inner_list[[1]])


###4. Define demographic functions###########################################################

#GROWTH FROM SIZE X TO Y
#Function produces a *probability density distribution* from each
#1. Size at time t, (the *x* argument) to each
#2. Size at time t+1 (the *y* argument)
gxy<-function(x,y,params,rfx,i){
	xb=pmin(pmax(x,params[[19]][[1]]),params[[20]][[1]]) #Transforms all values below/above limits in min/max size
	return(dnorm(y,mean=params[[1]][[i]]+params[[2]][[i]]*xb + rfx[1],sd=params[[26]][[i]]))
	}

#SURVIVAL AT SIZE X.
#This calculates the survival prob. of each size class 
sx<-function(x,params,rfx, s_tavg, s_tmin){
	xb=pmin(pmax(x,params[[19]][[1]]),params[[20]][[1]])
	return(invlogit(params[[6]][[1]] + params[[7]][[1]]*xb + params[[27]][[1]]*s_tavg + params[[28]][[1]]*s_tmin + rfx[2]))
	}

#SURVIVAL*GROWTH
pxy<-function(x,y,params,rfx, s_tavg, s_tmin){
	xb=pmin(pmax(x,params[[19]][[1]]),params[[20]][[1]])
	sx(xb,params,rfx, s_tavg, s_tmin)*gxy(xb,y,params,rfx,i)
	#gxy(xb,y,params,random) ## set survival equal to one and check colsums of T
	}

#PRODUCTION OF 1-YO SEEDS IN THE SEED BANK FROM X-SIZED MOMS
fx<-function(x,params,rfx, fp_tavg,i){
	xb=pmin(pmax(x,params[[19]][[1]]),params[[20]][[1]])
	p.flow<-invlogit(params[[8]][[1]] + params[[9]][[1]]*xb + params[[29]][[1]]*fp_tavg + rfx[3]) #Probability of flowering
	nfruits<-exp(params[[10]][[i]] + params[[11]][[i]]*xb + rfx[4])   #n. of flowers produced

  return(p.flow*nfruits*params[[12]][[1]]*params[[13]][[1]])  #Params[c(12,13)] are n. of seeds/fruit, 
                                                #plus seed survival to next year
}



#SIZE DISTRIBUTION OF RECRUITS
#If a recruit is produced, what is the likelihood of falling 
#being in a certain size (*y*) class?
recruits<-function(y,params){
	dnorm(x=y,mean=params[[17]][[1]],sd=params[[18]][[1]])
}


###5. Simulation parameters###############################################################
max.yrs       <- 50000
matsize       <- 200
extra.grid    <- 2        #How many row/col add to the part of the matrix referring to sizes?
random        <- T     
allparms      <- cholla
varCorr       <- vCor

#Eviction parameters
floor.extend   <- 0.5
ceiling.extend <- 2.25
lower          <- cholla[[19]][[1]] - floor.extend
upper          <- cholla[[20]][[1]] + ceiling.extend

```


## 2) Covariates
Climate data
```{r}
# min, max, sd, mean of covariates
# tavg_1_n5
min_tavg_1_n5=min(clim_df$tavg_1_n5)
max_tavg_1_n5=max(clim_df$tavg_1_n5)
mean_tavg_1_n5=mean(clim_df$tavg_1_n5)
sd_tavg_1_n5=sd(clim_df$tavg_1_n5)

# tmin_n1_n9
min_tmin_n1_n9=min(clim_df$tmin_n1_n9)
max_tmin_n1_n9=max(clim_df$tmin_n1_n9)
mean_tmin_n1_n9=mean(clim_df$tmin_n1_n9)
sd_tmin_n1_n9=sd(clim_df$tmin_n1_n9)

# tavg_17_12
min_tavg_17_12=min(clim_df$tavg_17_12)
max_tavg_17_12=max(clim_df$tavg_17_12)
mean_tavg_17_12=mean(clim_df$tavg_17_12)
sd_tavg_17_12=sd(clim_df$tavg_17_12)

# covariation among covariates
# tavg_1_n5
tmin_n1_n9_when_tavg_1_n5_max=clim_df$tmin_n1_n9[which(clim_df$tavg_1_n5==max(clim_df$tavg_1_n5))][1]
tmin_n1_n9_when_tavg_1_n5_min=clim_df$tmin_n1_n9[which(clim_df$tavg_1_n5==min(clim_df$tavg_1_n5))][1]

tavg_17_12_when_tavg_1_n5_max=clim_df$tavg_17_12[which(clim_df$tavg_1_n5==max(clim_df$tavg_1_n5))][1]
tavg_17_12_when_tavg_1_n5_min=clim_df$tavg_17_12[which(clim_df$tavg_1_n5==min(clim_df$tavg_1_n5))][1]

# tmin_n1_n9
tavg_1_n5_when_tmin_n1_n9_max=clim_df$tavg_1_n5[which(clim_df$tmin_n1_n9==max(clim_df$tmin_n1_n9))][1]
tavg_1_n5_when_tmin_n1_n9_min=clim_df$tavg_1_n5[which(clim_df$tmin_n1_n9==min(clim_df$tmin_n1_n9))][1]

tavg_17_12_when_tmin_n1_n9_max=clim_df$tavg_17_12[which(clim_df$tmin_n1_n9==max(clim_df$tmin_n1_n9))][1]
tavg_17_12_when_tmin_n1_n9_min=clim_df$tavg_17_12[which(clim_df$tmin_n1_n9==min(clim_df$tmin_n1_n9))][1]

# tavg_17_12
tmin_n1_n9_when_tavg_17_12_max=clim_df$tmin_n1_n9[which(clim_df$tavg_17_12==max(clim_df$tavg_17_12))][1]
tmin_n1_n9_when_tavg_17_12_min=clim_df$tmin_n1_n9[which(clim_df$tavg_17_12==min(clim_df$tavg_17_12))][1]

tavg_1_n5_when_tavg_17_12_max=clim_df$tavg_1_n5[which(clim_df$tavg_17_12==max(clim_df$tavg_17_12))][1]
tavg_1_n5_when_tavg_17_12_min=clim_df$tavg_1_n5[which(clim_df$tavg_17_12==min(clim_df$tavg_17_12))][1]
```


## 3) Population model
Construct IPM Kernel
```{r}
bigmatrix<-function(params,random,sigma,lower,upper,rand.seed, s_tavg, s_tmin, fp_tavg,i){ 
  
	n<-params[[21]][[1]]
	L<-lower; U<-upper;
	#these are the upper and lower integration limits
	h<-(U-L)/n                   #Bin size
	b<-L+c(0:n)*h;               #Lower boundaries of bins 
	y<-0.5*(b[1:n]+b[2:(n+1)]);  #Bins' midpoints
	#these are the boundary points (b) and mesh points (y)
	
	# Random effects
	rfx <- matrix(0,1,4)  #Set random effect to 0 (in case random=F,corr=F) 
	if(random==T){        
	  set.seed(rand.seed)
	  rfx[1,1:4] = rmvnorm(n=1, mean=rep(0,4), sigma=sigma)
	}
 
	# Fertility matrix
	Fmat<-matrix(0,(n+2),(n+2))

	# Banked seeds go in top row
	Fmat[1,3:(n+2)]<-fx(y,params,rfx, fp_tavg,i)

	# Growth/survival transition matrix
	Tmat<-matrix(0,(n+2),(n+2))

	# Graduation to 2-yo seed bank = pr(not germinating as 1-yo)
	Tmat[2,1]<-1-params[[14]][[1]]

	# Graduation from 1-yo bank to cts size = germination * size distn * pre-census survival
	Tmat[3:(n+2),1]<-params[[14]][[1]]*recruits(y,params)*h   

	# Graduation from 2-yo bank to cts size = germination * size distn * pre-census survival
	Tmat[3:(n+2),2]<-params[[15]][[1]]*recruits(y,params)*h   

	# Growth/survival transitions among cts sizes
	Tmat[3:(n+2),3:(n+2)]<-t(outer(y,y,pxy,params=params,rfx=rfx, s_tavg = s_tavg, s_tmin = s_tmin))*h

	# Put it all together
	IPMmat<-Fmat+Tmat     #Full Kernel is simply a summation ot fertility
                        #and transition matrix
	return(list(IPMmat=IPMmat,Fmat=Fmat,Tmat=Tmat,meshpts=y,rfx=rfx))
}

## Calculate deterministic lambda
#i = 2                                   ## Set this i from 1:13 for the different years

#lambda<-Re(eigen(bigmatrix(params=cholla,random=F,
#                           lower=cholla[19],upper=cholla[20],rand.seed=42,
#                           s_tavg = clim_df$tavg_1_n5[i], s_tmin = clim_df$tmin_n1_n9[i], 
#                           fp_tavg = clim_df$tavg_17_12[i])$IPMmat)$values[1])
#lambda

# IPM KERNEL FUNCTION
# I compute the sensitivities with this function:
kernel.fun<-function(tavg_1_n5,tmin_n1_n9,tavg_17_12,i){
 lam <-  Re(eigen(bigmatrix(params=cholla,random=F,
                     lower=cholla[[19]][[1]],upper=cholla[[20]][[1]],rand.seed=42,
                     s_tavg = tavg_1_n5, s_tmin = tmin_n1_n9, 
                     fp_tavg = tavg_17_12,i=i)$IPMmat)$values[1])
 return(lam)
  
}
```


## 4) Sensitivity analyses
Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: <https://doi.org/10.1073/pnas.1918363117>)

We compute sensitivites for each covariate with and without covariation. We resample parameters from post distributions 100 times. The analyses are done with the IPM KERNEL FUNCTION names kernel.fun() (line 355)
```{r}
#### Sensitivity to tavg_1_n5 -----------------------

# no covariation (mean covariates)
Delta.tavg_1_n5=NULL # define empty list for the sensitivities
for(i in 1:100){ # resampling
  max_lam<-kernel.fun(max_tavg_1_n5,mean_tmin_n1_n9,mean_tavg_17_12,i=i) # model with covariate set to max value
  min_lam<-kernel.fun(min_tavg_1_n5,mean_tmin_n1_n9,mean_tavg_17_12,i=i) # model with covariate set to min value
  
  # sensitivity
  Delta.tavg_1_n5[i] <- abs((max_lam-min_lam)/((max_tavg_1_n5-min_tavg_1_n5)/sd_tavg_1_n5))
  
}
# dist of sens
hist(Delta.tavg_1_n5)

# same as above but with covariation
# by covariation we mean, i.e. the value of tmin_n1_n9 when tavg_1_n5 is set to max or min value (see chunk 2)
Delta.tavg_1_n5.cov=NULL
for(i in 1:100){
  max_lam<-kernel.fun(max_tavg_1_n5,tmin_n1_n9_when_tavg_1_n5_max,tavg_17_12_when_tavg_1_n5_max,i=i)
  min_lam<-kernel.fun(min_tavg_1_n5,tmin_n1_n9_when_tavg_1_n5_min,tavg_17_12_when_tavg_1_n5_min,i=i)
  
  Delta.tavg_1_n5.cov[i] <- abs((max_lam-min_lam)/((max_tavg_1_n5-min_tavg_1_n5)/sd_tavg_1_n5))
  
}

hist(Delta.tavg_1_n5.cov)

# Differences between the two sensitivities (no cov-cov)
hist(Delta.tavg_1_n5-Delta.tavg_1_n5.cov)
mean(Delta.tavg_1_n5-Delta.tavg_1_n5.cov) #0.002312954


#### Sensitivity to tmin_n1_n9 -------------------------

# no covariation
Delta.tmin_n1_n9=NULL
for(i in 1:100){
  max_lam<-kernel.fun(mean_tavg_1_n5,max_tmin_n1_n9,mean_tavg_17_12,i=i)
  min_lam<-kernel.fun(mean_tavg_1_n5,min_tmin_n1_n9,mean_tavg_17_12,i=i)
  
  Delta.tmin_n1_n9[i] <- abs((max_lam-min_lam)/((max_tmin_n1_n9-min_tmin_n1_n9)/sd_tmin_n1_n9))
}

hist(Delta.tmin_n1_n9)

# with covariation
Delta.tmin_n1_n9.cov=NULL
for(i in 1:100){
  max_lam<-kernel.fun(tavg_1_n5_when_tmin_n1_n9_max,max_tmin_n1_n9,tavg_17_12_when_tmin_n1_n9_max,i=i)
  min_lam<-kernel.fun(tavg_1_n5_when_tmin_n1_n9_min,min_tmin_n1_n9,tavg_17_12_when_tmin_n1_n9_min,i=i)
  
  Delta.tmin_n1_n9.cov[i] <- abs((max_lam-min_lam)/((max_tmin_n1_n9-min_tmin_n1_n9)/sd_tmin_n1_n9))
}

hist(Delta.tmin_n1_n9.cov)

# Differences (no cov - cov)
hist(Delta.tmin_n1_n9-Delta.tmin_n1_n9.cov)


#### Sensitivity to tavg_17_12 ------------------------------

# no covariation
Delta.tavg_17_12=NULL
for(i in 1:100){
  max_lam<-kernel.fun(mean_tavg_1_n5,mean_tmin_n1_n9,max_tavg_17_12,i=i)
  min_lam<-kernel.fun(mean_tavg_1_n5,mean_tmin_n1_n9,min_tavg_17_12,i=i)
  
  Delta.tavg_17_12[i] <- abs((max_lam-min_lam)/((max_tavg_17_12-min_tavg_17_12)/sd_tavg_17_12))
}

hist(Delta.tavg_17_12)

# with covariation
Delta.tavg_17_12.cov=NULL
for(i in 1:100){
  max_lam<-kernel.fun(tavg_1_n5_when_tavg_17_12_max,tmin_n1_n9_when_tavg_17_12_max,max_tavg_17_12,i=i)
  min_lam<-kernel.fun(tavg_1_n5_when_tavg_17_12_min,tmin_n1_n9_when_tavg_17_12_min,min_tavg_17_12,i=i)
  
  Delta.tavg_17_12.cov[i] <- abs((max_lam-min_lam)/((max_tavg_17_12-min_tavg_17_12)/sd_tavg_17_12))
}

hist(Delta.tavg_17_12.cov)

# Differences (no cov - cov)
hist(Delta.tavg_17_12 - Delta.tavg_17_12.cov)
mean(Delta.tavg_17_12 - Delta.tavg_17_12.cov)
```


## 5) Vital-rate-specific sensitivities
```{r}
# Sensitivity of Reproduction:
# The only climate covariate linked to reproduction is fp_tavg which is tavg_17_12
# Therefore, only perturb fp_tavg and leave the rest at mean values
# But this would be the same as Sensitivity to tavg_17_12 with no covariation
Sens.Reprod.tavg=NULL
for(i in 1:100){
  max_lam<-kernel.fun(mean_tavg_1_n5,mean_tmin_n1_n9,max_tavg_17_12,i=i)
  min_lam<-kernel.fun(mean_tavg_1_n5,mean_tmin_n1_n9,min_tavg_17_12,i=i)
  
  Sens.Reprod.tavg[i] <- abs((max_lam-min_lam)/((max_tavg_17_12-min_tavg_17_12)/sd_tavg_17_12))
}

hist(Sens.Reprod.tavg) # basically 0

# Sensitivity of Survival:
# The two climate covariates linked to survival are s_tavg = tavg_1_n5, and s_tmin
# So, perturb these two but leave the third covariate (fp_tavg) at its mean value

# First get sens to s_tavg (tavg_1_n5)
Sens.Surv.stavg=NULL
for(i in 1:100){
  max_lam<-kernel.fun(max_tavg_1_n5,tmin_n1_n9_when_tavg_1_n5_max,mean_tavg_17_12,i=i)
  min_lam<-kernel.fun(min_tavg_1_n5,tmin_n1_n9_when_tavg_1_n5_min,mean_tavg_17_12,i=i)
  
  Sens.Surv.stavg[i] <- abs((max_lam-min_lam)/((max_tavg_1_n5-min_tavg_1_n5)/sd_tavg_1_n5))
}

hist(Sens.Surv.stavg) 

# Now get sens to tmin (tmin_n1_n9)
Sens.Surv.tmin=NULL
for(i in 1:100){
  max_lam<-kernel.fun(tavg_1_n5_when_tmin_n1_n9_max,max_tmin_n1_n9,mean_tavg_17_12,i=i)
  min_lam<-kernel.fun(tavg_1_n5_when_tmin_n1_n9_min,min_tmin_n1_n9,mean_tavg_17_12,i=i)
  
  Sens.Surv.tmin[i] <- abs((max_lam-min_lam)/((max_tmin_n1_n9-min_tmin_n1_n9)/sd_tmin_n1_n9))
}

hist(Sens.Surv.tmin) 

```


## 6) Save output
```{r}
# save output of the sensitivity analysis from 4)
Results=data.frame(study.doi="https://doi.org/10.1111/gcb.15519",
                  year.of.publication="2021",
                  group="Plants",
                  species="Opuntia imbricata",
                  continent="North America",
                  driver="temperature",
                  driver.type="C",
                  stage.age="all",
                  vital.rates="all",
                  sens=c(Delta.tavg_1_n5,Delta.tavg_1_n5.cov,
                         Delta.tavg_17_12,Delta.tavg_17_12.cov,
                         Delta.tmin_n1_n9,Delta.tmin_n1_n9.cov),
                  cov=rep(c(0,1),100),
                  gen.time=9,
                  n.vr=3, # number of vital rates with covariates
                  n.pam=28, # number of total parameters of these vital rates
                  dens=0,
                  biotic_interactions=0)

write.csv(Results,"Sens_OpunitaImbricata.csv",row.names = F)

# save output of the vital-rate-specific sensitivity analysis from 5)
ResultsVR=data.frame(study.doi="https://doi.org/10.1111/gcb.15519",
                  year.of.publication="2021",
                  group="Plants",
                  species="Opuntia imbricata",
                  continent="North America",
                  driver="temperature",
                  driver.type="C",
                  stage.age="all", # could not perturb covariates in specific stages
                  vital.rates=rep(c("reproduction","survival","survival"),each=100),
                  sens=c(Sens.Reprod.tavg,
                         Sens.Surv.stavg,
                         Sens.Surv.tmin),
                  gen.time=9,
                  n.vr=3, # number of vital rates with covariates
                  n.pam=28, # number of total parameters of these vital rates
                  dens=0,
                  biotic_interactions=0)

write.csv(ResultsVR,"SensVR_OpuntiaImbricata.csv",row.names = F)

```







