---
title: "Sensitivity_Analysis_IPM"
author: "Esin Ickin"
date: "4/4/2023"
output:
  pdf_document: default
  html_document: default
---

## 0) Prepare session

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Files_Marmot")
```

```{r}
rm(list=ls())

library(boot)
library(popbio)

# perturbation functions
source("perturbation_func_marmots_MCMC/pert.Sj.winter.R")
source("perturbation_func_marmots_MCMC/pert.Sy.winter.R")
source("perturbation_func_marmots_MCMC/pert.Snra.winter.R")
source("perturbation_func_marmots_MCMC/pert.Sra.winter.R")

source("perturbation_func_marmots_MCMC/pert.Ty.winter.R")
source("perturbation_func_marmots_MCMC/pert.Tnra.winter.R")
source("perturbation_func_marmots_MCMC/pert.Tra.winter.R")

source("perturbation_func_marmots_MCMC/pert.Gj.winter.R") 
source("perturbation_func_marmots_MCMC/pert.Gy.winter.R")
source("perturbation_func_marmots_MCMC/pert.Gnra.winter.R")
source("perturbation_func_marmots_MCMC/pert.Gra.winter.R") 

source("perturbation_func_marmots_MCMC/pert.Sy.summer.R")
source("perturbation_func_marmots_MCMC/pert.Snra.summer.R")
source("perturbation_func_marmots_MCMC/pert.Sra.summer.R")

source("perturbation_func_marmots_MCMC/pert.Gy.summer.R")
source("perturbation_func_marmots_MCMC/pert.Gnra.summer.R")
source("perturbation_func_marmots_MCMC/pert.Gra.summer.R")

source("perturbation_func_marmots_MCMC/pert.Off.summer.R")
source("perturbation_func_marmots_MCMC/pert.Rec.summer.R")
```

## 1) Vital rates

The example is from a paper on yellow-bellied marmots (Paniw et al. 2020, <https://doi.org/10.1111/ele.13459>)

In the paper, we parameterized vital rates as a function of a latent climatic variable (Q). Because of the way the model was conceptualized, we considered random year variation as a separate covariate.

The vital rates are defined as functions, and we need the coefficients from vital rate models (here mean posterior values from Bayesian GLMs). We would appreciate a similar input on your part. Alternatively, it possible to have vital-rate models as an object, and then just use the "predict" function, instead of building our own functions.

```{r}
# Coefficients (can be supplied as a list; written as R objects; or directly integrated into the functions below)

# LOAD GLM BAYESIAN MODEL PARAMETERS
load("outputBay1_latent3.rda")

#mcmc=out1J$mean # use mean values here
mcmc=out1J$sims.list


#mcmc_mean=out1J$mean
#mcmc_mean[["a1.pr"]]

#str(mcmc[["a1.pr"]])

#mcmc[["a1.pr"]][55, 1]
#mcmc[["a1.pr"]][79, 3]

### Parameter are scaled - need to backscale for the vital rate predictions

m_S1=12.8968
sd_S1=2.23326

m_S2=12.62647
sd_S2=1.556634

m_PR=14.89485
sd_PR=0.8104184

m_R=13.74764
sd_R=0.7182092

### Rescale parameters

# Winter survival
for(i in 1:3000){
  mcmc[["bc.surv1"]][i]=mcmc[["bc.surv1"]][i]/sd_S1
  mcmc[["a0.surv1"]][i]=mcmc[["a0.surv1"]][i]-mcmc[["bc.surv1"]][i]*m_S1 

  # Summer survival 
  mcmc[["bc.surv2"]][i]=mcmc[["bc.surv2"]][i]/sd_S2
  mcmc[["a0.surv2"]][i]=mcmc[["a0.surv2"]][i]-mcmc[["bc.surv2"]][i]*m_S2
  
  # Transition to reproductive 
  mcmc[["bc.pr"]][i]=mcmc[["bc.pr"]][i]/sd_PR
  mcmc[["a0.pr"]][i]=mcmc[["a0.pr"]][i]-mcmc[["bc.pr"]][i]*m_PR

  # Number recruits 
  mcmc[["bc.rec"]][i]=mcmc[["bc.rec"]][i]/sd_R
  mcmc[["a0.rec"]][i]=mcmc[["a0.rec"]][i]-mcmc[["bc.rec"]][i]*m_R
  
}



# Functions

# The marmot model is seasonal, so we have survival and growth in winter and summer, and reproduction in summer

# The input covariates are: 

# z - mass
# stage - life-history stage: juvenile (J), yearling (Y), non-reproductive adult (NRA), reproductive adult (RA)
# year - one of 40 study years (random year effect)
# Q.sim - latent climatic variable

# SURVIVAL: AUGUST JUNE NEXT 

S1.fun <- function(z,stage,year,Q.sim,i) {
  
  if(year==0){
    
    mu.surv=exp(mcmc[["a0.surv1"]][i]+ mcmc[["bc.surv1"]][i]*z + mcmc[["a1.surv1"]][i,stage]+
                  mcmc[["bq.surv1"]][i]*Q.sim)
  }else{
    mu.surv=exp(mcmc[["a0.surv1"]][i]+ mcmc[["bc.surv1"]][i]*z + mcmc[["a1.surv1"]][i,stage]+ mcmc[["aY.surv1"]][i,year]+
                  mcmc[["bq.surv1"]][i]*Q.sim)
  }
  
  return(mu.surv/(1+mu.surv))
}


# SURVIVAL: JUNE AUGUST  
S2.fun <- function(z,stage,year,Q.sim,i) {
  
  if(year==0){
    
    mu.surv=exp(mcmc[["a0.surv2"]][i]+ mcmc[["bc.surv2"]][i]*z + mcmc[["a1.surv2"]][i,stage]+
                  mcmc[["bq.surv2"]][i]*Q.sim)
  }else{
    mu.surv=exp(mcmc[["a0.surv2"]][i]+ mcmc[["bc.surv2"]][i]*z + mcmc[["a1.surv2"]][i,stage]+ mcmc[["aY.surv2"]][i,year]+
                  mcmc[["bq.surv2"]][i]*Q.sim)
  }
  
  
  return(mu.surv/(1+mu.surv))
}

# GROWTH AUGUST TO JUNE (we assume a constant variance)

GR1.fun <- function(z,zz,stage,year,Q.sim,i){
  
  if(year==0){
    
    growth.mu=(mcmc[["a0.gr"]][i]+ (mcmc[["bc.gr"]][i]+mcmc[["bcstage.gr"]][i,stage])*z + mcmc[["a1.gr"]][i,stage]+ 
                 mcmc[["bq.gr"]][i]*Q.sim)
  }else{
    growth.mu=(mcmc[["a0.gr"]][i]+ (mcmc[["bc.gr"]][i]+mcmc[["bcstage.gr"]][i,stage])*z + mcmc[["a1.gr"]][i,stage]+ 
                 mcmc[["aY.gr"]][i,year]+
                 mcmc[["bq.gr"]][i]*Q.sim)
  }
  
  
  # Get residual variance 
  var.res= (mcmc[["sigma.gr"]][i])^2
  # Density distribution function of the normal distribution
  gr1 = sqrt(2*pi*var.res)
  gr2 = ((zz-growth.mu)^2)/(2*var.res)
  
  return(exp(-gr2)/gr1)
  
}

# GROWTH JUNE TO AUGUST (we assume a contant variance)

GR2.fun <- function(z,zz,stage,year,Q.sim,i){
  
  if(year==0){
    
    growth.mu=(mcmc[["a0.gr2"]][i]+ (mcmc[["bc.gr2"]][i]+mcmc[["bcstage.gr2"]][i,stage])*z + mcmc[["a1.gr2"]][i,stage]+ 
                 mcmc[["bq.gr2"]][i]*Q.sim)
    
  }else{
    growth.mu=(mcmc[["a0.gr2"]][i]+ (mcmc[["bc.gr2"]][i]+mcmc[["bcstage.gr2"]][i,stage])*z + mcmc[["a1.gr2"]][i,stage]+ 
                 mcmc[["aY.gr2"]][i,year]+
                 mcmc[["bq.gr2"]][i]*Q.sim)
    
  }
  
  
  # Get residual variance 
  var.res= (mcmc[["sigma.gr2"]][i])^2
  # Density distribution function of the normal distribution
  gr1 = sqrt(2*pi*var.res)
  gr2 = ((zz-growth.mu)^2)/(2*var.res)
  
  return(exp(-gr2)/gr1)
  
}


# PROBABILITY OF REPRODUCING:  

PR.fun <- function(z,stage,year,Q.sim,i) {
  
  if(year==0){
    
    mu.rep=exp(mcmc[["a0.pr"]][i]+ mcmc[["bc.pr"]][i]*z+ mcmc[["a1.pr"]][i,stage]+
                 mcmc[["bq.pr"]][i]*Q.sim)
    
  }else{
    mu.rep=exp(mcmc[["a0.pr"]][i]+ mcmc[["bc.pr"]][i]*z+ mcmc[["a1.pr"]][i,stage]+mcmc[["aY.pr"]][i,year]+
                 mcmc[["bq.pr"]][i]*Q.sim)
    
  }
  
  
  return(mu.rep/(1+mu.rep))
}

# Number of recruits:  

R.fun <- function(z,year,Q.sim,i) {
  
  if(year==0){
    
    mu.rec=exp(mcmc[["a0.rec"]][i]+ mcmc[["bc.rec"]][i]*z+
                 mcmc[["bq.rec"]][i]*Q.sim)
    
  }else{
    mu.rec=exp(mcmc[["a0.rec"]][i]+ mcmc[["bc.rec"]][i]*z+  mcmc[["aY.rec"]][i,year]+
                 mcmc[["bq.rec"]][i]*Q.sim)
    
  }
  
  
  
  
  return(mu.rec)
}

## OFFSPRING MASS

OffMass.fun <- function(z,zz,year,Q.sim,i){
  
  if(year==0){
    
    growth.mu=(mcmc[["a0.off"]][i]+ mcmc[["bc.off"]][i]*z+
                 mcmc[["bq.off"]][i]*Q.sim)
    
  }else{
    growth.mu=(mcmc[["a0.off"]][i]+ mcmc[["bc.off"]][i]*z+  mcmc[["aY.off"]][i,year]+
                 mcmc[["bq.off"]][i]*Q.sim)
    
  }
  
  
  # Get residual variance 
  var.res= (mcmc[["sigma.off"]][i])^2
  # Density distribution function of the normal distribution
  gr1 = sqrt(2*pi*var.res)
  gr2 = ((zz-growth.mu)^2)/(2*var.res)
  
  return(exp(-gr2)/gr1)
  
}
```

## 2) Covariates

There are the input data for the vital rate functions. Here, we have a time series of data. This is the best format because it allows us to calculate not only means and variances, but also covariances, the latter being important to calculate scaled sensitivities.

However, it is also possible to send us just the values we are interested in.

```{r}
mean_mcmc=out1J$mean

max.Q=max(mean_mcmc$Q)
min.Q=min(mean_mcmc$Q)
mean.Q=mean(mean_mcmc$Q)
sd.Q=sd(mean_mcmc$Q)

# Covariation
# Covariation
year_when_Q_max=which(mean_mcmc$Q==max.Q)
year_when_Q_min=which(mean_mcmc$Q==min.Q)


# ...

# This type of information is required for all (a)biotic covariates in the model (here we only have Q)
# For an example on how to provide more covariates (if you don't provide a time series), please see the MPM code example
```

## 3) Population model

Here, we use the vital rate function to construct an annual population model that can give us the population growth rate (lambda).

In the following example, the IPM is constructed with mean abiotic (Q) covariate values.

In the perturbations, the covariate values are changed.

Please note that this part is very specific to each study. Here, we create a stage-by-mass IPM for winter and summer, and then join them into an annual model.

```{r}
### Necessary IPM input parameters 

# Define the lower and upper integration limit
L=7.791458 # minimum observed mass (log grams)
U=17.07776 # maximum observed mass (log grams)

n=100 # bins

b <- L+c(0:n)*(U-L)/n # interval that each cell of the matrix covers 
z <- 0.5*(b[1:n]+b[2:(n+1)]) # midpoint
h=(U-L)/n # bin width 

n.stage=4

# We formulate the IPM construction as a function, so that we can use it later for the perturbations

winter_ipm <- function(year,env,i){
 
  IPMaj=array(0,c(n*n.stage,n*n.stage)) # winter
  
  ### AUGUST - JUNE 
  
  year = year
  Q.sim=env
  
  Sj <- diag(S1.fun(z,stage=1,year,Q.sim,i)) # Survival juveniles
  Sy <- diag(S1.fun(z,stage=2,year,Q.sim,i)) # Survival yearlings
  Snra <- diag(S1.fun(z,stage=3,year,Q.sim,i)) # Survival non-reproductive adults 
  Sra <- diag(S1.fun(z,stage=4,year,Q.sim,i)) # Survival reproductive adults 
  
  # Transition To RA or NRA 
  Ty <- diag(PR.fun(z,stage=1,year,Q.sim,i))
  Tnra <- diag(PR.fun(z,stage=2,year,Q.sim,i))
  Tra <- diag(PR.fun(z,stage=3,year,Q.sim,i))
  
  # Growth - stage specific like for survival
  G <- h*t(outer(z,z,GR1.fun,stage=1,year,Q.sim,i)) 
  # Control for eviction:
  # this is equivalent to redistributing evicted sizes evenly among existing size classes 
  G=G/matrix(as.vector(apply(G,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  # FILL IPM
  IPMaj[(n+1):(2*n),1:n]=G%*%Sj # Juvenile to Yearling
  
  G <- h*t(outer(z,z,GR1.fun,stage=2,year,Q.sim,i)) 
  # Control for eviction:
  # this is equivalent to redistributing evicted sizes evenly among existing size classes 
  G=G/matrix(as.vector(apply(G,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  # FILL IPM
  IPMaj[(2*n+1):(3*n),(n+1):(2*n)]=G%*%(Sy*diag(1-PR.fun(z,stage=1,year,Q.sim,i))) # Yearling to Non-Reproductive Adult
  IPMaj[(3*n+1):(4*n),(n+1):(2*n)]=G%*%(Sy*Ty) # Yearling to Reproductive Adult
  
  
  G <- h*t(outer(z,z,GR1.fun,stage=3,year,Q.sim,i)) 
  G=G/matrix(as.vector(apply(G,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  # FILL IPM
  IPMaj[(2*n+1):(3*n),(2*n+1):(3*n)]=G%*%(Snra*diag(1-PR.fun(z,stage=2,year,Q.sim,i))) # Non-Reproductive Adult to Non-Reproductive Adult
  IPMaj[(3*n+1):(4*n),(2*n+1):(3*n)]=G%*%(Snra*Tnra) # Non-Reproductive Adult to Reproductive Adult
  
  G <- h*t(outer(z,z,GR1.fun,stage=4,year,Q.sim,i)) 
  G=G/matrix(as.vector(apply(G,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  # FILL IPM
  IPMaj[(2*n+1):(3*n),(3*n+1):(4*n)]=G%*%(Sra*diag(1-PR.fun(z,stage=3,year,Q.sim,i))) # Reproductive Adult to non-Reproductive Adult
  IPMaj[(3*n+1):(4*n),(3*n+1):(4*n)]=G%*%(Sra*Tra) # Reproductive Adult to Reproductive Adult
  
  return(IPMaj)
}

########################################################

summer_ipm <- function(year,env,i){
  IPMja=array(0,c(n*n.stage,n*n.stage)) # summer
  year = year
  Q.sim=env
  
  # SURVIVAL 
  S2y <- diag(S2.fun(z,stage=1,year,Q.sim,i),nrow=n,ncol=n) # Survival yearlings
  S2nra <- diag(S2.fun(z,stage=2,year,Q.sim,i),nrow=n,ncol=n) # Survival non-reproductive adults 
  S2ra <- diag(S2.fun(z,stage=3,year,Q.sim,i),nrow=n,ncol=n) # Survival reproductive adults 
  
  # GROWTH 
  
  # Yearlings 
  Gy <- h*t(outer(z,z,GR2.fun,stage=1,year,Q.sim,i)) 
  
  # Reproductive Adults 
  Gnra <- h*t(outer(z,z,GR2.fun,stage=2,year,Q.sim,i)) 
  
  # Non-Reproductive Adults 
  Gra <- h*t(outer(z,z,GR2.fun,stage=3,year,Q.sim,i)) 
  
  # Control for eviction:
  # this is equivalent to redistributing evictd sizes evenly among existing size classes 
  
  Gy=Gy/matrix(as.vector(apply(Gy,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  Gra=Gra/matrix(as.vector(apply(Gra,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  Gnra=Gnra/matrix(as.vector(apply(Gnra,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  
  
  # Offspring mass 
  
  OffMass <- h*t(outer(z,z,OffMass.fun,year,Q.sim,i)) 
  # Control for eviction:
  # this is equivalent to redistributing evictd sizes evenly among existing size classes 
  OffMass=OffMass/matrix(as.vector(apply(OffMass,2,sum)),nrow=n,ncol=n,byrow=TRUE)
  
  # recruitment (as individuals that die in August still produce offspring, recruitment does not need to be multiplied by S2)
  
  Rec=(diag(R.fun(z,year,Q.sim,i)))/2
  
  Fkernel <- as.matrix(OffMass%*%(Rec*S2ra))
  
  # FILL IPM
  
  IPMja[(n+1):(2*n),(n+1):(2*n)]=Gy%*%S2y # Yearling stay Yearling
  IPMja[(2*n+1):(3*n),(2*n+1):(3*n)]=Gnra%*%S2nra # Non-Reproductive Adults stay Non-Reproductive Adults
  IPMja[(3*n+1):(4*n),(3*n+1):(4*n)]=Gra%*%S2ra # Reproductive Adults stay Reproductive Adults
  
  IPMja[1:n,(3*n+1):(4*n)]=Fkernel # Adults producing juveniles
  
  return(IPMja)
}

#### Asymptotic lambda at mean environmental values
lam=NULL

for(i in 1:100){
  
  K = summer_ipm(0,mean.Q,i)%*%winter_ipm(0,mean.Q,i)
  lam[i]=lambda(K)
}

  hist(lam)
```

## 4) Scaled sensitivity analyses

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: <https://doi.org/10.1073/pnas.1918363117>)

Note that this is a step that Esin will implement in her MS thesis. With the information given in 1-3, we should be able to run these analyses.

```{r}
delta.Q=NULL
delta.Q.cov=NULL

for(i in 1:100){
  #1. Sensitivity to Q assuming 0 year effect 

  max.ipm=summer_ipm(0,max.Q,i)%*%winter_ipm(0,max.Q,i)
  min.ipm=summer_ipm(0,min.Q,i)%*%winter_ipm(0,min.Q,i)

  delta.Q[i]=abs((lambda(max.ipm)-lambda(min.ipm))/((max.Q-min.Q)/sd.Q))

  # 2. Sensitivity to Q assuming covariation among covariates
  max.ipm=summer_ipm(year_when_Q_max,max.Q,i)%*%winter_ipm(year_when_Q_max,max.Q,i)
  min.ipm=summer_ipm(year_when_Q_min,min.Q,i)%*%winter_ipm(year_when_Q_min,min.Q,i)

  delta.Q.cov[i]=abs((lambda(max.ipm)-lambda(min.ipm))/((max.Q-min.Q)/sd.Q))
}

hist(delta.Q-delta.Q.cov) # should be positive but it's not :(

```

## 5) Scaled sensitivities per vital rate

```{r}
delta.Sj.winter=NULL
delta.Sy.winter=NULL
delta.Snra.winter=NULL
delta.Sra.winter=NULL
delta.Sy.summer=NULL
delta.Snra.summer=NULL
delta.Sra.summer=NULL

delta.Gj.winter=NULL
delta.Gy.winter=NULL
delta.Gnra.winter=NULL
delta.Gra.winter=NULL

delta.Ty.winter=NULL
delta.Tnra.winter=NULL
delta.Tra.winter=NULL

delta.Off.summer=NULL
delta.Rec.summer=NULL

for(i in 1:100){
  # WINTER SURVIVAL
# juveniles
max.pert.Sj.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSj_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Sj.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSj_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Sj.winter[i]=abs((lambda(max.pert.Sj.winter)-lambda(min.pert.Sj.winter))/((max.Q-min.Q)/sd.Q))

# yearlings
max.pert.Sy.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSy_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Sy.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSy_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Sy.winter[i]=abs((lambda(max.pert.Sy.winter)-lambda(min.pert.Sy.winter))/((max.Q-min.Q)/sd.Q))

# non-reproductive adults
max.pert.Snra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSnra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Snra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSnra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Snra.winter[i]=abs((lambda(max.pert.Snra.winter)-lambda(min.pert.Snra.winter))/((max.Q-min.Q)/sd.Q))

# reproductive adults
max.pert.Sra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Sra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertSra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Sra.winter[i]=abs((lambda(max.pert.Sra.winter)-lambda(min.pert.Sra.winter))/((max.Q-min.Q)/sd.Q))

# SUMMER SURVIVAL
# no juvenile summer survival

# yearlings
max.pert.Sy.summer=summer_pertSy_ipm(0,mean.Q,max.Q,year_when_Q_max,i)%*%winter_ipm(0,mean.Q,i)
min.pert.Sy.summer=summer_pertSy_ipm(0,mean.Q,min.Q,year_when_Q_min,i)%*%winter_ipm(0,mean.Q,i)

delta.Sy.summer[i]=abs((lambda(max.pert.Sy.summer)-lambda(min.pert.Sy.summer))/((max.Q-min.Q)/sd.Q))

# non-reproductive adults
max.pert.Snra.summer=summer_pertSnra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)%*%winter_ipm(0,mean.Q,i)
min.pert.Snra.summer=summer_pertSnra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)%*%winter_ipm(0,mean.Q,i)

delta.Snra.summer[i]=abs((lambda(max.pert.Snra.summer)-lambda(min.pert.Snra.summer))/((max.Q-min.Q)/sd.Q))

# reproductive adults
max.pert.Sra.summer=summer_pertSra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)%*%winter_ipm(0,mean.Q,i)
min.pert.Sra.summer=summer_pertSra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)%*%winter_ipm(0,mean.Q,i)

delta.Sra.summer[i]=abs((lambda(max.pert.Sra.summer)-lambda(min.pert.Sra.summer))/((max.Q-min.Q)/sd.Q))

# WINTER GROWTH
# juveniles
max.pert.Gj.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGj_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Gj.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGj_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Gj.winter[i]=abs((lambda(max.pert.Gj.winter)-lambda(min.pert.Gj.winter))/((max.Q-min.Q)/sd.Q))

# yearlings
max.pert.Gy.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGy_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Gy.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGy_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Gy.winter[i]=abs((lambda(max.pert.Gy.winter)-lambda(min.pert.Gy.winter))/((max.Q-min.Q)/sd.Q))

# non-reproductive adults
max.pert.Gnra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGnra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Gnra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGnra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Gnra.winter[i]=abs((lambda(max.pert.Gnra.winter)-lambda(min.pert.Gnra.winter))/((max.Q-min.Q)/sd.Q))

# reproductive adults
max.pert.Gra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Gra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertGra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Gra.winter[i]=abs((lambda(max.pert.Gra.winter)-lambda(min.pert.Gra.winter))/((max.Q-min.Q)/sd.Q))

# winter TRANSITION
# yearlings
max.pert.Ty.winter=summer_ipm(0,mean.Q,i)%*%winter_pertTy_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Ty.winter=summer_ipm(0,mean.Q,i)%*%winter_pertTy_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Ty.winter[i]=abs((lambda(max.pert.Ty.winter)-lambda(min.pert.Ty.winter))/((max.Q-min.Q)/sd.Q))

# non-reproductive adults
max.pert.Tnra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertTnra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Tnra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertTnra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Tnra.winter[i]=abs((lambda(max.pert.Tnra.winter)-lambda(min.pert.Tnra.winter))/((max.Q-min.Q)/sd.Q))

# reproductive adults
max.pert.Tra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertTra_ipm(0,mean.Q,max.Q,year_when_Q_max,i)
min.pert.Tra.winter=summer_ipm(0,mean.Q,i)%*%winter_pertTra_ipm(0,mean.Q,min.Q,year_when_Q_min,i)

delta.Tra.winter[i]=abs((lambda(max.pert.Tra.winter)-lambda(min.pert.Tra.winter))/((max.Q-min.Q)/sd.Q))

# summer offspring mass
max.pert.Off.summer=summer_pertOff_ipm(0,mean.Q,max.Q,year_when_Q_max,i)%*%winter_ipm(0,mean.Q,i)
min.pert.Off.summer=summer_pertOff_ipm(0,mean.Q,min.Q,year_when_Q_min,i)%*%winter_ipm(0,mean.Q,i)

delta.Off.summer[i]=abs((lambda(max.pert.Off.summer)-lambda(min.pert.Off.summer))/((max.Q-min.Q)/sd.Q))

# summer recruitment
max.pert.Rec.summer=summer_pertRec_ipm(0,mean.Q,max.Q,year_when_Q_max,i)%*%winter_ipm(0,mean.Q,i)
min.pert.Rec.summer=summer_pertRec_ipm(0,mean.Q,min.Q,year_when_Q_min,i)%*%winter_ipm(0,mean.Q,i)

delta.Rec.summer[i]=abs((lambda(max.pert.Rec.summer)-lambda(min.pert.Rec.summer))/((max.Q-min.Q)/sd.Q))
  
}

sens.per.vr.winter=data.frame(sens=c(delta.Gj.winter,delta.Gy.winter,delta.Gnra.winter,delta.Gra.winter,delta.Sj.winter,delta.Sy.winter,delta.Snra.winter,delta.Sra.winter,delta.Ty.winter,delta.Tnra.winter,delta.Tra.winter),driver=c("winter Q"),vital.rate=rep(c("Gj","Gy","Gnra","Gra","Sj","Sy","Snra","Sra","Ty","Tnra","Tra"),each=100))

sens.per.vr.summer=data.frame(sens=c(delta.Sy.summer,delta.Snra.summer,delta.Sra.summer,delta.Off.summer,delta.Rec.summer),
                              driver=c("summer Q"),vital.rate=c("Sy","Snra","Sra","OffspringMass","Recruitment"))

sens.per.vr=rbind(sens.per.vr.winter,sens.per.vr.summer)

all.sens=data.frame(sens=c(delta.Q,delta.Q.cov),
                           driver=c("Q"),vital.rate=rep(c("All","All.cov"),each=100))

all.sens=rbind(all.sens,sens.per.vr)

write.csv(all.sens,"AllSens_Marmots_MCMC.csv") # this file is then reorganized in the "ReOrganize.R" script for later analyses

sens.per.vr2=data.frame(Gj.winter=delta.Gj.winter,Gnra.winter=delta.Gnra.winter,Gra.winter=delta.Gra.winter,Gy.winter=delta.Gy.winter,Off.summer=delta.Off.summer,All.vr=delta.Q,All.vr.cov=delta.Q.cov,Rec.summer=delta.Rec.summer,Sj.winter=delta.Sj.winter,Snra.summer=delta.Snra.summer,Snra.winter=delta.Snra.winter,Sra.winter=delta.Sra.winter,Sra.summer=delta.Sra.summer,Sy.winter=delta.Sy.winter,Sy.summer=delta.Sy.summer,Tnra.winter=delta.Tnra.winter,Tra.winter=delta.Tra.winter,Ty.winter=delta.Ty.winter)
write.csv(sens.per.vr2,"AllSens2_Marmots_MCMC.csv")
#df2=data.frame(sens=c(delta.Sj.winter,delta.Sy.winter,delta.Snra.winter,delta.Sra.winter,
#                       delta.Gj.winter,delta.Gy.winter,delta.Gnra.winter,delta.Gra.winter,
#                       delta.Ty.winter,delta.Tnra.winter,delta.Tra.winter),
#                vital.rate=c("Sj","Sy","Snra","Sra","Gj","Gy","Gnra","Gra","Ty","Tnra","Tra"),
#               driver=c("winter Q"))

#df3=data.frame(sens=c(delta.Sy.summer,delta.Snra.summer,delta.Sra.summer,
#                      delta.Rec.summer,delta.Off.summer),
#               vital.rate=c("Sy","Snra","Sra","Rec","Offspring mass"),
#               driver=c("summer Q"))

#df4=rbind(df2,df3)
#write.csv(df4,"AllScaledSens.csv")
```
