---
title: "SensitivityAnalysis_MPM"
author: "Esin Ickin"
date: "2/23/2023"
output: html_document
---

This code is an example on how to parameterize vital-rate functions using coefficients and covariate inputs to then build a matrix population model and compute sensitivities.

## 0) Prepare session
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/OneDrive - Universität Zürich UZH/Master Thesis/pert_analyses/mouse lemurs")
```

```{r}
library(boot)
library(popbio)
library(MASS)
rm(list=ls())
set.seed(1234)
```

## 1) Vital rates

The example is from a paper on gray mouse lemurs: Ozgul et al. 2023
In the paper, we parameterized five vital rates as a function of population density, rainfall, and temperature.

The vital rates are defined as functions, and we need the coefficients from vital rate models (here GLMs). We would appreciate a similar input on your part. Alternatively, it possible to have GLM models as an object, and then just use the "predict" function, instead of building out own functions. 

```{r}
# Female adult survival
intercept.sfa=rnorm(10000,mean=9.95,sd=4.153)
coef_dens.sfa=rnorm(10000,mean=0.04,sd=0.013)
coef_rain.sfa=rnorm(10000,mean=0.0031,sd=0.001)
coef_temp.sfa=rnorm(10000,mean=-0.38,sd=0.135)
coef_int.sfa=rnorm(10000,mean=-0.00005,sd=0.00001) #densrain

# Female juvenile survival
intercept.sfj=rnorm(10000,mean=9.95,sd=4.153)-rnorm(10000,mean=-1.86,sd=0.356)
coef_dens.sfj=rnorm(10000,mean=0.04,sd=0.013)
coef_densjuv.sfj=rnorm(10000,mean=0.01,sd=0.004)
coef_rain.sfj=rnorm(10000,mean=0.0031,sd=0.001)
coef_temp.sfj=rnorm(10000,mean=-0.38,sd=0.135)
coef_int.sfj=rnorm(10000,mean=-0.00005,sd=0.00001)

# Male adult survival
intercept.sma=rnorm(10000,mean=9.95,sd=4.153)-rnorm(10000,mean=-1.74,sd=0.465)
coef_dens.sma=rnorm(10000,mean=0.04,sd=0.013)
coef_rain.sma=rnorm(10000,mean=0.0031,sd=0.001)
coef_rainsex.sma=rnorm(10000,mean=0.0013,sd=0.0005)
coef_temp.sma=rnorm(10000,mean=-0.38,sd=0.135)
coef_int.sma=rnorm(10000,mean=-0.00005,sd=0.00001)

# Male juvenile survival
intercept.smj=rnorm(10000,mean=9.95,sd=4.153)-rnorm(10000,mean=-1.74,sd=0.465)-rnorm(10000,mean=-1.86,sd=0.356)  
coef_dens.smj=rnorm(10000,mean=0.04,sd=0.013)
coef_densjuv.smj=rnorm(10000,mean=0.01,sd=0.004)
coef_rain.smj=rnorm(10000,mean=0.0031,sd=0.001)
coef_rainsex.smj=rnorm(10000,mean=0.0013,sd=0.0005)
coef_temp.smj=rnorm(10000,mean=-0.38,sd=0.135)
coef_int.smj=rnorm(10000,mean=-0.00005,sd=0.00001)

# Recruitment
intercept.rec=rnorm(10000,mean=-10.97,sd=5.32)
coef_dens.rec=rnorm(10000,mean=-0.024,sd=0.009)
coef_temp.rec=rnorm(10000,mean=0.364,sd=0.171)

# Functions
# Female adult survival
Sfa <- function(dens,rain,temp,i){
  surv <- inv.logit(
    intercept.sfa[i] + coef_dens.sfa[i]*dens + coef_rain.sfa[i]*rain + coef_temp.sfa[i]*temp + coef_int.sfa[i]*dens*rain
  )
  return(surv)
}

# Female juvenile survival
Sfj <- function(dens,rain,temp,i){
  surv <- inv.logit(
    intercept.sfj[i] + coef_dens.sfj[i]*dens + coef_densjuv.sfj[i]*dens + coef_rain.sfj[i]*rain + coef_temp.sfj[i]*temp + coef_int.sfj[i]*dens*rain
  )
  return(surv)
}

# Male adult survival
Sma <- function(dens,rain,temp,i){
  surv <- inv.logit(
    intercept.sma[i] + coef_dens.sma[i]*dens + coef_rain.sma[i]*rain + coef_rainsex.sma[i] + coef_temp.sma[i]*temp + coef_int.sma[i]*dens*rain
  )
  return(surv)
}

# Male juvenile survival
Smj <- function(dens,rain,temp,i){
  surv <- inv.logit(
    intercept.smj[i] + coef_dens.smj[i]*dens + coef_densjuv.smj[i] + coef_rain.smj[i]*rain + coef_rainsex.smj[i]*rain + coef_temp.smj[i]*temp + coef_int.smj[i]*dens*rain
  )
  return(surv)
}

# Recruitment
rec <- function(dens,temp,i){
  Rec <- exp(
    intercept.rec[i] + coef_dens.rec[i]*dens + coef_temp.rec[i]*temp
  )
  return(Rec)
}

```

## 2) Covariates

There are the input data for the vital rate functions. Here, we have a time series of data. This is the best format because it allows us to calculate not only means and variances, but also covariances, the latter being important to calculate scaled sensitivities. 

However, it is also possible to send us just the values we are interested in.

```{r}
cov=read.csv("covariates.csv") # raw data

# Rain
max.rain=max(cov$rain) # here and elsewhere, this can be calculated from the raw data or a single value can be provided
min.rain=min(cov$rain)
mean.rain=mean(cov$rain)
sd.rain=sd(cov$rain)

# Covariation
temp_when_rain_max=cov$tmaxc[which(cov$rain==max(cov$rain))][1]
temp_when_rain_min=cov$tmaxc[which(cov$rain==min(cov$rain))][1]

dens_when_rain_max=cov$pop[which(cov$rain==max(cov$rain))][1]
dens_when_rain_min=cov$pop[which(cov$rain==min(cov$rain))][1]

# Temperature
max.temp=max(cov$tmaxc)
min.temp=min(cov$tmaxc)
mean.temp=mean(cov$tmaxc)
sd.temp=sd(cov$tmaxc)

# Covariation
dens_when_temp_max=cov$pop[which(cov$tmaxc==max(cov$tmaxc))][1]
dens_when_temp_min=cov$pop[which(cov$tmaxc==min(cov$tmaxc))][1]

rain_when_temp_max=cov$rain[which(cov$tmaxc==max(cov$tmaxc))][1]
rain_when_temp_min=cov$rain[which(cov$tmaxc==min(cov$tmaxc))][1]

# Density
max.dens=max(cov$pop)
min.dens=min(cov$pop)
mean.dens= mean(cov$pop)
sd.dens=sd(cov$pop)

# Covariation
rain_when_dens_max=cov$rain[which(cov$pop==max(cov$pop))][1]
rain_when_dens_min=cov$rain[which(cov$pop==min(cov$pop))][1]

temp_when_dens_max=cov$tmaxc[which(cov$pop==max(cov$pop))][1]
temp_when_dens_min=cov$tmaxc[which(cov$pop==min(cov$pop))][1]
```

## 3) Population model

Here, we use the vital rate function to construct an annual population model that can give us the population growth rate (lambda).

In the following example, the MPM is constructed with mean covariate values.
In the perturbations, the covariate values are changed.

```{r}
n.lambda=NULL
n.stage=4

for(i in 1:100){
  mpm = matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  n.lambda[i]=lambda(mpm)
}
#n.lambda=n.lambda[n.lambda > 0 & n.lambda < 2]
hist(n.lambda)
length(n.lambda)
max(n.lambda)
min(n.lambda)
mean(n.lambda)

# filter lambda <3
n.lambda = n.lambda[n.lambda <= 3]
#n.lambda
hist(n.lambda)
```

## 4) Sensitivtiy analysis

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117)

With the information given in 1-3, we should be able to run these analyses.

```{r}
# empty vectors for lambdas
# rain
l.max.R=NULL
l.min.R=NULL
l.max.R2=NULL
l.min.R2=NULL

# temp
l.max.T=NULL
l.min.T=NULL
l.max.T2=NULL
l.min.T2=NULL

# dens
l.max.D=NULL
l.min.D=NULL
l.max.D2=NULL
l.min.D2=NULL

# empty vectors for sensitivities
# rain
delta.R=NULL
deltaR.cov=NULL

# temp
delta.T=NULL
deltaT.cov=NULL

# dens
delta.D=NULL
deltaD.cov=NULL

# empty vectors for log ratios
# rain
l_ratio.R=NULL
l_ratio.R.cov=NULL

# temp
l_ratio.T=NULL
l_ratio.T.cov=NULL

# dens
l_ratio.D=NULL
l_ratio.D.cov=NULL


for(i in 1:1000){
  # RAIN
  # 1. Sensitivity to rain assuming mean temperature and mean density
  mpm.max = matrix(c(0,Smj(mean.dens,max.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,max.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,max.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,max.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,max.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,max.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,max.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,max.rain,mean.temp,i)),n.stage,n.stage)
  
  mpm.min = matrix(c(0,Smj(mean.dens,min.rain,mean.temp,i),0,0,
                   0,Sma(mean.dens,min.rain,mean.temp,i),0,0,
                   0.5*Sfj(mean.dens,min.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,min.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,min.rain,mean.temp,i),
                   0.5*Sfa(mean.dens,min.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,min.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,min.rain,mean.temp,i)),n.stage,n.stage)
  
  l.max.R[i]=lambda(mpm.max)
  l.min.R[i]=lambda(mpm.min)
 
  # 2. Sensitivity to rain assuming covariation among covariates
  mpm.max = matrix(c(0,Smj(dens_when_rain_max,max.rain,temp_when_rain_max,i),0,0,
                   0,Sma(dens_when_rain_max,max.rain,temp_when_rain_max,i),0,0,
                   0.5*Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),0,0.5*Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i),
                   0.5*Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),0,0.5*Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)),n.stage,n.stage)
  
  mpm.min = matrix(c(0,Smj(dens_when_rain_min,min.rain,temp_when_rain_min,i),0,0,
                   0,Sma(dens_when_rain_min,min.rain,temp_when_rain_min,i),0,0,
                   0.5*Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),0,0.5*Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i),
                   0.5*Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),0,0.5*Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)),n.stage,n.stage)
  
  l.max.R2[i]=lambda(mpm.max)
  l.min.R2[i]=lambda(mpm.min)
  

  
  # TEMPERATURE
  # 1. Sensitivity to temperature assuming mean rain and mean density
  mpm.max = matrix(c(0,Smj(mean.dens,mean.rain,max.temp,i),0,0,
                   0,Sma(mean.dens,mean.rain,max.temp,i),0,0,
                   0.5*Sfj(mean.dens,mean.rain,max.temp,i)*rec(mean.dens,max.temp,i),0,0.5*Sfj(mean.dens,mean.rain,max.temp,i)*rec(mean.dens,max.temp,i),Sfj(mean.dens,mean.rain,max.temp,i),
                   0.5*Sfa(mean.dens,mean.rain,max.temp,i)*rec(mean.dens,max.temp,i),0,0.5*Sfa(mean.dens,mean.rain,max.temp,i)*rec(mean.dens,max.temp,i),Sfa(mean.dens,mean.rain,max.temp,i)),n.stage,n.stage)
  
  mpm.min = matrix(c(0,Smj(mean.dens,mean.rain,min.temp,i),0,0,
                   0,Sma(mean.dens,mean.rain,min.temp,i),0,0,
                   0.5*Sfj(mean.dens,mean.rain,min.temp,i)*rec(mean.dens,min.temp,i),0,0.5*Sfj(mean.dens,mean.rain,min.temp,i)*rec(mean.dens,min.temp,i),Sfj(mean.dens,mean.rain,min.temp,i),
                   0.5*Sfa(mean.dens,mean.rain,min.temp,i)*rec(mean.dens,min.temp,i),0,0.5*Sfa(mean.dens,mean.rain,min.temp,i)*rec(mean.dens,min.temp,i),Sfa(mean.dens,mean.rain,min.temp,i)),n.stage,n.stage)
  
  l.max.T[i]=lambda(mpm.max)
  l.min.T[i]=lambda(mpm.min)
  
  
  # 2. Sensitivity to temperature assuming covariation among covariates
  mpm.max = matrix(c(0,Smj(dens_when_temp_max,rain_when_temp_max,max.temp,i),0,0,
                   0,Sma(dens_when_temp_max,rain_when_temp_max,max.temp,i),0,0,
                   0.5*Sfj(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(dens_when_temp_max,max.temp,i),0,0.5*Sfj(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(dens_when_temp_max,max.temp,i),Sfj(dens_when_temp_max,rain_when_temp_max,max.temp,i),
                   0.5*Sfa(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(dens_when_temp_max,max.temp,i),0,0.5*Sfa(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(dens_when_temp_max,max.temp,i),Sfa(dens_when_temp_max,rain_when_temp_max,max.temp,i)),n.stage,n.stage)
  
  mpm.min = matrix(c(0,Smj(dens_when_temp_min,rain_when_temp_min,min.temp,i),0,0,
                   0,Sma(dens_when_temp_min,rain_when_temp_min,min.temp,i),0,0,
                   0.5*Sfj(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(dens_when_temp_min,min.temp,i),0,0.5*Sfj(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(dens_when_temp_min,min.temp,i),Sfj(dens_when_temp_min,rain_when_temp_min,min.temp,i),
                   0.5*Sfa(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(dens_when_temp_min,min.temp,i),0,0.5*Sfa(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(dens_when_temp_min,min.temp,i),Sfa(dens_when_temp_min,rain_when_temp_min,min.temp,i)),n.stage,n.stage)
  
  l.max.T2[i]=lambda(mpm.max)
  l.min.T2[i]=lambda(mpm.min)
  

  
  # DENSITY
  # 1. Sensitivity to density assuming mean rain and mean temperature
  mpm.max = matrix(c(0,Smj(max.dens,mean.rain,mean.temp,i),0,0,
                   0,Sma(max.dens,mean.rain,mean.temp,i),0,0,
                   0.5*Sfj(max.dens,mean.rain,mean.temp,i)*rec(max.dens,mean.temp,i),0,0.5*Sfj(max.dens,mean.rain,mean.temp,i)*rec(max.dens,mean.temp,i),Sfj(max.dens,mean.rain,mean.temp,i),
                   0.5*Sfa(max.dens,mean.rain,mean.temp,i)*rec(max.dens,mean.temp,i),0,0.5*Sfa(max.dens,mean.rain,mean.temp,i)*rec(max.dens,mean.temp,i),Sfa(max.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
  mpm.min = matrix(c(0,Smj(min.dens,mean.rain,mean.temp,i),0,0,
                   0,Sma(min.dens,mean.rain,mean.temp,i),0,0,
                   0.5*Sfj(min.dens,mean.rain,mean.temp,i)*rec(min.dens,mean.temp,i),0,0.5*Sfj(min.dens,mean.rain,mean.temp,i)*rec(min.dens,mean.temp,i),Sfj(min.dens,mean.rain,mean.temp,i),
                   0.5*Sfa(min.dens,mean.rain,mean.temp,i)*rec(min.dens,mean.temp,i),0,0.5*Sfa(min.dens,mean.rain,mean.temp,i)*rec(min.dens,mean.temp,i),Sfa(min.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
  l.max.D[i]=lambda(mpm.max)
  l.min.D[i]=lambda(mpm.min)
  

  
  # 2. Sensitivity to density assuming covariation among covariates
  mpm.max = matrix(c(0,Smj(max.dens,rain_when_dens_max,temp_when_dens_max,i),0,0,
                   0,Sma(max.dens,rain_when_dens_max,temp_when_dens_max,i),0,0,
                   0.5*Sfj(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(max.dens,temp_when_dens_max,i),0,0.5*Sfj(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(max.dens,temp_when_dens_max,i),Sfj(max.dens,rain_when_dens_max,temp_when_dens_max,i),
                   0.5*Sfa(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(max.dens,temp_when_dens_max,i),0,0.5*Sfa(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(max.dens,temp_when_dens_max,i),Sfa(max.dens,rain_when_dens_max,temp_when_dens_max,i)),n.stage,n.stage)
  
  mpm.min = matrix(c(0,Smj(min.dens,rain_when_dens_min,temp_when_dens_min,i),0,0,
                   0,Sma(min.dens,rain_when_dens_min,temp_when_dens_min,i),0,0,
                   0.5*Sfj(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(min.dens,temp_when_dens_min,i),0,0.5*Sfj(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(min.dens,temp_when_dens_min,i),Sfj(min.dens,rain_when_dens_min,temp_when_dens_min,i),
                   0.5*Sfa(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(min.dens,temp_when_dens_min,i),0,0.5*Sfa(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(min.dens,temp_when_dens_min,i),Sfa(min.dens,rain_when_dens_min,temp_when_dens_min,i)),n.stage,n.stage)
  
  l.max.D2[i]=lambda(mpm.max)
  l.min.D2[i]=lambda(mpm.min)
  
}

# filter lambdas & calculate sensitivities ###############

# RAIN
# filter:
filter.maxmin.R=unique(c(which(l.max.R>3),which(l.min.R>3)))  
l.max.R=l.max.R[-filter.maxmin.R]
l.min.R=l.min.R[-filter.maxmin.R]

filter.maxmin.R2=unique(c(which(l.max.R2>3),which(l.min.R2>3)))
l.max.R2=l.max.R2[-filter.maxmin.R2]
l.min.R2=l.min.R2[-filter.maxmin.R2]

# sensitivities:
delta.R=abs((l.max.R-l.min.R)/((max.rain-min.rain)/sd.rain))
deltaR.cov=abs((l.max.R2-l.min.R2)/((max.rain-min.rain)/sd.rain))

# log ratios of lambdas
l_ratio.R=abs(log(l.max.R/l.min.R))
l_ratio.R.cov=abs(log(l.max.R2/l.min.R2))

# TEMP
# filter:
filter.maxmin.T=unique(c(which(l.max.T>3),which(l.min.T>3)))  
l.max.T=l.max.T[-filter.maxmin.T]
l.min.T=l.min.T[-filter.maxmin.T]

filter.maxmin.T2=unique(c(which(l.max.T2>3),which(l.min.T2>3)))
l.max.T2=l.max.T2[-filter.maxmin.T2]
l.min.T2=l.min.T2[-filter.maxmin.T2]

# sensitivities:
delta.T=abs((l.max.T-l.min.T)/((max.temp-min.temp)/sd.temp))
deltaT.cov=abs((l.max.T2-l.min.T2)/((max.temp-min.temp)/sd.temp))

# log ratios of lambdas
l_ratio.T=abs(log(l.max.T/l.min.T))
l_ratio.T.cov=abs(log(l.max.T2/l.min.T2))

# DENS
# filter:
filter.maxmin.D=unique(c(which(l.max.D>3),which(l.min.D>3)))  
l.max.D=l.max.D[-filter.maxmin.D]
l.min.D=l.min.D[-filter.maxmin.D]

filter.maxmin.D2=unique(c(which(l.max.D2>3),which(l.min.D2>3)))
l.max.D2=l.max.D2[-filter.maxmin.D2]
l.min.D2=l.min.D2[-filter.maxmin.D2]

# sensitivities:
delta.D=abs((l.max.D-l.min.D)/((max.dens-min.dens)/sd.dens))
deltaD.cov=abs((l.max.D2-l.min.D2)/((max.dens-min.dens)/sd.dens))

# log ratios of lambdas
l_ratio.D=abs(log(l.max.D/l.min.D))
l_ratio.D.cov=abs(log(l.max.D2/l.min.D2))

# SAVE OUTPUT #################
# also only get the first 100 iterations so that it matches with the rows
# quick save
df=data.frame(delta.R=delta.R[1:100],
              deltaR.cov=deltaR.cov[1:100],
              delta.T=delta.T[1:100],
              deltaT.cov=deltaT.cov[1:100],
              delta.D=delta.D[1:100],
              deltaD.cov=deltaD.cov[1:100],
              
              l_ratio.R=l_ratio.R[1:100],
              l_ratio.R.cov=l_ratio.R.cov[1:100],
              l_ratio.T=l_ratio.T[1:100],
              l_ratio.T.cov=l_ratio.T.cov[1:100],
              l_ratio.D=l_ratio.D[1:100],
              l_ratio.D.cov=l_ratio.D.cov[1:100]
              )
write.csv(df, "quicksave_mouselemur.csv",row.names = F)

# actual save
new_df <- data.frame(study.doi="10.1073/pnas.2214244120",
                     year.of.publication="2023",
                     group="Mammals",
                     species="Microcebus murinus",
                     continent="Africa",
                     driver=rep(c("rain","temperature","density"),each=200),
                     driver.type=rep(c("C","C","D"),each=200),
                     stage.age="all",
                     vital.rates="all",
                     sens=c(df$delta.R,df$deltaR.cov,
                            df$delta.T,df$deltaT.cov,
                            df$delta.D,df$deltaD.cov),
                     cov=rep(c(0,1),each=100),
                     mat=1, # age at sexual maturity Ozgul et al. 2023
                     n.vr=5, # number of vital rates with covariates
                     n.pam=27, # number of parameters
                     dens=1, # density dependence in it?
                     biotic_interactions=0, # any biotic interactions?
                      lambda.sim=0, # lambda analytically calculated
                      study.length=26,
                     l_ratio=c(df$l_ratio.R,df$l_ratio.R.cov,
                            df$l_ratio.T,df$l_ratio.T.cov,
                            df$l_ratio.D,df$l_ratio.D.cov))
                     

write.csv(new_df, "Sens_MouseLemurs.csv",row.names = F)

sens_no_cov=dplyr::filter(new_df,cov==0)
hist(sens_no_cov$sens)

sens_cov=dplyr::filter(new_df,cov==1)
hist(sens_cov$sens)
hist(sens_no_cov$sens-sens_cov$sens)

```



## 5) Vital-rate specific sensitivities
#### 5.1 Rain
```{r}
# empty vectors for lambdas
l.max.smj=NULL
l.min.smj=NULL

l.max.sfj=NULL
l.min.sfj=NULL

l.max.sma=NULL
l.min.sma=NULL

l.max.sfa=NULL
l.min.sfa=NULL

# empty vectors for sensitivities
deltaR.smj=NULL
deltaR.sfj=NULL
deltaR.sma=NULL
deltaR.sfa=NULL

for(i in 1:1000){
  
  # perturb male juvenile survival
  smj.max=matrix(c(0,Smj(dens_when_rain_max,max.rain,temp_when_rain_max,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    smj.min=matrix(c(0,Smj(dens_when_rain_min,min.rain,temp_when_rain_min,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
    
    l.max.smj[i]=lambda(smj.max)
    l.min.smj[i]=lambda(smj.min)

# perturb female juvenile survival
  sfj.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(mean.dens,mean.temp,i),Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
  sfj.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(mean.dens,mean.temp,i),Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)

  l.max.sfj[i]=lambda(sfj.max)
  l.min.sfj[i]=lambda(sfj.min)
  
  
deltaR.sfj[i]=abs((lambda(sfj.max)-lambda(sfj.min))/((max.rain-min.rain)/sd.rain))


# perturb male adult survival
  sma.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(dens_when_rain_max,max.rain,temp_when_rain_max,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    sma.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(dens_when_rain_min,min.rain,temp_when_rain_min,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    l.max.sma[i]=lambda(sma.max)
    l.min.sma[i]=lambda(sma.min)
    
    
deltaR.sma[i]=abs((lambda(sma.max)-lambda(sma.min))/((max.rain-min.rain)/sd.rain))


# perturb female adult survival
  sfa.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(mean.dens,mean.temp,i),Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)),n.stage,n.stage)
  
    sfa.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(mean.dens,mean.temp,i),Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)),n.stage,n.stage)
    
    l.max.sfa[i]=lambda(sfa.max)
    l.min.sfa[i]=lambda(sfa.min)
    
deltaR.sfa[i]=abs((lambda(sfa.max)-lambda(sfa.min))/((max.rain-min.rain)/sd.rain))


# perturb recruitment
# no rain in recruitment

}

# calculate sensitivities
# but first need to filter out the super large lambdas

# filter:
filter.maxmin.smj=unique(c(which(l.max.smj>3),which(l.min.smj>3)))  
l.max.smj=l.max.smj[-filter.maxmin.smj]
l.min.smj=l.min.smj[-filter.maxmin.smj]
# sens:
deltaR.smj=abs((l.max.smj-l.min.smj)/((max.rain-min.rain)/sd.rain))


# filter:
filter.maxmin.sfj=unique(c(which(l.max.sfj>3),which(l.min.sfj>3)))  
l.max.sfj=l.max.sfj[-filter.maxmin.sfj]
l.min.sfj=l.min.sfj[-filter.maxmin.sfj]
# sens:
deltaR.sfj=abs((l.max.sfj-l.min.sfj)/((max.rain-min.rain)/sd.rain))


# filter:
filter.maxmin.sma=unique(c(which(l.max.sma>3),which(l.min.sma>3)))  
l.max.sma=l.max.sma[-filter.maxmin.sma]
l.min.sma=l.min.sma[-filter.maxmin.sma]
# sens:
deltaR.sma=abs((l.max.sma-l.min.sma)/((max.rain-min.rain)/sd.rain))


# filter:
filter.maxmin.sfa=unique(c(which(l.max.sfa>3),which(l.min.sfa>3)))  
l.max.sfa=l.max.sfa[-filter.maxmin.sfa]
l.min.sfa=l.min.sfa[-filter.maxmin.sfa]
# sens:
deltaR.sfa=abs((l.max.sfa-l.min.sfa)/((max.rain-min.rain)/sd.rain))



# ALTERNATIVE grouping of vital rates and stages
# what if we already grouped them??

# perturb juvenile survival (both female and male)
deltaR.sj=NULL
for(i in 1:1000){
  sj.max=matrix(c(0,Smj(dens_when_rain_max,max.rain,temp_when_rain_max,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    sj.min=matrix(c(0,Smj(dens_when_rain_min,min.rain,temp_when_rain_min,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)

deltaR.sj[i]=abs((lambda(sj.max)-lambda(sj.min))/((max.rain-min.rain)/sd.rain))
}
hist(deltaR.sj)

# perturb reproduction (instead of just recruitment rate) of both juvenile and adult
deltaR.r=NULL
for(i in 1:1000){
  r.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),0,0.5*Sfj(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 
                 0.5*Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),0,0.5*Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)*rec(dens_when_rain_max,temp_when_rain_max,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    r.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),0,0.5*Sfj(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 
                 0.5*Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),0,0.5*Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)*rec(dens_when_rain_min,temp_when_rain_min,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
deltaR.r[i]=abs((lambda(r.max)-lambda(r.min))/((max.rain-min.rain)/sd.rain))
}
hist(deltaR.r)

# perturb adult survival (both female and male)
deltaR.sa=NULL
for(i in 1:1000){
  sa.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(dens_when_rain_max,max.rain,temp_when_rain_max,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(dens_when_rain_max,max.rain,temp_when_rain_max,i)),n.stage,n.stage)
  
    sa.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(dens_when_rain_min,min.rain,temp_when_rain_min,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(dens_when_rain_min,min.rain,temp_when_rain_min,i)),n.stage,n.stage)
  

deltaR.sa[i]=abs((lambda(sa.max)-lambda(sa.min))/((max.rain-min.rain)/sd.rain))
}
hist(deltaR.sa)


```


#### 5.2 Temperature
```{r}
# empty vectors for lambdas
l.max.smj=NULL
l.min.smj=NULL

l.max.sfj=NULL
l.min.sfj=NULL

l.max.sma=NULL
l.min.sma=NULL

l.max.sfa=NULL
l.min.sfa=NULL

l.max.rec=NULL
l.min.rec=NULL


# empty vectors for sensitivities
deltaT.smj=NULL
deltaT.sfj=NULL
deltaT.sma=NULL
deltaT.sfa=NULL
deltaT.rec=NULL



for(i in 1:1000){
  
  # perturb male juvenile survival
  smj.max=matrix(c(0,Smj(dens_when_temp_max,rain_when_temp_max,max.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    smj.min=matrix(c(0,Smj(dens_when_temp_min,rain_when_temp_min,min.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)

    l.max.smj[i]=lambda(smj.max)
    l.min.smj[i]=lambda(smj.min)
    


# perturb female juvenile survival
  sfj.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(mean.dens,mean.temp,i),Sfj(dens_when_temp_max,rain_when_temp_max,max.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
  sfj.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(mean.dens,mean.temp,i),Sfj(dens_when_temp_min,rain_when_temp_min,min.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    l.max.sfj[i]=lambda(sfj.max)
    l.min.sfj[i]=lambda(sfj.min)



# perturb male adult survival
  sma.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(dens_when_temp_max,rain_when_temp_max,max.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    sma.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(dens_when_temp_min,rain_when_temp_min,min.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  

    
    l.max.sma[i]=lambda(sma.max)
    l.min.sma[i]=lambda(sma.min)

# perturb female adult survival
  sfa.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(dens_when_temp_max,rain_when_temp_max,max.temp,i)*rec(mean.dens,mean.temp,i),Sfa(dens_when_temp_max,rain_when_temp_max,max.temp,i)),n.stage,n.stage)
  
    sfa.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(dens_when_temp_min,rain_when_temp_min,min.temp,i)*rec(mean.dens,mean.temp,i),Sfa(dens_when_temp_min,rain_when_temp_min,min.temp,i)),n.stage,n.stage)
    
    l.max.sfa[i]=lambda(sfa.max)
    l.min.sfa[i]=lambda(sfa.min)
    
    
  # perturb recruitment
  rec.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_max,max.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_max,max.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_max,max.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_max,max.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    rec.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_min,min.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_min,min.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_min,min.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(dens_when_temp_min,min.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  

    l.max.rec[i]=lambda(rec.max)
    l.min.rec[i]=lambda(rec.min)
    
}
# I feel like we have a problem with the recruitment vital rate, the coefficient is way to large (-10.97) and I think this makes large lambdas and thus sensitivities


# calculate sensitivities
# but first need to filter out the super large lambdas

# filter:
filter.maxmin.smj=unique(c(which(l.max.smj>3),which(l.min.smj>3)))  
l.max.smj=l.max.smj[-filter.maxmin.smj]
l.min.smj=l.min.smj[-filter.maxmin.smj]
# sens:
deltaT.smj=abs((l.max.smj-l.min.smj)/((max.temp-min.temp)/sd.temp))


# filter:
filter.maxmin.sfj=unique(c(which(l.max.sfj>3),which(l.min.sfj>3)))  
l.max.sfj=l.max.sfj[-filter.maxmin.sfj]
l.min.sfj=l.min.sfj[-filter.maxmin.sfj]
# sens:
deltaT.sfj=abs((l.max.sfj-l.min.sfj)/((max.temp-min.temp)/sd.temp))


# filter:
filter.maxmin.sma=unique(c(which(l.max.sma>3),which(l.min.sma>3)))  
l.max.sma=l.max.sma[-filter.maxmin.sma]
l.min.sma=l.min.sma[-filter.maxmin.sma]
# sens:
deltaT.sma=abs((l.max.sma-l.min.sma)/((max.temp-min.temp)/sd.temp))


# filter:
filter.maxmin.sfa=unique(c(which(l.max.sfa>3),which(l.min.sfa>3)))  
l.max.sfa=l.max.sfa[-filter.maxmin.sfa]
l.min.sfa=l.min.sfa[-filter.maxmin.sfa]
# sens:
deltaT.sfa=abs((l.max.sfa-l.min.sfa)/((max.temp-min.temp)/sd.temp))


# filter:
filter.maxmin.rec=unique(c(which(l.max.rec>3),which(l.min.rec>3)))
l.max.rec=l.max.rec[-filter.maxmin.rec]
l.min.rec=l.min.rec[-filter.maxmin.rec]
# sens:
deltaT.rec=abs((l.max.rec-l.min.rec)/((max.temp-min.temp)/sd.temp))


```


#### 5.3 Density
```{r}
# empty vectors for lambdas
l.max.smj=NULL
l.min.smj=NULL

l.max.sfj=NULL
l.min.sfj=NULL

l.max.sma=NULL
l.min.sma=NULL

l.max.sfa=NULL
l.min.sfa=NULL

l.max.rec=NULL
l.min.rec=NULL


# empty vectors for sensitivities
deltaD.smj=NULL
deltaD.sfj=NULL
deltaD.sma=NULL
deltaD.sfa=NULL
deltaD.rec=NULL


for(i in 1:1000){
  # perturb male juvenile survival
  smj.max=matrix(c(0,Smj(max.dens,rain_when_dens_max,temp_when_dens_max,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    smj.min=matrix(c(0,Smj(min.dens,rain_when_dens_min,temp_when_dens_min,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
    
    l.max.smj[i]=lambda(smj.max)
    l.min.smj[i]=lambda(smj.min)

    
deltaD.smj[i]=abs((lambda(smj.max)-lambda(smj.min))/((max.dens-min.dens)/sd.dens))


# perturb female juvenile survival
  sfj.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(mean.dens,mean.temp,i),Sfj(max.dens,rain_when_dens_max,temp_when_dens_max,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
  sfj.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(mean.dens,mean.temp,i),Sfj(min.dens,rain_when_dens_min,temp_when_dens_min,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    l.max.sfj[i]=lambda(sfj.max)
    l.min.sfj[i]=lambda(sfj.min)

deltaD.sfj[i]=abs((lambda(sfj.max)-lambda(sfj.min))/((max.dens-min.dens)/sd.dens))


# perturb male adult survival
  sma.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(max.dens,rain_when_dens_max,temp_when_dens_max,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    sma.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(min.dens,rain_when_dens_min,temp_when_dens_min,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
    
    l.max.sma[i]=lambda(sma.max)
    l.min.sma[i]=lambda(sma.min)
  
deltaD.sma[i]=abs((lambda(sma.max)-lambda(sma.min))/((max.dens-min.dens)/sd.dens))


# perturb female adult survival
  sfa.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(max.dens,rain_when_dens_max,temp_when_dens_max,i)*rec(mean.dens,mean.temp,i),Sfa(max.dens,rain_when_dens_max,temp_when_dens_max,i)),n.stage,n.stage)
  
    sfa.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(mean.dens,mean.temp,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(mean.dens,mean.temp,i),0,0.5*Sfa(min.dens,rain_when_dens_min,temp_when_dens_min,i)*rec(mean.dens,mean.temp,i),Sfa(min.dens,rain_when_dens_min,temp_when_dens_min,i)),n.stage,n.stage)
    
    l.max.sfa[i]=lambda(sfa.max)
    l.min.sfa[i]=lambda(sfa.min)
    
deltaD.sfa[i]=abs((lambda(sfa.max)-lambda(sfa.min))/((max.dens-min.dens)/sd.dens))


  # perturb recruitment
  rec.max=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(max.dens,temp_when_dens_max,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(max.dens,temp_when_dens_max,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(max.dens,temp_when_dens_max,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(max.dens,temp_when_dens_max,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    rec.min=matrix(c(0,Smj(mean.dens,mean.rain,mean.temp,i),0,0,
                 0,Sma(mean.dens,mean.rain,mean.temp,i),0,0,
                 0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(min.dens,temp_when_dens_min,i),0,0.5*Sfj(mean.dens,mean.rain,mean.temp,i)*rec(min.dens,temp_when_dens_min,i),Sfj(mean.dens,mean.rain,mean.temp,i),
                 0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(min.dens,temp_when_dens_min,i),0,0.5*Sfa(mean.dens,mean.rain,mean.temp,i)*rec(min.dens,temp_when_dens_min,i),Sfa(mean.dens,mean.rain,mean.temp,i)),n.stage,n.stage)
  
    l.max.rec[i]=lambda(rec.max)
    l.min.rec[i]=lambda(rec.min)

deltaD.rec[i]=abs((lambda(rec.max)-lambda(rec.min))/((max.dens-min.dens)/sd.dens))

}


# calculate sensitivities
# but first need to filter out the super large lambdas

# filter:
filter.maxmin.smj=unique(c(which(l.max.smj>3),which(l.min.smj>3)))  
l.max.smj=l.max.smj[-filter.maxmin.smj]
l.min.smj=l.min.smj[-filter.maxmin.smj]
# sens:
deltaD.smj=abs((l.max.smj-l.min.smj)/((max.dens-min.dens)/sd.dens))


# filter:
filter.maxmin.sfj=unique(c(which(l.max.sfj>3),which(l.min.sfj>3)))  
l.max.sfj=l.max.sfj[-filter.maxmin.sfj]
l.min.sfj=l.min.sfj[-filter.maxmin.sfj]
# sens:
deltaD.sfj=abs((l.max.sfj-l.min.sfj)/((max.dens-min.dens)/sd.dens))


# filter:
filter.maxmin.sma=unique(c(which(l.max.sma>3),which(l.min.sma>3)))  
l.max.sma=l.max.sma[-filter.maxmin.sma]
l.min.sma=l.min.sma[-filter.maxmin.sma]
# sens:
deltaD.sma=abs((l.max.sma-l.min.sma)/((max.dens-min.dens)/sd.dens))


# filter:
filter.maxmin.sfa=unique(c(which(l.max.sfa>3),which(l.min.sfa>3)))  
l.max.sfa=l.max.sfa[-filter.maxmin.sfa]
l.min.sfa=l.min.sfa[-filter.maxmin.sfa]
# sens:
deltaD.sfa=abs((l.max.sfa-l.min.sfa)/((max.dens-min.dens)/sd.dens))


# filter:
filter.maxmin.rec=unique(c(which(l.max.rec>3),which(l.min.rec>3)))
l.max.rec=l.max.rec[-filter.maxmin.rec]
l.min.rec=l.min.rec[-filter.maxmin.rec]
# sens:
deltaD.rec=abs((l.max.rec-l.min.rec)/((max.dens-min.dens)/sd.dens))


```

#### 5.4 Save Output
```{r}
surv_vr=data.frame(study.doi="10.1073/pnas.2214244120",
                     year.of.publication="2023",
                     group="Mammals",
                     species="Microcebus murinus",
                     continent="Africa",
                     driver=rep(c("rain","temperature","density"),each=400),
                     driver.type=rep(c("C","C","D"),each=400),
                     stage.age=rep(c("male juvenile","female juvenile","male adult","female adult"),each=100),
                     vital.rates="survival",
                     sens=c(deltaR.smj[1:100],deltaR.sfj[1:100],deltaR.sma[1:100],deltaR.sfa[1:100],
                            deltaT.smj[1:100],deltaT.sfj[1:100],deltaT.sma[1:100],deltaT.sfa[1:100],
                            deltaD.smj[1:100],deltaD.sfj[1:100],deltaD.sma[1:100],deltaD.sfa[1:100]),
                     mat=1, # age at sexual maturity Ozgul et al. 2023
                     n.vr=5, # number of vital rates with covariates
                     n.pam=27, # number of parameters
                     dens=1, # density dependence in it?
                     biotic_interactions=0, # any biotic interactions?
                      lambda.sim=0, # lambda analytically calculated
                      study.length=26)

rec_vr=data.frame(study.doi="10.1073/pnas.2214244120",
                     year.of.publication="2023",
                     group="Mammals",
                     species="Microcebus murinus",
                     continent="Africa",
                     driver=rep(c("temperature","density"),100),
                     driver.type=rep(c("C","D"),100),
                     stage.age="reproductive",
                     vital.rates="recruitment",
                     sens=c(deltaT.rec[1:100],
                            deltaD.rec[1:100]),
                     mat=1, # age at sexual maturity Ozgul et al. 2023
                     n.vr=5, # number of vital rates with covariates
                     n.pam=27, # number of parameters
                     dens=1, # density dependence in it?
                     biotic_interactions=0, # any biotic interactions?
                      lambda.sim=0, # lambda analytically calculated
                      study.length=26)

results_vr=rbind(surv_vr,rec_vr)

write.csv(results_vr,"SensVR_MouseLemurs.csv",row.names = F)
```



```{r}
rm(df)
df=read.csv("/Users/esinickin/Documents/Master Thesis/pert_analyses/mouse lemurs/SensVR_MouseLemurs.csv")
names(df)[names(df) == "gen.time"] <- "mat"
df$mat=1
df$lambda.sim=0
df$study.length=26
write.csv(df,"/Users/esinickin/Documents/Master Thesis/pert_analyses/mouse lemurs/SensVR_MouseLemurs.csv",row.names = F)
```

