---
title: "Malchow et al. 2023"
author: "Esin Ickin"
date: "2023-06-02"
output: html_document
---

This is the code for sensitivity analyses conducted on the Alpine accentor (*Prunella collaris*) which is species 10 of this study: https://doi.org/10.1098/rstb.2022.0194


## 0) Prepare session
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Documents/Master Thesis/pert_analyses/birds")
```

```{r}
rm(list=ls()) # reset

# packages
library(popbio)
library(tidyverse)
```

## 1) Vital rates

This part is a modified version of 2.) from the "analysis_main.R" where you compose demography-environment relationships from calibrated parameters for each species

```{r}
## a) Fecundity
loc_dens <- 1
DER_Fecund <- function(Fec0, Fc.pr1, Fc.pr2, Fc.tm1, Fc.tm2, DDp, Tbr=0, Pbr=0, Hab=100){ 
	exp(
		log(Fec0) + 
		Fc.pr1*Pbr + 
		Fc.pr2*Pbr^2 + 
		Fc.tm1*Tbr + 
		Fc.tm2*Tbr^2 ) *
	# density-dependence
	exp( - 1 * eval(loc_dens) / DDp / Hab )
}


## b) Juvenile survival
DER_juvSurv <- function(jSv0, jS.pr1, jS.pr2, jS.tm1, jS.tm2, Tat=0, Pwn=0){
	plogis(
		jS.pr1*Pwn + 
		jS.pr2*Pwn^2 + 
		jS.tm1*Tat + 
		jS.tm2*Tat^2 ,
		location = log((1-jSv0)/jSv0))
}


## c) Adult survival
DER_adSurv <- function(aSv0, aS.pr1, aS.pr2, aS.tm1, Twn=0, Pwn=0){
	plogis(
		aS.pr1*Pwn + 
		aS.pr2*Pwn^2 + 
		aS.tm1*Twn ,
		location = log((1-aSv0)/aSv0))
}

# SPECIES 10 parameters
post_sample_list=readRDS("Post_sample_list.rds")
# Juvenile survival
jSv0.sp10=post_sample_list$`Alpine accentor`$juvSurv
jS.pr1.sp10=post_sample_list$`Alpine accentor`$jSb.pr1
jS.pr2.sp10=post_sample_list$`Alpine accentor`$jSb.pr2
jS.tm1.sp10=post_sample_list$`Alpine accentor`$jSb.tm1
jS.tm2.sp10=post_sample_list$`Alpine accentor`$jSb.tm2

# Fecundity
Fec0.sp10=post_sample_list$`Alpine accentor`$Fecund
Fc.pr1.sp10=post_sample_list$`Alpine accentor`$Fcb.pr1
Fc.pr2.sp10=post_sample_list$`Alpine accentor`$Fcb.pr2
Fc.tm1.sp10=post_sample_list$`Alpine accentor`$Fcb.tm1
Fc.tm2.sp10=post_sample_list$`Alpine accentor`$Fcb.tm2
DDp.sp10=post_sample_list$`Alpine accentor`$DensDep

# Adult survival
aSv0.sp10=post_sample_list$`Alpine accentor`$adSurv
aS.pr1.sp10=post_sample_list$`Alpine accentor`$aSb.pr1
aS.pr2.sp10=post_sample_list$`Alpine accentor`$aSb.pr2
aS.tm1.sp10=post_sample_list$`Alpine accentor`$aSb.tm
```

## 2) Covariates

species-specific covariates

```{r}
cov=read.csv("PrunellaCollarisCovariates.csv")

# Tbr = mean temp during breeding season (april-july) for fecundity
max.Tbr=max(cov$breeding_tmp)
min.Tbr=min(cov$breeding_tmp)
mean.Tbr=0
sd.Tbr=1
# Covariation
Pbr_when_Tbr_max=cov$breeding_pr[which(cov$breeding_tmp==max(cov$breeding_tmp))][1]
Pbr_when_Tbr_min=cov$breeding_pr[which(cov$breeding_tmp==min(cov$breeding_tmp))][1]
Tat_when_Tbr_max=cov$postbreed_tmp[which(cov$breeding_tmp==max(cov$breeding_tmp))][1]
Tat_when_Tbr_min=cov$postbreed_tmp[which(cov$breeding_tmp==min(cov$breeding_tmp))][1]
Twn_when_Tbr_max=cov$winter_tmin[which(cov$breeding_tmp==max(cov$breeding_tmp))][1]
Twn_when_Tbr_min=cov$winter_tmin[which(cov$breeding_tmp==min(cov$breeding_tmp))][1]
Pwn_when_Tbr_max=cov$winter_pr[which(cov$breeding_tmp==max(cov$breeding_tmp))][1]
Pwn_when_Tbr_min=cov$winter_pr[which(cov$breeding_tmp==min(cov$breeding_tmp))][1]

# Pwn = tot precip during winter (dec-feb) for juvenile survival, adult survival
max.Pwn=max(cov$winter_pr)
min.Pwn=min(cov$winter_pr)
mean.Pwn=0
sd.Pwn=1
# Covariation
Pbr_when_Pwn_max=cov$breeding_pr[which(cov$winter_pr==max(cov$winter_pr))][1]
Pbr_when_Pwn_min=cov$breeding_pr[which(cov$winter_pr==min(cov$winter_pr))][1]
Tat_when_Pwn_max=cov$postbreed_tmp[which(cov$winter_pr==max(cov$winter_pr))][1]
Tat_when_Pwn_min=cov$postbreed_tmp[which(cov$winter_pr==min(cov$winter_pr))][1]
Twn_when_Pwn_max=cov$winter_tmin[which(cov$winter_pr==max(cov$winter_pr))][1]
Twn_when_Pwn_min=cov$winter_tmin[which(cov$winter_pr==min(cov$winter_pr))][1]
Tbr_when_Pwn_max=cov$breeding_tmp[which(cov$winter_pr==max(cov$winter_pr))][1]
Tbr_when_Pwn_min=cov$breeding_tmp[which(cov$winter_pr==min(cov$winter_pr))][1]

# Pbr = tot precip during breeding season (april-july) for fecundity
max.Pbr=max(cov$breeding_pr)
min.Pbr=min(cov$breeding_pr)
mean.Pbr=0
sd.Pbr=1
# Covariation
Pwn_when_Pbr_max=cov$winter_pr[which(cov$breeding_pr==max(cov$breeding_pr))][1]
Pwn_when_Pbr_min=cov$winter_pr[which(cov$breeding_pr==min(cov$breeding_pr))][1]
Tat_when_Pbr_max=cov$postbreed_tmp[which(cov$breeding_pr==max(cov$breeding_pr))][1]
Tat_when_Pbr_min=cov$postbreed_tmp[which(cov$breeding_pr==min(cov$breeding_pr))][1]
Twn_when_Pbr_max=cov$winter_tmin[which(cov$breeding_pr==max(cov$breeding_pr))][1]
Twn_when_Pbr_min=cov$winter_tmin[which(cov$breeding_pr==min(cov$breeding_pr))][1]
Tbr_when_Pbr_max=cov$breeding_tmp[which(cov$breeding_pr==max(cov$breeding_pr))][1]
Tbr_when_Pbr_min=cov$breeding_tmp[which(cov$breeding_pr==min(cov$breeding_pr))][1]

# Tat = mean temp during autumn (sep-nov) for juvenile survival
max.Tat=max(cov$postbreed_tmp)
min.Tat=min(cov$postbreed_tmp)
mean.Tat=0
sd.Tat=1
# Covariation
Pwn_when_Tat_max=cov$winter_pr[which(cov$postbreed_tmp==max(cov$postbreed_tmp))][1]
Pwn_when_Tat_min=cov$winter_pr[which(cov$postbreed_tmp==min(cov$postbreed_tmp))][1]
Pbr_when_Tat_max=cov$breeding_pr[which(cov$postbreed_tmp==max(cov$postbreed_tmp))][1]
Pbr_when_Tat_min=cov$breeding_pr[which(cov$postbreed_tmp==min(cov$postbreed_tmp))][1]
Twn_when_Tat_max=cov$winter_tmin[which(cov$postbreed_tmp==max(cov$postbreed_tmp))][1]
Twn_when_Tat_min=cov$winter_tmin[which(cov$postbreed_tmp==min(cov$postbreed_tmp))][1]
Tbr_when_Tat_max=cov$breeding_tmp[which(cov$postbreed_tmp==max(cov$postbreed_tmp))][1]
Tbr_when_Tat_min=cov$breeding_tmp[which(cov$postbreed_tmp==min(cov$postbreed_tmp))][1]

# Twn = min temp during winter (dec-feb) for (juvenile survival), adult survival
max.Twn=max(cov$winter_tmin)
min.Twn=min(cov$winter_tmin)
mean.Twn=0
sd.Twn=1
# Covariation
Pwn_when_Twn_max=cov$winter_pr[which(cov$winter_tmin==max(cov$winter_tmin))][1]
Pwn_when_Twn_min=cov$winter_pr[which(cov$winter_tmin==min(cov$winter_tmin))][1]
Pbr_when_Twn_max=cov$breeding_pr[which(cov$winter_tmin==max(cov$winter_tmin))][1]
Pbr_when_Twn_min=cov$breeding_pr[which(cov$winter_tmin==min(cov$winter_tmin))][1]
Tat_when_Twn_max=cov$postbreed_tmp[which(cov$winter_tmin==max(cov$winter_tmin))][1]
Tat_when_Twn_min=cov$postbreed_tmp[which(cov$winter_tmin==min(cov$winter_tmin))][1]
Tbr_when_Twn_max=cov$breeding_tmp[which(cov$winter_tmin==max(cov$winter_tmin))][1]
Tbr_when_Twn_min=cov$breeding_tmp[which(cov$winter_tmin==min(cov$winter_tmin))][1]
```

## 3) Population model

Female-only, 2-staged matrix model with mean covariates

```{r}
# SPECIES 10
# mpm <- matrix(c(0,Sj,
#                fec,Sa),n.stage,n.stage)
lambda_mean_popmodel=NULL
for(u in 1:100){
  mpm <- matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)
  lambda_mean_popmodel[u]=lambda(mpm)
}

hist(lambda_mean_popmodel)
mean(lambda_mean_popmodel)
```

## 4) Scaled Sensitivity analyses

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: <https://doi.org/10.1073/pnas.1918363117>)

Note that this is a step that Esin will implement in her MS thesis. With the information given in 1-3, we should be able to run these analyses. 
### 4.1 Tbr

```{r}
# empty vectors for sensitivities
delta.Tbr=NULL
deltaTbr.cov=NULL

# empty vectors for lambdas for Tbr & cov
l.max.Tbr2=NULL
l.min.Tbr2=NULL

# empty vectors for log ratios of the lambdas:
l_ratio.Tbr=NULL
l_ratio.Tbr.cov=NULL

# Tbr (breeding temperature)
for(u in 1:1000){
  # 1. Sensitivity to Tbr assuming mean covariates
# matrix with max Tbr values
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=max.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)
# matrix with min Tbr values
mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=min.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

# sensitivity to Tbr without covariation
l.Tbr.max=lambda(mpm.max)
l.Tbr.min=lambda(mpm.min)

delta.Tbr[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Tbr-min.Tbr)/sd.Tbr))

l_ratio.Tbr[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))

# 2. Sensitivity to Tbr assuming covariation
# matrix with max values
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Tbr_max,Pwn=Pwn_when_Tbr_max),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=max.Tbr,Pbr=Pbr_when_Tbr_max),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Tbr_max,Pwn=Pwn_when_Tbr_max)),2,2)
# matrix with min values
mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Tbr_min,Pwn=Pwn_when_Tbr_min),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=min.Tbr,Pbr=Pbr_when_Tbr_min),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Tbr_min,Pwn=Pwn_when_Tbr_min)),2,2)

# lambda Tbr cov 
l.max.Tbr2[u]=lambda(mpm.max)
l.min.Tbr2[u]=lambda(mpm.min)

}

# after checking the sensitivities, it's wayy too large for perturbations of Tbr with covariation
# therefore, we will filter the lambdas for this one

# filter:
filter.maxmin.Tbr2=unique(c(which(l.max.Tbr2>3),which(l.min.Tbr2>3)))  
l.max.Tbr2=l.max.Tbr2[-filter.maxmin.Tbr2]
l.min.Tbr2=l.min.Tbr2[-filter.maxmin.Tbr2]

# calculate sensitivities:
deltaTbr.cov=abs((l.max.Tbr2-l.min.Tbr2)/((max.Tbr-min.Tbr)/sd.Tbr))
l_ratio.Tbr.cov=abs(log(l.max.Tbr2/l.min.Tbr2))

# now get the first 1:100 so that the length matches with the other ones
deltaTbr.cov=deltaTbr.cov[1:100]
# same with delta.Tbr
delta.Tbr=delta.Tbr[1:100]

# same with log ratios
l_ratio.Tbr=l_ratio.Tbr[1:100]
l_ratio.Tbr.cov=l_ratio.Tbr.cov[1:100]



# check sens that were weird:
hist(deltaTbr.cov) # whyyyy is it sill sooo big?

hist(l.max.Tbr2) # these lambdas are really large
hist(l.min.Tbr2) # these are fine

# can't find an error but I also don't want to filter way too much or differently than I did with the mouse lemurs because of standardization etc.
# so I'll leave it for now...
# or we will just omit these sensitivities to Tbr with and without covariation
```

### 4.2 Pwn

```{r}
delta.Pwn=NULL
deltaPwn.cov=NULL

l_ratio.Pwn=NULL
l_ratio.Pwn.cov=NULL

# Pwn (winter precipitation)
for(u in 1:100){
  # 1. Sensitivity to Pwn assuming mean covariates
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=max.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=max.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=min.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=min.Pwn)),2,2)

delta.Pwn[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pwn-min.Pwn)/sd.Pwn))

l_ratio.Pwn[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))

# 2. Sensitivity to Pwn assuming covariation
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Pwn_max,Pwn=max.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Pwn_max,Pbr=Pbr_when_Pwn_max), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Pwn_max,Pwn=max.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Pwn_min,Pwn=min.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Pwn_min,Pbr=Pbr_when_Pwn_min), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Pwn_min,Pwn=min.Pwn)),2,2)

deltaPwn.cov[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pwn-min.Pwn)/sd.Pwn))

l_ratio.Pwn.cov[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))

}

hist(delta.Pwn)
hist(deltaPwn.cov)
```

### 4.3 Pbr

```{r}
# Pbr = breeding precip
delta.Pbr=NULL
deltaPbr.cov=NULL

l_ratio.Pbr=NULL
l_ratio.Pbr.cov=NULL

for(u in 1:100){
  # 1. Sensitivity to Pbr assuming mean covariates
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=max.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=min.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

# sensitivity to Pbr without covariation
delta.Pbr[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pbr-min.Pbr)/sd.Pbr))

l_ratio.Pbr[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))


# 2. Sensitivity to Pbr assuming covariation
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Pbr_max,Pwn=Pwn_when_Pbr_max),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Pbr_max,Pbr=max.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Pbr_max,Pwn=Pwn_when_Pbr_max)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Pbr_min,Pwn=Pwn_when_Pbr_min),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Pbr_min,Pbr=min.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Pbr_min,Pwn=Pwn_when_Pbr_min)),2,2)

# sensitivity to Pbr with covariation
deltaPbr.cov[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pbr-min.Pbr)/sd.Pbr))

l_ratio.Pbr.cov[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))

}
hist(delta.Pbr)
hist(deltaPbr.cov)
```

### 4.4 Twn

```{r}
# Twn (winter temp)
delta.Twn=NULL
deltaTwn.cov=NULL

l_ratio.Twn=NULL
l_ratio.Twn.cov=NULL

for(u in 1:100){
  # 1. Sensitivity to Twn assuming mean covariates
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=max.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=min.Twn,Pwn=mean.Pwn)),2,2)

delta.Twn[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Twn-min.Twn)/sd.Twn))

l_ratio.Twn[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))

# 2. Sensitivity to Twn assuming covariation
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Twn_max,Pwn=Pwn_when_Twn_max),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Twn_max,Pbr=Pbr_when_Twn_max), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=max.Twn,Pwn=Pwn_when_Twn_max)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Twn_min,Pwn=Pwn_when_Twn_min),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Twn_min,Pbr=Pbr_when_Twn_min), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=min.Twn,Pwn=Pwn_when_Twn_min)),2,2)

deltaTwn.cov[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Twn-min.Twn)/sd.Twn))

l_ratio.Twn.cov[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))
}

hist(delta.Twn)
hist(deltaTwn.cov)
```

### 4.5 Tat

```{r}
# Tat (postbreeding temp)
delta.Tat=NULL
deltaTat.cov=NULL

l_ratio.Tat=NULL
l_ratio.Tat.cov=NULL

for(u in 1:100){
  # 1. Sensitivity to Tat assuming mean covariates
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=max.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=min.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

delta.Tat[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Tat-min.Tat)/sd.Tat))

l_ratio.Tat[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))

# 2. Sensitivity to Twn assuming covariation
mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=max.Tat,Pwn=Pwn_when_Tat_max),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Tat_max,Pbr=Pbr_when_Tat_max), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Tat_max,Pwn=Pwn_when_Tat_max)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=min.Tat,Pwn=Pwn_when_Tat_min),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Tat_min,Pbr=Pbr_when_Tat_min), 
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Tat_min,Pwn=Pwn_when_Tat_min)),2,2)

deltaTat.cov[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Tat-min.Tat)/sd.Tat))

l_ratio.Tat.cov[u]=abs(log(lambda(mpm.max)/lambda(mpm.min)))
}

hist(delta.Tat)
hist(deltaTat.cov)
```

# SAVE OUTPUT
```{r}

# no cov vs with cov
ScaledSens=data.frame(study.doi="10.1098/rstb.2022.0194",
                      year.of.publication="2023",
                      group="Birds",
                      species="Prunella collaris",
                      continent="Europe",
                      driver=rep(c("Tbr","Tat","Twn","Pbr","Pwn"),each=200),
                      driver.type="C",
                      stage.age="all",
                      vital.rates="all",
                      sens=c(delta.Tbr,deltaTbr.cov,
                      	   delta.Tat,deltaTat.cov,
                      	   delta.Twn,deltaTwn.cov,
                      	   delta.Pbr,deltaPbr.cov,
                      	   delta.Pwn,deltaPwn.cov),
					  cov=rep(c(0,1),100),
					  mat=1, # age at first repro
                	  n.vr=3, # number of vital rates with covariates
                      n.pam=15, # number of parameters
                	  dens=0, # density dependence in it?
                	  biotic_interactions=0, # any biotic interactions?
					  lambda.sim=0,
					  study.length=20,
					  l_ratio=c(delta.Tbr,deltaTbr.cov,
                      	   delta.Tat,deltaTat.cov,
                      	   delta.Twn,deltaTwn.cov,
                      	   delta.Pbr,deltaPbr.cov,
                      	   delta.Pwn,deltaPwn.cov))

write.csv(ScaledSens,"output/PrunellaCollaris_Sens.csv",row.names = F)

```



## 5) Scaled sensitivity analyses per vr

where you perturb (considering covariation) each vital rate separately. Say, you have juvenile and adult survival; then you calculate scaled sensitivity perturbing juvenile survival (max, min) only, leaving the effect of the drivers at their mean for adult survival; and do the same for adult survival, leaving the effect of drivers at their mean for juvenile survival.

### 5.1) Tbr

```{r}
# 1. Sensitivity to Tbr (breeding temperature) assuming covariation
# FECUNDITY
deltaTbr.Fec=NULL
for(u in 1:100){
  mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=max.Tbr,Pbr=Pbr_when_Tbr_max),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=min.Tbr,Pbr=Pbr_when_Tbr_min),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

deltaTbr.Fec[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Tbr-min.Tbr)/sd.Tbr))
  
}

# JUVENILE SURVIVAL
# no Tbr in juvenile surv

# ADULT SURVIVAL
# no Tbr in adult surv
```

### 5.2) Pwn

```{r}
# 2. Sensitivity to Pwn (winter precip) assuming covariation
# JUVENILE SURVIVAL
deltaPwn.JSuv=NULL

for(u in 1:100){
  mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Pwn_max,Pwn=max.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=Tat_when_Pwn_min,Pwn=min.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

deltaPwn.JSuv[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pwn-min.Pwn)/sd.Pwn))
}


# ADULT SURVIVAL
deltaPwn.ASuv=NULL
for(u in 1:100){
  mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Pwn_max,Pwn=max.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=Twn_when_Pwn_min,Pwn=min.Pwn)),2,2)

deltaPwn.ASuv[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pwn-min.Pwn)/sd.Pwn))
}

```

### 5.3) Pbr

```{r}
# 3. Sensitivity to Pbr
# only in fecundity
deltaPbr.Fec=NULL
for(u in 1:100){
  mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=Tbr_when_Pbr_max,Pbr=max.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=min.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Twn)),2,2)

deltaPbr.Fec[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Pbr-min.Pbr)/sd.Pbr))
}
```

### 5.4) Twn

```{r}
# 4. Sensitivity to Twn
# only in adult survival
deltaTwn.ASurv=NULL
for(u in 1:100){
  mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=max.Twn,Pwn=Pwn_when_Twn_max)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=mean.Tat,Pwn=mean.Pwn),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=min.Twn,Pwn=Pwn_when_Twn_min)),2,2)

deltaTwn.ASurv[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Twn-min.Twn)/sd.Twn))
  
}

```

### 5.5) Tat

```{r}
# 3. Sensitivity to Tat (autumn temp)
# only in juvenile survival
deltaTat.JSurv=NULL
for(u in 1:100){
  mpm.max=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=max.Tat,Pwn=Pwn_when_Tat_max),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Pwn)),2,2)

mpm.min=matrix(c(0,DER_juvSurv(jSv0.sp10[u],jS.pr1.sp10[u],jS.pr2.sp10[u],jS.tm1.sp10[u],jS.tm2.sp10[u],Tat=min.Tat,Pwn=Pwn_when_Tat_min),
                DER_Fecund(Fec0.sp10[u],Fc.pr1.sp10[u],Fc.pr2.sp10[u],Fc.tm1.sp10[u],Fc.tm2.sp10[u],DDp.sp10[u],Tbr=mean.Tbr,Pbr=mean.Pbr),
                DER_adSurv(aSv0.sp10[u],aS.pr1.sp10[u],aS.pr2.sp10[u],aS.tm1.sp10[u],Twn=mean.Twn,Pwn=mean.Twn)),2,2)

deltaTat.JSurv[u]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.Tat-min.Tat)/sd.Tat))
}
```

## 6) Check sensitivities
```{r}
sens.no.cov=c(delta.Tbr,delta.Tat,delta.Twn,delta.Pbr,delta.Pwn)
                      sens.cov=c(deltaTbr.cov,deltaTat.cov,deltaTwn.cov,deltaPbr.cov,deltaPwn.cov)
hist(sens.cov)

filtered.sens.cov=which(sens.cov<=quantile(sens.cov[!is.nan(sens.cov)],0.9))
hist(filtered.sens.cov)
                           sens.per.vr=c(deltaTbr.Fec,deltaPbr.Fec,deltaPwn.JSuv,deltaPwn.ASuv,deltaTat.JSurv,deltaTwn.ASurv)                     
                      
plot(sens.no.cov)
plot(sens.cov) # way too big
plot(sens.per.vr) # also way too big

```


## 7) Save output

```{r}
# per vr
ScaledSensVR=data.frame(study.doi="10.1098/rstb.2022.0194",
                        year.of.publication="2023",
                        group="Birds",
                        species="Prunella collaris",
                        continent="Europe",
                        driver=rep(c("Tbr","Pbr","Pwn","Pwn","Tat","Twn"),each=100),
                        driver.type="C",
                        stage.age=rep(c("adult","adult","juvenile","adult","juvenile","adult"),each=100),
                        vital.rates=rep(c("fecundity","fecundity","survival","survival","survival","survival")),
                        sens=c(deltaTbr.Fec,deltaPbr.Fec,deltaPwn.JSuv,deltaPwn.ASuv,deltaTat.JSurv,deltaTwn.ASurv),
						mat=1, # age at first repro
                	  n.vr=3, # number of vital rates with covariates
                      n.pam=15, # number of parameters
                	  dens=0, # density dependence in it?
                	  biotic_interactions=0, # any biotic interactions?
					  lambda.sim=0,
					  study.length=20)

write.csv(sensvr,"output/PrunellaCollaris_Sens_per_VR.csv",row.names = F)

```

