---
title: "Sensitivity analyses"
author: "Esin Ickin, Chloé Nater"
date: "4/11/2023"
output: html_document
---

This code is an implementation of Esin's template for Nater et al.'s study on striped mice.
The data files needed for running the code below, as well as the original files from which the code below is taken, have been deposited in a folder "Files_StripedMouse" within the repository. Questions about details of the study and/or code in sections 1-3 below can be sent to chloe.nater@nina.no.

## 0) Prepare session
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Files_StripedMouse")
```


```{r}
#rm(list=ls())
library(lme4)
library(ggplot2)
library(popbio)
library(plyr)
library(gridExtra)
library(boot)
```


## 1) Vital rates

The following estimates and code are from two papers on African striped mice (Rhabdomys pumilio):

- Nater et al. 2016; https://doi.org/10.1111/oik.03384
- Nater et al. 2018; https://doi.org/10.1111/1365-2656.12888


The first paper established the vital rate models. The second paper ran updated versions of the vital rate models, incorporated them into a stage-structured matrix population model, and performed a variety of analyses of the MPM. The striped mouse has a short life cycle, and both vital rate predictions and the MPM have a monthly time interval. 

The vital rates are defined as sets of functions, one each to calculate the predicted vital rate on the link scale (logit or log) and one to back-transform to the actual probabilities/rates. 
The functions use z-standardized environmental covariates as input and require the fitted model objects to be loaded into the workspace. 

Note that custom prediction functions are only provided for the vital rates originating from the mark-recapture analysis (survival S and transition/maturation probabilities Psi). For the remaining vital rates (breeding probability Bp, litter probability Lp, and litter size LS), R's internal `predict()` function is used. 

```{r}
# Load workspace containing fitted models
load('170503_allmodels_final.RData')

#################################
#### CMR PREDICTION FORMULAS ####
#################################

# SURVIVAL
# sample parameters
random_est_surv=matrix(ncol=18,nrow=1000) # matrix 18 cols for the 18 parameters and the rows are than the sampled ones

for(i in 1:length(est$estimate[1:18])){
  random_est_surv[,i]=rnorm(1000,mean=est$estimate[i],sd=est$se[i]) # fill the cols
}


# Provides the y value (linear predictor) of S, where logit(S) = y and y = beta1*x1 + beta2*x2
S.y = function(stage,temp,food,dens_F,i){
  
  X <- c(1, as.numeric(stage=='I'), as.numeric(stage=='P'), temp, food, dens_F,
         (stage=='I')*temp, (stage=='P')*temp,
         (stage=='I')*food, (stage=='P')*food,
         (stage=='I')*dens_F, (stage=='P')*dens_F,
         temp*food, food*dens_F,
         (stage=='I')*temp*food, (stage=='P')*temp*food,
         (stage=='I')*food*dens_F, (stage=='P')*food*dens_F)
  
  X %*% random_est_surv[i,]
}

# Transforms the y value with the correct link function
S.predict = function(stage,temp,food,dens_F,i){
  return(plogis(S.y(stage,temp,food,dens_F,i)))
}


# MATURATION
# sample parameters
random_est_m=matrix(ncol=64,nrow=1000)

for(i in 57:64){
  random_est_m[,i]=rnorm(1000,mean=est$estimate[i],sd=est$se[i]) # fill the cols
}


Psi.y = function(stage,temp,food,dens_F,i){
  X <- c(as.numeric(stage=='I'), as.numeric(stage=='P'),
         temp, food, dens_F,
         food*dens_F, temp*dens_F, temp*food)
  X %*% random_est_m[i,57:64]
}

Psi.predict = function(stage,temp,food,dens_F,i){
  return(1/(1+2*exp(-Psi.y(stage,temp,food,dens_F,i))))    
}



# Bp (breeding probability)
# model: bpmodF
summary(bpmodF)$coefficients[,"Estimate"]
summary(bpmodF)$coefficients[,"Std. Error"]
intercept.bp=rnorm(1000,mean=summary(bpmodF)$coefficients[,"Estimate"][1],sd=summary(bpmodF)$coefficients[,"Std. Error"][1])
coef_tempt.bp=rnorm(1000,mean=summary(bpmodF)$coefficients[,"Estimate"][2],sd=summary(bpmodF)$coefficients[,"Std. Error"][2])
coef_foodt.bp=rnorm(1000,mean=summary(bpmodF)$coefficients[,"Estimate"][3],sd=summary(bpmodF)$coefficients[,"Std. Error"][3])
coef_densFt.bp=rnorm(1000,mean=summary(bpmodF)$coefficients[,"Estimate"][4],sd=summary(bpmodF)$coefficients[,"Std. Error"][4])
coef_tempt_foodt.bp=rnorm(1000,mean=summary(bpmodF)$coefficients[,"Estimate"][5],sd=summary(bpmodF)$coefficients[,"Std. Error"][5])
coef_tempt_densFt.bp=rnorm(1000,mean=summary(bpmodF)$coefficients[,"Estimate"][6],sd=summary(bpmodF)$coefficients[,"Std. Error"][6])

bp <- function(temp_t,food_t,dens_F_t,i){
  breedProb <- inv.logit(intercept.bp[i] + coef_tempt.bp[i]*temp_t + coef_foodt.bp[i]*food_t + coef_densFt.bp[i]*dens_F_t + coef_tempt_foodt.bp[i]*temp_t*food_t + coef_tempt_densFt.bp[i]*temp_t*dens_F_t) # what to do with random effects? ignore?
  return(breedProb)
}

# Lp (litter prob)
# model: lpmodF
summary(lpmodF)
summary(lpmodF)$coefficients[,"Estimate"]
summary(lpmodF)$coefficients[,"Std. Error"]
intercept.lp=rnorm(1000,mean=summary(lpmodF)$coefficients[,"Estimate"][1],sd=summary(lpmodF)$coefficients[,"Std. Error"][1])
coef_foodt.lp=rnorm(1000,mean=summary(lpmodF)$coefficients[,"Estimate"][2],sd=summary(lpmodF)$coefficients[,"Std. Error"][2])
coef_densFt.lp=rnorm(1000,mean=summary(lpmodF)$coefficients[,"Estimate"][3],sd=summary(lpmodF)$coefficients[,"Std. Error"][3])
coef_foodt_densFt.lp=rnorm(1000,mean=summary(lpmodF)$coefficients[,"Estimate"][4],sd=summary(lpmodF)$coefficients[,"Std. Error"][4])

lp <- function(food_t,dens_F_t,i){
  litterProb <- inv.logit(intercept.lp[i] + coef_foodt.lp[i]*food_t + coef_densFt.lp[i]*dens_F_t + coef_foodt_densFt.lp[i]*food_t*dens_F_t)
  return(litterProb)
}

# Ls (litter size)
# model: lsmod_Wlag
summary(lsmod_Wlag)
summary(lsmod_Wlag)$coefficients[,"Estimate"]
summary(lsmod_Wlag)$coefficients[,"Std. Error"]

intercept.ls=rnorm(1000,mean=summary(lsmod_Wlag)$coefficients[,"Estimate"][1],sd=summary(lsmod_Wlag)$coefficients[,"Std. Error"][1])
coef_tempt.ls=rnorm(1000,mean=summary(lsmod_Wlag)$coefficients[,"Estimate"][2],sd=summary(lsmod_Wlag)$coefficients[,"Std. Error"][2])
coef_foodt.ls=rnorm(1000,mean=summary(lsmod_Wlag)$coefficients[,"Estimate"][3],sd=summary(lsmod_Wlag)$coefficients[,"Std. Error"][3])
coef_densFt.ls=rnorm(1000,mean=summary(lsmod_Wlag)$coefficients[,"Estimate"][4],sd=summary(lsmod_Wlag)$coefficients[,"Std. Error"][4])
coef_tempt_foodt.ls=rnorm(1000,mean=summary(lsmod_Wlag)$coefficients[,"Estimate"][5],sd=summary(lsmod_Wlag)$coefficients[,"Std. Error"][5])
coef_foodt_densFt.ls=rnorm(1000,mean=summary(lsmod_Wlag)$coefficients[,"Estimate"][6],sd=summary(lsmod_Wlag)$coefficients[,"Std. Error"][6])

ls <- function(temp_t, food_t, dens_F_t, i){
  litterSize <- exp(intercept.ls[i] + coef_tempt.ls[i]*temp_t + coef_foodt.ls[i]*food_t + coef_densFt.ls[i]*dens_F_t + coef_tempt_foodt.ls[i]*temp_t*food_t + coef_foodt_densFt.ls[i]*dens_F_t*food_t)
  return(litterSize)
}



```


## 2) Covariates

There are the input data for the vital rate functions, and the time-series used in the study is provided as a dataframe named "env" in the workspace. This contains both raw and z-standardized versions of covariates temperature (mean.temp, temp) and food availability (life_cover, food). The density covariate is only provided in the standardized form (dens_all, dens_F). Additionally, the dataframe contains an unused NDVI variable, and season and Period variables that are only relevant for plotting and/or calculation of recapture probabilities.


```{r}
######################################
#### Preparing Environmental Data ####
######################################

# Extract scaling parameters for population density
m_dens <- 19.22219
sd_dens <- 8.199321

# Set "baseline" area
area.fix <- 1.9434353 # 19343.53/10000 

# Add "lagged" environmental covariates (t-1)
env$temp_t <- NA
env$food_t <- NA
env$dens_F_t <- NA

for(i in 2:nrow(env)){
  env$temp_t[i] <- env$temp[i-1]
  env$food_t[i] <- env$food[i-1]
  env$dens_F_t[i] <- env$dens_F[i-1]
}
```


```{r}
# z-standardized environmental covariates

# TEMP
max.temp=max(env$temp)
min.temp=min(env$temp)
mean.temp=0
sd.temp=1

# covariation
food_when_temp_max=env$food[which(env$temp==max(env$temp))]
food_when_temp_min=env$food[which(env$temp==min(env$temp))]

densF_when_temp_max=env$dens_F[which(env$temp==max(env$temp))]
densF_when_temp_min=env$dens_F[which(env$temp==min(env$temp))]

tempt_when_temp_max=env$temp_t[which(env$temp==max(env$temp))]
tempt_when_temp_min=env$temp_t[which(env$temp==min(env$temp))]

foodt_when_temp_max=env$food_t[which(env$temp==max(env$temp))]
foodt_when_temp_min=env$food_t[which(env$temp==min(env$temp))]

densFt_when_temp_max=env$dens_F_t[which(env$temp==max(env$temp))]
densFt_when_temp_min=env$dens_F_t[which(env$temp==min(env$temp))]

# TEMP_t
max.temp_t=max(env_merge_lag$temp_t)
min.temp_t=min(env_merge_lag$temp_t)
mean.temp_t=0
sd.temp_t=1
# covariation
temp_when_tempt_max=env$temp[which(env$temp_t==max(env_merge_lag$temp_t))]
temp_when_tempt_min=env$temp[which(env$temp_t==min(env_merge_lag$temp_t))]

food_when_tempt_max=env$food[which(env$temp_t==max(env_merge_lag$temp_t))]
food_when_tempt_min=env$food[which(env$temp_t==min(env_merge_lag$temp_t))]

densF_when_tempt_max=env$dens_F[which(env$temp_t==max(env_merge_lag$temp_t))]
densF_when_tempt_min=env$dens_F[which(env$temp_t==min(env_merge_lag$temp_t))]

foodt_when_tempt_max=env$food_t[which(env$temp_t==max(env_merge_lag$temp_t))]
foodt_when_tempt_min=env$food_t[which(env$temp_t==min(env_merge_lag$temp_t))]

densFt_when_tempt_max=env$dens_F_t[which(env$temp_t==max(env_merge_lag$temp_t))]
densFt_when_tempt_min=env$dens_F_t[which(env$temp_t==min(env_merge_lag$temp_t))]

# FOOD
max.food=max(env$food)
min.food=min(env$food)
mean.food=0
sd.food=1
# covariation
temp_when_food_max=env$temp[which(env$food==max(env$food))]
temp_when_food_min=env$temp[which(env$food==min(env$food))]

densF_when_food_max=env$dens_F[which(env$food==max(env$food))]
densF_when_food_min=env$dens_F[which(env$food==min(env$food))]

tempt_when_food_max=env$temp_t[which(env$food==max(env$food))]
tempt_when_food_min=env$temp_t[which(env$food==min(env$food))]

foodt_when_food_max=env$food_t[which(env$food==max(env$food))]
foodt_when_food_min=env$food_t[which(env$food==min(env$food))]

densFt_when_food_max=env$dens_F_t[which(env$food==max(env$food))]
densFt_when_food_min=env$dens_F_t[which(env$food==min(env$food))]

# FOOD_t
max.food_t=max(env_merge_lag$food_t)
min.food_t=min(env_merge_lag$food_t)
mean.food_t=0
sd.food_t=1

tempt_when_foodt_max=env$temp_t[which(env$food_t==max(env_merge_lag$food_t))]
tempt_when_foodt_min=env$temp_t[which(env$food_t==min(env_merge_lag$food_t))]
densFt_when_foodt_max=env$dens_F_t[which(env$food_t==max(env_merge_lag$food_t))]
densFt_when_foodt_min=env$dens_F_t[which(env$food_t==min(env_merge_lag$food_t))]
food_when_foodt_max=env$food[which(env$food_t==max(env_merge_lag$food_t))]
food_when_foodt_min=env$food[which(env$food_t==min(env_merge_lag$food_t))]
densF_when_foodt_max=env$dens_F[which(env$food_t==max(env_merge_lag$food_t))]
densF_when_foodt_min=env$dens_F[which(env$food_t==min(env_merge_lag$food_t))]
temp_when_foodt_max=env$temp[which(env$food_t==max(env_merge_lag$food_t))]
temp_when_foodt_min=env$temp[which(env$food_t==min(env_merge_lag$food_t))]


# DENS
max.dens_F=max(env$dens_F)
min.dens_F=min(env$dens_F)
mean.dens_F=0
sd.dens_F=1
# covariation
temp_when_densF_max=env$temp[which(env$dens_F==max(env$dens_F))]
temp_when_densF_min=env$temp[which(env$dens_F==min(env$dens_F))]

tempt_when_densF_max=env$temp_t[which(env$dens_F==max(env$dens_F))]
tempt_when_densF_min=env$temp_t[which(env$dens_F==min(env$dens_F))]

food_when_densF_max=env$food[which(env$dens_F==max(env$dens_F))]
food_when_densF_min=env$food[which(env$dens_F==min(env$dens_F))]

foodt_when_densF_max=env$food_t[which(env$dens_F==max(env$dens_F))]
foodt_when_densF_min=env$food_t[which(env$dens_F==min(env$dens_F))]

densFt_when_densF_max=env$dens_F_t[which(env$dens_F==max(env$dens_F))]
densFt_when_densF_min=env$dens_F_t[which(env$dens_F==min(env$dens_F))]

# DENS_t
max.dens_F_t=max(env_merge_lag$dens_F_t)
min.dens_F_t=min(env_merge_lag$dens_F_t)
mean.dens_F_t=0
sd.dens_F_t=1

food_when_densFt_max=env$food[which(env$dens_F_t==max(env_merge_lag$dens_F_t))]
food_when_densFt_min=env$food[which(env$dens_F_t==min(env_merge_lag$dens_F_t))]

foodt_when_densFt_max=env$food_t[which(env$dens_F_t==max(env_merge_lag$dens_F_t))]
foodt_when_densFt_min=env$food_t[which(env$dens_F_t==min(env_merge_lag$dens_F_t))]

temp_when_densFt_max=env$temp[which(env$dens_F_t==max(env_merge_lag$dens_F_t))]
temp_when_densFt_min=env$temp[which(env$dens_F_t==min(env_merge_lag$dens_F_t))]

tempt_when_densFt_max=env$temp_t[which(env$dens_F_t==max(env_merge_lag$dens_F_t))]
tempt_when_densFt_min=env$temp_t[which(env$dens_F_t==min(env_merge_lag$dens_F_t))]

densF_when_densFt_max=env$dens_F[which(env$dens_F_t==max(env_merge_lag$dens_F_t))]
densF_when_densFt_min=env$dens_F[which(env$dens_F_t==min(env_merge_lag$dens_F_t))]

```


## 3) Population model

Below are functions for assembling the monthly projection matrices. It's a set of two nested functions, where the first one returns a matrix based on specified covariate values for temperature, food, and density, while the second one calls upon the first, but allows for incorporation of the density feedback (relevant for projections):

```{r}
# Original
#my.matrix = function(temp, temp_t, food, food_t, dens_F, dens_F_t, i){
  
#  Si <- S.predict('I',temp,food,dens_F,i) # immature
#  Sp <- S.predict('P',temp,food,dens_F,i) # philopatric
#  Sb <- S.predict('B',temp,food,dens_F,i) # breeder
  
#  PsiIB <- Psi.predict('I',temp,food,dens_F,i)
#  PsiPB <- Psi.predict('P',temp,food,dens_F,i)
  
#  newd <- data.frame(food_t = food_t, dens_F_t = dens_F_t, temp_t = temp_t) 
  
#  Bp <- unname(predict(bpmodF, newdata = newd, re.form=NA, type='response')) # breeding prob
  
#  Lp <- predict(lpmodF, type = "response", newdata = newd) # litter prob
#  LS <- predict(lsmod_Wlag, type = "response", newdata = newd) # litter size
  #LS <- 2.65
  
#  A <- matrix(data = c(Si*PsiIB*Lp*LS, Si*(1-PsiIB), Si*PsiIB,
#                       Sp*PsiPB*Lp*LS, Sp*(1-PsiPB), Sp*PsiPB,
#                       Sb*Bp*Lp*LS,            0,       Sb),nrow=3)
#  return(A)
#}

my.matrix = function(temp, temp_t, food, food_t, dens_F, dens_F_t, i){
  
  Si <- S.predict('I',temp,food,dens_F,i) # immature
  Sp <- S.predict('P',temp,food,dens_F,i) # philopatric
  Sb <- S.predict('B',temp,food,dens_F,i) # breeder
  
  PsiIB <- Psi.predict('I',temp,food,dens_F,i)
  PsiPB <- Psi.predict('P',temp,food,dens_F,i)
  
  Bp <- bp(temp_t = temp_t, food_t = food_t, dens_F_t = dens_F_t, i) # breeding prob
  
  Lp <- lp(food_t = food_t, dens_F_t = dens_F_t, i) # litter prob
  LS <- ls(temp_t = temp_t, food_t = food_t, dens_F_t = dens_F_t, i) # litter size
  
  A <- matrix(data = c(Si*PsiIB*Lp*LS, Si*(1-PsiIB), Si*PsiIB,
                       Sp*PsiPB*Lp*LS, Sp*(1-PsiPB), Sp*PsiPB,
                       Sb*Bp*Lp*LS,            0,       Sb),nrow=3)
  return(A)
}

n.lambda=NULL
for(i in 1:1000){
  n.lambda[i]=lambda(my.matrix(mean.temp,mean.temp_t,mean.food,mean.food_t,mean.dens_F,mean.dens_F_t,i))

}
hist(n.lambda)
mean(n.lambda)

```


## 4) Scaled sensitivity analyses 

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117) for all the vital rates and also for each vital rate alone. Since the population model is a function, I created a new file that includes new functions to perturb temp, food, and dens for each vital rate.

```{r}
# for all vital rates
delta.T=NULL
deltaT.cov=NULL
delta.T_t=NULL
deltaT_t.cov=NULL

delta.F=NULL
deltaF.cov=NULL
delta.F_t=NULL
deltaF_t.cov=NULL

delta.D=NULL
deltaD.cov=NULL
delta.D_t=NULL
deltaD_t.cov=NULL

# for each vital rate
source("pert.functions.striped.mouse2.mcmc.R")
delta.temp.Si=NULL
delta.temp.Sp=NULL
delta.temp.Sb=NULL
delta.temp.PsiIB=NULL
delta.temp.PsiPB=NULL 
delta.tempt.Bp=NULL
delta.tempt.Lp=NULL
delta.tempt.Ls=NULL

delta.food.Si=NULL
delta.food.Sp=NULL
delta.food.Sb=NULL
delta.food.PsiIB=NULL
delta.food.PsiPB=NULL
delta.foodt.Bp=NULL
delta.foodt.Lp=NULL
delta.foodt.Ls=NULL

delta.dens.Si=NULL
delta.dens.Sp=NULL
delta.dens.Sb=NULL
delta.dens.PsiIB=NULL
delta.dens.PsiPB=NULL
delta.denst.Bp=NULL
delta.denst.Lp=NULL
delta.denst.Ls=NULL

set.seed(1234)
for(i in 1:100){
  # ALL VITAL RATES ####################
  # TEMP ------------------------
  # 1. Sensitivity to temp assuming mean food, mean dens, and mean lagged covariates
  mpm.max=my.matrix(max.temp,mean.temp_t,mean.food,mean.food_t,mean.dens_F,mean.dens_F_t,i)
  mpm.min=my.matrix(min.temp,mean.temp_t,mean.food,mean.food_t,mean.dens_F,mean.dens_F_t,i)
  delta.T[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp-min.temp)/sd.temp))
  
  # 2. Sensitivity to temp assuming covariation among covariates
  mpm.max=my.matrix(max.temp,tempt_when_temp_max,food_when_temp_max,foodt_when_temp_max,densF_when_temp_max,densFt_when_temp_max,i)
  mpm.min=my.matrix(min.temp,tempt_when_temp_min,food_when_temp_min,foodt_when_temp_min,densF_when_temp_min,densFt_when_temp_min,i)
  deltaT.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp-min.temp)/sd.temp))
  
  # 3. Sensitivity to lagged temp assuming mean everything else
  mpm.max=my.matrix(mean.temp,max.temp_t,mean.food,mean.food_t,mean.dens_F,mean.dens_F_t,i)
  mpm.min=my.matrix(mean.temp,min.temp_t,mean.food,mean.food_t,mean.dens_F,mean.dens_F_t,i)
  delta.T_t[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp_t-min.temp_t)/sd.temp_t))
  
  # 4. Sensitivity to lagged temp assuming covariation among covariates
  mpm.max=my.matrix(temp_when_tempt_max,max.temp_t,food_when_tempt_max,foodt_when_tempt_max,densF_when_tempt_max,densFt_when_tempt_max,i)
  mpm.min=my.matrix(temp_when_tempt_min,min.temp_t,food_when_tempt_min,foodt_when_tempt_min,densF_when_tempt_min,densFt_when_tempt_min,i)
  deltaT_t.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp_t-min.temp_t)/sd.temp_t))

  # FOOD --------------------------
  # 1. Sensitivity to food assuming mean everything else
  mpm.max=my.matrix(mean.temp,mean.temp_t,max.food,mean.food_t,mean.dens_F,mean.dens_F_t,i)
  mpm.min=my.matrix(mean.temp,mean.temp_t,min.food,mean.food_t,mean.dens_F,mean.dens_F_t,i)
  delta.F[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.food-min.food)/sd.food))

  # 2. Sensitivity to food assuming covariation among covariates
  mpm.max=my.matrix(temp_when_food_max,tempt_when_food_max,max.food,foodt_when_food_max,densF_when_food_max,densFt_when_food_max,i)
  mpm.min=my.matrix(temp_when_food_min,tempt_when_food_min,min.food,foodt_when_food_min,densF_when_food_min,densFt_when_food_min,i)
  deltaF.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.food-min.food)/sd.food))
  
  # 3. Sensitivity to lagged food assuming mean covariates
  mpm.max=my.matrix(mean.temp,mean.temp_t,mean.food,max.food_t,mean.dens_F,mean.dens_F_t,i)
  mpm.min=my.matrix(mean.temp,min.temp_t,mean.food,min.food_t,mean.dens_F,mean.dens_F_t,i)
  delta.F_t[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.food_t-min.food_t)/sd.food_t))

  # 4. Sensitivity to lagged food assuming covariation among covariates
  mpm.max=my.matrix(temp_when_foodt_max,tempt_when_foodt_max,food_when_foodt_max,max.food_t,densF_when_foodt_max,densFt_when_foodt_max,i)
  mpm.min=my.matrix(temp_when_foodt_min,tempt_when_foodt_min,food_when_foodt_min,max.food_t,densF_when_foodt_min,densFt_when_foodt_min,i)
  deltaF_t.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.food_t-min.food_t)/sd.food_t))
  
  # DENS -------------------------
  # 1. Sensitivity to female density assuming mean everything else
  mpm.max=my.matrix(mean.temp,mean.temp_t,mean.food,mean.food_t,max.dens_F,mean.dens_F_t,i)
  mpm.min=my.matrix(mean.temp,mean.temp_t,mean.food,mean.food_t,min.dens_F,mean.dens_F_t,i)
  delta.D[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens_F-min.dens_F)/sd.dens_F))

  # 2. Sensitivity to female density assuming covariation among covariates
  mpm.max=my.matrix(temp_when_densF_max,tempt_when_densF_max,food_when_densF_max,foodt_when_densF_max,max.dens_F,densFt_when_densF_max,i)
  mpm.min=my.matrix(temp_when_densF_min,tempt_when_densF_min,food_when_densF_min,foodt_when_densF_min,min.dens_F,densFt_when_densF_min,i)
  deltaD.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens_F-min.dens_F)/sd.dens_F))
  
  # 3. Sensitivity to lagged density assuming mean covariates
  mpm.max=my.matrix(mean.temp,mean.temp_t,mean.food,mean.food_t,mean.dens_F,max.dens_F_t,i)
  mpm.min=my.matrix(mean.temp,mean.temp_t,mean.food,mean.food_t,mean.dens_F,min.dens_F_t,i)
  delta.D_t[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens_F_t-min.dens_F_t)/sd.dens_F_t))
  
   # 4. Sensitivity to lagged density assuming covariation among covariates
  mpm.max=my.matrix(temp_when_densFt_max,tempt_when_densFt_max,food_when_densFt_max,foodt_when_densFt_max,densF_when_densFt_max,max.dens_F_t,i)
  mpm.min=my.matrix(temp_when_densFt_min,tempt_when_densFt_min,food_when_densFt_min,foodt_when_densFt_min,densF_when_densFt_min,min.dens_F_t,i)
  deltaD_t.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens_F_t-min.dens_F_t)/sd.dens_F_t))
  
  # PERTUBATIONS PER VITAL RATE ############################
  # TEMP --------------------------------------
  # Si (immature survival)
  pert.Si.max=pert.Si(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_temp=max.temp,pert_food=food_when_temp_max,pert_dens_F=densF_when_temp_max,i)
  pert.Si.min=pert.Si(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_min,pert_temp=min.temp,pert_dens_F=densF_when_temp_min,i)
delta.temp.Si[i]=abs((lambda(pert.Si.max)-lambda(pert.Si.min))/((max.temp-min.temp)/sd.temp))

  # Sp philopatric survival
  pert.Sp.max=pert.Sp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_max,pert_temp=max.temp,pert_dens_F=densF_when_temp_max,i)
  pert.Sp.min=pert.Sp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_min,pert_temp=min.temp,pert_dens_F=densF_when_temp_min,i)
delta.temp.Sp[i]=abs((lambda(pert.Sp.max)-lambda(pert.Sp.min))/((max.temp-min.temp)/sd.temp))

  # Sb (breeder survival)
  pert.Sb.max=pert.Sb(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_max,pert_temp=max.temp,pert_dens_F=densF_when_temp_max,i)
  pert.Sb.min=pert.Sb(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_min,pert_temp=min.temp,pert_dens_F=densF_when_temp_min,i)
delta.temp.Sb[i]=abs((lambda(pert.Sb.max)-lambda(pert.Sb.min))/((max.temp-min.temp)/sd.temp))

  # PsiIB
  pert.PsiIB.max=pert.PsiIB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_max,pert_temp=max.temp,pert_dens_F=densF_when_temp_max,i)
  pert.PsiIB.min=pert.PsiIB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_min,pert_temp=min.temp,pert_dens_F=densF_when_temp_min,i)
delta.temp.PsiIB[i]=abs((lambda(pert.PsiIB.max)-lambda(pert.PsiIB.min))/((max.temp-min.temp)/sd.temp))

  # PsiPB
  pert.PsiPB.max=pert.PsiPB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_max,pert_temp=max.temp,pert_dens_F=densF_when_temp_max,i)
  pert.PsiPB.min=pert.PsiPB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_temp_min,pert_temp=min.temp,pert_dens_F=densF_when_temp_min,i)

delta.temp.PsiPB[i]=abs((lambda(pert.PsiPB.max)-lambda(pert.PsiPB.min))/((max.temp-min.temp)/sd.temp))

  # Bp (breeding probability) with lagged covariates
  pert.Bp.max=pert.Bp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_tempt_max,pert_temp_t=max.temp_t,pert_dens_F_t=densFt_when_tempt_max,i)
  pert.Bp.min=pert.Bp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_tempt_min,pert_temp_t=min.temp_t,pert_dens_F_t=densFt_when_tempt_min,i)
  delta.tempt.Bp[i]=abs((lambda(pert.Bp.max)-lambda(pert.Bp.min))/((max.temp_t-min.temp_t)/sd.temp_t))

  # Lp (litter probability) with lagged covariates
  # no temp_t in Lp
  
  # Ls (litter size) with lagged covariates
  pert.Ls.max=pert.Ls(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_tempt_max,pert_temp_t=max.temp_t,pert_dens_F_t=densFt_when_tempt_max,i)
  pert.Ls.min=pert.Ls(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_tempt_min,pert_temp_t=min.temp_t,pert_dens_F_t=densFt_when_tempt_min,i)
  delta.tempt.Ls[i]=abs((lambda(pert.Ls.max)-lambda(pert.Ls.min))/((max.temp_t-min.temp_t)/sd.temp_t))


  # FOOD -------------------------------
  # Si (immature survival)
  pert.Si.max=pert.Si(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=max.food,pert_temp=temp_when_food_max,pert_dens_F=densF_when_food_max,i)
  pert.Si.min=pert.Si(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=min.food,pert_temp=temp_when_food_min,pert_dens_F=densF_when_food_min,i)
  delta.food.Si[i]=abs((lambda(pert.Si.max)-lambda(pert.Si.min))/((max.food-min.food)/sd.food))

  # Sp philopatric survival
  pert.Sp.max=pert.Sp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=max.food,pert_temp=temp_when_food_max,pert_dens_F=densF_when_food_max,i)
  pert.Sp.min=pert.Sp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=min.food,pert_temp=temp_when_food_min,pert_dens_F=densF_when_food_min,i)
  delta.food.Sp[i]=abs((lambda(pert.Sp.max)-lambda(pert.Sp.min))/((max.food-min.food)/sd.food))

  # Sb (breeder survival)
  pert.Sb.max=pert.Sb(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=max.food,pert_temp=temp_when_food_max,pert_dens_F=densF_when_food_max,i)
  pert.Sb.min=pert.Sb(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=min.food,pert_temp=temp_when_food_min,pert_dens_F=densF_when_food_min,i)
  delta.food.Sb[i]=abs((lambda(pert.Sb.max)-lambda(pert.Sb.min))/((max.food-min.food)/sd.food))

  # PsiIB
  pert.PsiIB.max=pert.PsiIB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=max.food,pert_temp=temp_when_food_max,pert_dens_F=densF_when_food_max,i)
  pert.PsiIB.min=pert.PsiIB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=min.food,pert_temp=temp_when_food_min,pert_dens_F=densF_when_food_min,i)
  delta.food.PsiIB[i]=abs((lambda(pert.PsiIB.max)-lambda(pert.PsiIB.min))/((max.food-min.food)/sd.food))

  # PsiPB
  pert.PsiPB.max=pert.PsiPB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=max.food,pert_temp=temp_when_food_max,pert_dens_F=densF_when_food_max,i)
  pert.PsiPB.min=pert.PsiPB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=min.food,pert_temp=temp_when_food_min,pert_dens_F=densF_when_food_min,i)
  delta.food.PsiPB[i]=abs((lambda(pert.PsiPB.max)-lambda(pert.PsiPB.min))/((max.food-min.food)/sd.food))

  # Bp (breeding probability) with lagged covariates
  pert.Bp.max=pert.Bp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=max.food_t,pert_temp_t=tempt_when_foodt_max,pert_dens_F_t=densFt_when_foodt_max,i)
  pert.Bp.min=pert.Bp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=min.food_t,pert_temp_t=tempt_when_foodt_min,pert_dens_F_t=densFt_when_foodt_min,i)
  delta.foodt.Bp[i]=abs((lambda(pert.Bp.max)-lambda(pert.Bp.min))/((max.food_t-min.food_t)/sd.food_t))

  # Lp (litter prob) with lagged covariates
  pert.Lp.max=pert.Lp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=max.food_t,pert_dens_F_t=densFt_when_foodt_max,i)
  pert.Lp.min=pert.Lp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=min.food_t,pert_dens_F_t=densFt_when_foodt_min,i)
  delta.foodt.Lp[i]=abs((lambda(pert.Lp.max)-lambda(pert.Lp.min))/((max.food_t-min.food_t)/sd.food_t))

  # Ls (litter size) with lagged covariates
  pert.Ls.max=pert.Ls(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=max.food_t,pert_temp_t=tempt_when_foodt_max,pert_dens_F_t=densFt_when_foodt_max,i)
  pert.Ls.min=pert.Ls(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=min.food_t,pert_temp_t=tempt_when_foodt_min,pert_dens_F_t=densFt_when_foodt_min,i)
  delta.foodt.Ls[i]=abs((lambda(pert.Ls.max)-lambda(pert.Ls.min))/((max.food_t-min.food_t)/sd.food_t))



# DENS --------------------------------
  # Si (immature survival)
  pert.Si.max=pert.Si(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_max,pert_temp=temp_when_densF_max,pert_dens_F=max.dens_F,i)
  pert.Si.min=pert.Si(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_min,pert_temp=temp_when_densF_min,pert_dens_F=min.dens_F,i)
  delta.dens.Si[i]=abs((lambda(pert.Si.max)-lambda(pert.Si.min))/((max.dens_F-min.dens_F)/sd.dens_F))

  # Sp philopatric survival
  pert.Sp.max=pert.Sp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_max,pert_temp=temp_when_densF_max,pert_dens_F=max.dens_F,i)
  pert.Sp.min=pert.Sp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_min,pert_temp=temp_when_densF_min,pert_dens_F=min.dens_F,i)
  delta.dens.Sp[i]=abs((lambda(pert.Sp.max)-lambda(pert.Sp.min))/((max.dens_F-min.dens_F)/sd.dens_F))

  # Sb (breeder survival)
  pert.Sb.max=pert.Sb(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_max,pert_temp=temp_when_densF_max,pert_dens_F=max.dens_F,i)
  pert.Sb.min=pert.Sb(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_min,pert_temp=temp_when_densF_min,pert_dens_F=min.dens_F,i)
  delta.dens.Sb[i]=abs((lambda(pert.Sb.max)-lambda(pert.Sb.min))/((max.dens_F-min.dens_F)/sd.dens_F))

  # PsiIB
  pert.PsiIB.max=pert.PsiIB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_max,pert_temp=temp_when_densF_max,pert_dens_F=max.dens_F,i)
  pert.PsiIB.min=pert.PsiIB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_min,pert_temp=temp_when_densF_min,pert_dens_F=min.dens_F,i)
  delta.dens.PsiIB[i]=abs((lambda(pert.PsiIB.max)-lambda(pert.PsiIB.min))/((max.dens_F-min.dens_F)/sd.dens_F))

  # PsiPB
  pert.PsiPB.max=pert.PsiPB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_max,pert_temp=temp_when_densF_max,pert_dens_F=max.dens_F,i)
  pert.PsiPB.min=pert.PsiPB(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food=food_when_densF_min,pert_temp=temp_when_densF_min,pert_dens_F=min.dens_F,i)
  delta.dens.PsiPB[i]=abs((lambda(pert.PsiPB.max)-lambda(pert.PsiPB.min))/((max.dens_F-min.dens_F)/sd.dens_F))

  # Bp (breeding probability) with lagged covariates
  pert.Bp.max=pert.Bp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_densFt_max,pert_temp_t=tempt_when_densFt_max,pert_dens_F_t=max.dens_F_t,i)
  pert.Bp.min=pert.Bp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_densFt_min,pert_temp_t=tempt_when_densFt_min,pert_dens_F_t=min.dens_F_t,i)
  delta.denst.Bp[i]=abs((lambda(pert.Bp.max)-lambda(pert.Bp.min))/((max.dens_F_t-min.dens_F_t)/sd.dens_F_t))

  # Lp (litter prob) with lagged covariates
  pert.Lp.max=pert.Lp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_densFt_max,pert_dens_F_t=max.dens_F_t,i)
  pert.Lp.min=pert.Lp(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_densFt_min,pert_dens_F_t=min.dens_F_t,i)
  delta.denst.Lp[i]=abs((lambda(pert.Lp.max)-lambda(pert.Lp.min))/((max.dens_F_t-min.dens_F_t)/sd.dens_F_t))

  # Ls (litter size) with lagged covariates
  pert.Ls.max=pert.Ls(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_densFt_max,pert_temp_t=tempt_when_densFt_max,pert_dens_F_t=max.dens_F_t,i)
  pert.Ls.min=pert.Ls(temp=mean.temp, temp_t=mean.temp_t,food=mean.food,food_t=mean.food_t,dens_F=mean.dens_F,dens_F_t=mean.dens_F_t,pert_food_t=foodt_when_densFt_min,pert_temp_t=tempt_when_densFt_min,pert_dens_F_t=min.dens_F_t,i)
  delta.denst.Ls[i]=abs((lambda(pert.Ls.max)-lambda(pert.Ls.min))/((max.dens_F_t-min.dens_F_t)/sd.dens_F_t))
  
}

hist(delta.T-deltaT.cov)
hist(delta.F-deltaF.cov)
hist(delta.D-deltaD.cov)
hist(delta.T_t-deltaT_t.cov)
hist(delta.F_t-deltaF_t.cov)
hist(delta.D_t-deltaD_t.cov)
```

## 5) Save output
```{r}
all.sens <- data.frame(study.doi="10.1111/1365-2656.12888",
                   year.of.publication="2018",
                   group="Mammals",
                   species="Rhabdomys pumilio",
                   continent="Africa",
                   driver=rep(c("temperature","food availability","density","lagged temperature","lagged food availability","lagged density"),each=200),
                   driver.type=rep(c("C","B","D","C","B","D"),each=200),
                   stage.age="all",
                   vital.rates="all",
                   sens=c(delta.T,deltaT.cov,
                          delta.F,deltaF.cov,
                          delta.D,deltaD.cov,
                          delta.T_t,deltaT_t.cov,
                          delta.F_t,deltaF_t.cov,
                          delta.D_t,deltaD_t.cov),
                   cov=rep(c(0,1),100),
                   gen.time=1, # generation time
                   n.vr=8, # number of vital rates with covariatee
                   n.pam=98, # number of parameters
                   dens=1, # density dependence in it?
                   biotic_interactions=1) # any biotic interactions?



write.csv(all.sens,"Sens_StripedMouse_MCMC.csv",row.names = F)

#  Si <- S.predict('I',temp,food,dens_F,i) # immature
#  Sp <- S.predict('P',temp,food,dens_F,i) # philopatric
#  Sb <- S.predict('B',temp,food,dens_F,i) # breeder
  
#  PsiIB <- Psi.predict('I',temp,food,dens_F,i)
#  PsiPB <- Psi.predict('P',temp,food,dens_F,i)
  
#  newd <- data.frame(food_t = food_t, dens_F_t = dens_F_t, temp_t = temp_t) 
  
#  Bp <- unname(predict(bpmodF, newdata = newd, re.form=NA, type='response')) # breeding prob
  
#  Lp <- predict(lpmodF, type = "response", newdata = newd) # litter prob
#  LS <- predict(lsmod_Wlag, type = "response", newdata = newd) # litter size
  #LS <- 2.65

# sens per vr
sens.vr.temp <- data.frame(study.doi="10.1111/1365-2656.12888",
                   year.of.publication="2018",
                   group="Mammals",
                   species="Rhabdomys pumilio",
                   continent="Africa",
                   driver=c(rep("temperature",500),rep("lagged temperature",200)),
                   driver.type="C",
                   stage.age=c(rep("immature",100),rep("non-reproductive adult",100),rep("reproductive adult",100),
                               rep("immature",100),rep("non-reproductive adult",100),
                               rep("all",100),rep("all",100)),
                   vital.rates=c(rep("survival",100),rep("survival",100),rep("survival",100),
                                 rep("maturation probability",100),rep("maturation probability",100),
                                 rep("breeding probability",100),rep("litter size",100)),
                   sens=c(delta.temp.Si,delta.temp.Sp,delta.temp.Sb,
                          delta.temp.PsiIB,delta.temp.PsiPB,
                          delta.tempt.Bp,delta.tempt.Ls))

sens.vr.food <- data.frame(study.doi="10.1111/1365-2656.12888",
                   year.of.publication="2018",
                   group="Mammals",
                   species="Rhabdomys pumilio",
                   continent="Africa",
                   driver=c(rep("food",500),rep("lagged food",300)),
                   driver.type="C",
                   stage.age=c(rep("immature",100),rep("non-reproductive adult",100),rep("reproductive adult",100),
                               rep("immature",100),rep("non-reproductive adult",100),
                               rep("all",100),rep("all",100),rep("all",100)),
                   vital.rates=c(rep("survival",100),rep("survival",100),rep("survival",100),
                                 rep("maturation probability",100),rep("maturation probability",100),
                                 rep("breeding probability",100),rep("litter size",100),rep("litter probability",100)),
                   sens=c(delta.food.Si,delta.food.Sp,delta.food.Sb,
                          delta.food.PsiIB,delta.food.PsiPB,
                          delta.foodt.Bp,delta.foodt.Ls,delta.foodt.Lp))

sens.vr.dens <- data.frame(study.doi="10.1111/1365-2656.12888",
                   year.of.publication="2018",
                   group="Mammals",
                   species="Rhabdomys pumilio",
                   continent="Africa",
                   driver=c(rep("density",500),rep("lagged density",300)),
                   driver.type="C",
                   stage.age=c(rep("immature",100),rep("non-reproductive adult",100),rep("reproductive adult",100),
                               rep("immature",100),rep("non-reproductive adult",100),
                               rep("all",100),rep("all",100),rep("all",100)),
                   vital.rates=c(rep("survival",100),rep("survival",100),rep("survival",100),
                                 rep("maturation probability",100),rep("maturation probability",100),
                                 rep("breeding probability",100),rep("litter size",100),rep("litter probability",100)),
                   sens=c(delta.dens.Si,delta.dens.Sp,delta.dens.Sb,
                          delta.dens.PsiIB,delta.dens.PsiPB,
                          delta.denst.Bp,delta.denst.Ls,delta.denst.Lp))

sens.vr=rbind(sens.vr.temp,sens.vr.food,sens.vr.dens)

sens.vr$gen.time=1
sens.vr$n.vr=8 # number of vital rates with covariates
sens.vr$n.pam=98 # number of parameters
sens.vr$dens=1
sens.vr$biotic_interactions=1

write.csv(sens.vr,"SensVR_StripedMouse_MCMC.csv",row.names = F)
```

