---
title: "SensitivityAnalysis_MPM_Dipper"
author: "Esin Ickin"
date: "2/23/2023"
output: html_document
---

## 0) Prepare session
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Files_Dipper")
```

```{r}
rm(list=ls())
library(boot)
library(popbio)
library(tidyverse)

cov <- read.csv("cov_Dipper.csv")
cov$N <- round(cov$N)

cov2 <- read.csv("Env_cov_dipper.txt", sep="")

# annual number of local birds
cov$Nlocal <- round(cov2$Nlocal)

# annual number of immigrants
cov$immigrants <- cov$N - cov$Nlocal

# immigrants per capita
cov$imm.per.capita <- cov$immigrants/cov$N
```


## 1) Vital rates


```{r}
set.seed(1234)
# SURVIVAL
# Parameter estimates table S2
mu=rnorm(100,mean=0.456,sd=0.212)
beta1.1=0
beta1.2=rnorm(100,mean=0.477,sd=0.048)
beta1.3=rnorm(100,mean=0.167,sd=0.048)
beta1.4=rnorm(100,mean=0.196,sd=0.048)
beta2=rnorm(100,mean=0.16,sd=0.122)
beta3=rnorm(100,mean=-0.008,sd=0.003)
beta5.1=0
beta5.2=rnorm(100,mean=-0.002,sd=0.001)
beta5.3=rnorm(100,mean=-0.001,sd=0.001)
beta5.4=rnorm(100,mean=-0.002,sd=0.001)
beta.year=0.287 # random year effect
res.S=0.013 # random residual effects
kR.all=0.57 

# age class 1
S1 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.1*1 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.1*1*dens*exp(-kR.all*temp))
  return(surv)
}

# age class 2
S2 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.2[i]*2 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.2[i]*2*dens*exp(-kR.all*temp))
  return(surv)
  }

# age class 3
S3 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.3[i]*3 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.3[i]*3*dens*exp(-kR.all*temp))
  return(surv)
  }

# age class 4
S4 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.4[i]*4 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.4[i]*4*dens*exp(-kR.all*temp))
  return(surv)
}

# test: S4(1.6,100,4)

# RECRUITMENT

# parameter estimates from table S2
v=rnorm(100,mean=-2.246,sd=0.249)
beta1.1=0
beta1.2=rnorm(100,mean=0.433,sd=0.052)
beta1.3=rnorm(100,mean=0.453,sd=0.052)
beta1.4=rnorm(100,mean=0.705,sd=0.052)
beta2=rnorm(100,mean=0.236,sd=0.158)
beta3=rnorm(100,mean=-0.017,sd=0.003)
beta.year=0.796 # random year effect
res.R=0.045 # random residual effects
kR.1=0.519
kR.2=0.403
kR.3=0.42
kR.4=0.34

# age class 1

# mean immigrants per capita rate
imm <- mean(cov$imm.per.capita)

R1 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.1*1 + beta2[i]*temp + beta3[i]*dens*exp(-kR.1*temp)) + imm
  return(rec)
}

# age class 2
R2 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.2[i]*2 + beta2[i]*temp + beta3[i]*dens*exp(-kR.2*temp))
  return(rec)
}

# age class 3
R3 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.3[i]*3 + beta2[i]*temp + beta3[i]*dens*exp(-kR.3*temp))
  return(rec)
}

# age class 4
R4 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.4[i]*4 + beta2[i]*temp + beta3[i]*dens*exp(-kR.4*temp))
  return(rec)
}

# IMMIGRATION
# eq 6
# without year effect
#I <- function(Nlocal, temp){
#  i <- exp(41.145/(1+exp(0.101-0.738*temp)*exp(-0.037*Nlocal)))
#  return(i)
#}

#I(mean.Nlocal.scaled, mean.temp)
#I(mean.Nlocal, mean.temp)
```

## 2) Covariates

```{r}
# density
mean.dens=mean(cov$N)
sd.dens=sd(cov$N)
min.dens=min(cov$N)
max.dens=max(cov$N)

# scaled density
#min.dens.scaled=(min.dens-mean.dens)/sd.dens
#max.dens.scaled=(max.dens-mean.dens)/sd.dens
#sd.dens.scaled=1
#mean.dens.scaled=0

# standardized mean winter temperature
#mean.temp=0
#sd.temp=1
#min.temp=-2
#max.temp=2
mean.temp=mean(cov$temp)
sd.temp=sd(cov$temp)
min.temp=min(cov$temp)
max.temp=max(cov$temp)

# covariation
dens_when_temp_max=cov$N[which(cov$temp==max(cov$temp))][1]
dens_when_temp_min=cov$N[which(cov$temp==min(cov$temp))][1]

temp_when_dens_max=cov$temp[which(cov$N==max(cov$N))][1]
temp_when_dens_min=cov$temp[which(cov$N==min(cov$N))][1]
```

## 3) Population model

Here, we use the vital rate function to construct an annual population model that can give us the population growth rate (lambda).

In the following example, the MPM is constructed with mean covariate values.
In the perturbations, the covariate values are changed.

```{r}
n.stage=4
lam=NULL
for(i in 1:100){
  mpm = matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)
  lam[i]=lambda(mpm)
  
}

hist(lam)
```


## 4) Scaled sensitivity analyses 

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117)

Note that this is a step that Esin will implement in her MS thesis. With the information given in 1-3, we should be able to run these analyses.

```{r}
delta.T=NULL
deltaT.cov=NULL

delta.D=NULL
deltaD.cov=NULL

deltaT.R1=NULL
deltaT.R2=NULL
deltaT.R3=NULL
deltaT.R4=NULL
deltaT.S1=NULL
deltaT.S2=NULL
deltaT.S3=NULL
deltaT.S4=NULL

deltaD.R1=NULL
deltaD.R2=NULL
deltaD.R3=NULL
deltaD.R4=NULL
deltaD.S1=NULL
deltaD.S2=NULL
deltaD.S3=NULL
deltaD.S4=NULL

for(i in 1:100){
  # TEMP
# 1. Sensitivity to temperature assuming mean density
mpm.max=matrix(c(R1(max.temp,mean.dens,i),S1(max.temp,mean.dens,i),0,0,
               R2(max.temp,mean.dens,i),0,S2(max.temp,mean.dens,i),0,
               R3(max.temp,mean.dens,i),0,0,S3(max.temp,mean.dens,i),
               R4(max.temp,mean.dens,i),0,0,S4(max.temp,mean.dens,i)),n.stage,n.stage)

mpm.min=matrix(c(R1(min.temp,mean.dens,i),S1(min.temp,mean.dens,i),0,0,
               R2(min.temp,mean.dens,i),0,S2(min.temp,mean.dens,i),0,
               R3(min.temp,mean.dens,i),0,0,S3(min.temp,mean.dens,i),
               R4(min.temp,mean.dens,i),0,0,S4(min.temp,mean.dens,i)),n.stage,n.stage)

delta.T[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp-min.temp)/sd.temp))


# 2. Sensitivity to temperature assuming covariation among covariates
mpm.max=matrix(c(R1(max.temp,dens_when_temp_max,i),S1(max.temp,dens_when_temp_max,i),0,0,
               R2(max.temp,dens_when_temp_max,i),0,S2(max.temp,dens_when_temp_max,i),0,
               R3(max.temp,dens_when_temp_max,i),0,0,S3(max.temp,dens_when_temp_max,i),
               R4(max.temp,dens_when_temp_max,i),0,0,S4(max.temp,dens_when_temp_max,i)),n.stage,n.stage)

mpm.min=matrix(c(R1(min.temp,dens_when_temp_min,i),S1(min.temp,dens_when_temp_min,i),0,0,
               R2(min.temp,dens_when_temp_min,i),0,S2(min.temp,dens_when_temp_min,i),0,
               R3(min.temp,dens_when_temp_min,i),0,0,S3(min.temp,dens_when_temp_min,i),
               R4(min.temp,dens_when_temp_min,i),0,0,S4(min.temp,dens_when_temp_min,i)),n.stage,n.stage)


deltaT.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp-min.temp)/sd.temp))
  
# DENSITY
# 1. Sensitivity to density assuming mean temperature
mpm.max = matrix(c(R1(mean.temp,max.dens,i),S1(mean.temp,max.dens,i),0,0,
               R2(mean.temp,max.dens,i),0,S2(mean.temp,max.dens,i),0,
               R3(mean.temp,max.dens,i),0,0,S3(mean.temp,max.dens,i),
               R4(mean.temp,max.dens,i),0,0,S4(mean.temp,max.dens,i)),n.stage,n.stage)

mpm.min = matrix(c(R1(mean.temp,min.dens,i),S1(mean.temp,min.dens,i),0,0,
               R2(mean.temp,min.dens,i),0,S2(mean.temp,min.dens,i),0,
               R3(mean.temp,min.dens,i),0,0,S3(mean.temp,min.dens,i),
               R4(mean.temp,min.dens,i),0,0,S4(mean.temp,min.dens,i)),n.stage,n.stage)

delta.D[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens-min.dens)/sd.dens))
    
  
# 2. Sensitivity to density assuming covariation among covariates
mpm.max = matrix(c(R1(temp_when_dens_max,max.dens,i),S1(temp_when_dens_max,max.dens,i),0,0,
               R2(temp_when_dens_max,max.dens,i),0,S2(temp_when_dens_max,max.dens,i),0,
               R3(temp_when_dens_max,max.dens,i),0,0,S3(temp_when_dens_max,max.dens,i),
               R4(temp_when_dens_max,max.dens,i),0,0,S4(temp_when_dens_max,max.dens,i)),n.stage,n.stage)

mpm.min = matrix(c(R1(temp_when_dens_max,min.dens,i),S1(temp_when_dens_max,min.dens,i),0,0,
               R2(temp_when_dens_max,min.dens,i),0,S2(temp_when_dens_max,min.dens,i),0,
               R3(temp_when_dens_max,min.dens,i),0,0,S3(temp_when_dens_max,min.dens,i),
               R4(temp_when_dens_max,min.dens,i),0,0,S4(temp_when_dens_max,min.dens,i)),n.stage,n.stage)

deltaD.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens-min.dens)/sd.dens))

# Perturb R1

# TEMP

R1.max=matrix(c(R1(max.temp,dens_when_temp_max,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R1.min=matrix(c(R1(min.temp,dens_when_temp_min,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R1[i]=abs((lambda(R1.max)-lambda(R1.min))/((max.temp-min.temp)/sd.temp))

# DENS
R1.max=matrix(c(R1(temp_when_dens_max,max.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R1.min=matrix(c(R1(temp_when_dens_min,min.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R1[i]=abs((lambda(R1.max)-lambda(R1.min))/((max.dens-min.dens)/sd.dens))


# Perturb R2

# TEMP
R2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(max.temp,dens_when_temp_max,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.dens,mean.dens,i),0,0,
               R2(min.temp,dens_when_temp_min,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R2[i]=abs((lambda(R2.max)-lambda(R2.min))/((max.temp-min.temp)/sd.temp))

# DENS
R2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(temp_when_dens_max,max.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(temp_when_dens_min,min.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R2[i]=abs((lambda(R2.max)-lambda(R2.min))/((max.dens-min.dens)/sd.dens))


# Perturb R3

# TEMP
R3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(max.temp,dens_when_temp_max,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(min.temp,dens_when_temp_min,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R3[i]=abs((lambda(R3.max)-lambda(R3.min))/((max.temp-min.temp)/sd.temp))

# DENS
R3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(temp_when_dens_max,max.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(temp_when_dens_min,min.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R3[i]=abs((lambda(R3.max)-lambda(R3.min))/((max.dens-min.dens)/sd.dens))


# Perturb R4

# TEMP
R4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(max.temp,dens_when_temp_max,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(min.temp,dens_when_temp_min,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R4[i]=abs((lambda(R4.max)-lambda(R4.min))/((max.temp-min.temp)/sd.temp))

# DENS
R4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(temp_when_dens_max,max.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(temp_when_dens_min,min.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R4[i]=abs((lambda(R4.max)-lambda(R4.min))/((max.dens-min.dens)/sd.dens))

# Perturb S1

# TEMP
S1.max=matrix(c(R1(mean.temp,mean.dens,i),S1(max.temp,dens_when_temp_max,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S1.min=matrix(c(R1(mean.temp,mean.dens,i),S1(min.temp,dens_when_temp_min,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.S1[i]=abs((lambda(S1.max)-lambda(S1.min))/((max.temp-min.temp)/sd.temp))

# DENS
S1.max=matrix(c(R1(mean.temp,mean.dens,i),S1(temp_when_dens_max,max.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S1.min=matrix(c(R1(mean.temp,mean.dens,i),S1(temp_when_dens_min,min.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.S1[i]=abs((lambda(S1.max)-lambda(S1.min))/((max.dens-min.dens)/sd.dens))


# Perturb S2

# TEMP
S2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(max.temp,dens_when_temp_max,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(min.temp,dens_when_temp_min,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.S2[i]=abs((lambda(S2.max)-lambda(S2.min))/((max.temp-min.temp)/sd.temp))

# DENS
S2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(temp_when_dens_max,max.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(temp_when_dens_min,min.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.S2[i]=abs((lambda(S2.max)-lambda(S2.min))/((max.dens-min.dens)/sd.dens))


# Perturb S3

# TEMP
S3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(max.temp,dens_when_temp_max,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(min.temp,dens_when_temp_min,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.S3[i]=abs((lambda(S3.max)-lambda(S3.min))/((max.temp-min.temp)/sd.temp))

# DENS
S3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(temp_when_dens_max,max.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(temp_when_dens_min,min.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.S3[i]=abs((lambda(S3.max)-lambda(S3.min))/((max.dens-min.dens)/sd.dens))


# Perturb S4

# TEMP
S4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(max.temp,dens_when_temp_max,i)),n.stage,n.stage)

S4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(min.temp,dens_when_temp_min,i)),n.stage,n.stage)

deltaT.S4[i]=abs((lambda(S4.max)-lambda(S4.min))/((max.temp-min.temp)/sd.temp))

# DENS
S4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(temp_when_dens_max,max.dens,i)),n.stage,n.stage)

S4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(temp_when_dens_min,min.dens,i)),n.stage,n.stage)

deltaD.S4[i]=abs((lambda(S4.max)-lambda(S4.min))/((max.dens-min.dens)/sd.dens))
}

hist(delta.T-deltaT.cov) # should be positive
hist(delta.D-deltaD.cov)


```

## 5) Save output
```{r}
sens.temp=data.frame(sens=c(delta.T,deltaT.cov,deltaT.R1,deltaT.R2,deltaT.R3,deltaT.R4,deltaT.S1,deltaT.S2,deltaT.S3,deltaT.S4),
              driver="temp",
              vital.rate=rep(c("all","all with cov", "R1", "R2", "R3", "R4", "S1", "S2", "S3", "S4"), each = 100))

sens.dens=data.frame(sens=c(delta.D,deltaD.cov,deltaD.R1,deltaD.R2,deltaD.R3,deltaD.R4,deltaD.S1,deltaD.S2,deltaD.S3,deltaD.S4),
              driver="dens",
              vital.rate=rep(c("all","all with cov", "R1", "R2", "R3", "R4", "S1", "S2", "S3", "S4"), each = 100))

all.sens=rbind(sens.temp,sens.dens)
tail(all.sens)

write.csv(all.sens,"all_sensitivities_dipper.csv")


sens=data.frame(sens.no.cov=c(delta.T,delta.D),
                sens.with.cov=c(deltaT.cov,deltaD.cov),
                driver=rep(c("temp","dens"),each=100))

write.csv(sens,"sensitivities_with_covariation_and_without.csv")
```

Reorganize the dataframes (from the "ReOrganize.R" script)
```{r}
# re-organise dipper sensitivities per vital rate
rm(list=ls())
# load data
df=read.csv("all_sensitivities_dipper.csv")
# We have these vital rates R1-R4 and S1-S4
# And 2 drivers
# For each driver and vital rate combo we have 100 sensitivities

# filter data
df <- df %>% filter(!(vital.rate %in% c("all", "all with cov")))

driver=df$driver

df2 <- data.frame(study.doi="10.1126/sciadv.1602298",
                  year.of.publication="2017",
                  group="Birds",
                  species="Cinclus cinclus",
                  continent="Europe",
                  driver=rep(c("temperature","density"),each=800),
                  driver.type=rep(c("C","D"),each=800),
                  stage.age=rep(c(c(rep(c("1","2","3","4"),each=100),rep(c("1","2","3","4"),each=100))),2),
                  vital.rates=rep(c("reproduction","survival"),each=400),
                  sens=df$sens)

write.csv(df2,"Output/Sens_VR_Dipper_Resampling.csv",row.names = F)


# re-organise dipper sens (no covariation vs covariation)
rm(list=ls())

# load data
df=read.csv("sensitivities_with_covariation_and_without.csv")

df2 <- data.frame(study.doi="10.1126/sciadv.1602298",
                  year.of.publication="2017",
                  group="Birds",
                  species="Cinclus cinclus",
                  continent="Europe",
                  driver=df$driver,
                  driver.type=rep(c("C","D"),each=100),
                  stage.age="all",
                  vital.rates="all",
                  sens.no.cov=df$sens.no.cov,
                  sens.cov=df$sens.with.cov)

write.csv(df2,"Output/Sens_Dipper_Resampling.csv",row.names=F)

```

 
 
 
 
 
