---
title: "SensitivityAnalysis_MPM_Dipper"
author: "Esin Ickin"
date: "2/23/2023"
output: html_document
---

This code is can be used as an example on how to parameterize vital-rate functions using coefficients and covariate inputs to then build a matrix population model and compute sensitivities.

## 0) Prepare session
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Library/CloudStorage/OneDrive-UniversitätZürichUZH/Master Thesis/pert_analyses/dipper")
```

```{r}
rm(list=ls())
library(boot)
library(popbio)
library(tidyverse)

# load data
cov <- read.csv("cov_Dipper.csv")
cov$N <- round(cov$N)
cov2 <- read.csv("Env_cov_dipper.txt", sep="")

# annual number of local birds
cov$Nlocal <- round(cov2$Nlocal)

# annual number of immigrants
cov$immigrants <- cov$N - cov$Nlocal

# immigrants per capita
cov$imm.per.capita <- cov$immigrants/cov$N
```


## 1) Vital rates

The example is from a paper on dippers: Gamelon et al. 2017. 
The vital rates are defined as functions. To compute the uncertainties around our sensitivity estimates, we need the distributions of the parameters.

```{r}
set.seed(1234)
# SURVIVAL
# Parameter estimates table S2
mu=rnorm(100,mean=0.456,sd=0.212)
beta1.1=0
beta1.2=rnorm(100,mean=0.477,sd=0.048)
beta1.3=rnorm(100,mean=0.167,sd=0.048)
beta1.4=rnorm(100,mean=0.196,sd=0.048)
beta2=rnorm(100,mean=0.16,sd=0.122)
beta3=rnorm(100,mean=-0.008,sd=0.003)
beta5.1=0
beta5.2=rnorm(100,mean=-0.002,sd=0.001)
beta5.3=rnorm(100,mean=-0.001,sd=0.001)
beta5.4=rnorm(100,mean=-0.002,sd=0.001)
kR.all=0.57 
#beta.year=0.287 # random year effect
#res.S=0.013 # random residual effects

# age class 1
S1 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.1*1 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.1*1*dens*exp(-kR.all*temp))
  return(surv)
}

# age class 2
S2 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.2[i]*2 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.2[i]*2*dens*exp(-kR.all*temp))
  return(surv)
  }

# age class 3
S3 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.3[i]*3 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.3[i]*3*dens*exp(-kR.all*temp))
  return(surv)
  }

# age class 4
S4 <- function(temp, dens, i){
  surv <- inv.logit(mu[i] + beta1.4[i]*4 + beta2[i]*temp + beta3[i]*dens*exp(-kR.all*temp) + beta5.4[i]*4*dens*exp(-kR.all*temp))
  return(surv)
}

# test: S4(1.6,100,4)

# RECRUITMENT

# parameter estimates from table S2
v=rnorm(100,mean=-2.246,sd=0.249)
beta1.1=0
beta1.2=rnorm(100,mean=0.433,sd=0.052)
beta1.3=rnorm(100,mean=0.453,sd=0.052)
beta1.4=rnorm(100,mean=0.705,sd=0.052)
beta2=rnorm(100,mean=0.236,sd=0.158)
beta3=rnorm(100,mean=-0.017,sd=0.003)
kR.1=0.519
kR.2=0.403
kR.3=0.42
kR.4=0.34
#beta.year=0.796 # random year effect
#res.R=0.045 # random residual effects

# age class 1

# mean immigrants per capita rate
imm <- mean(cov$imm.per.capita)

R1 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.1*1 + beta2[i]*temp + beta3[i]*dens*exp(-kR.1*temp)) + imm
  return(rec)
}

# age class 2
R2 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.2[i]*2 + beta2[i]*temp + beta3[i]*dens*exp(-kR.2*temp))
  return(rec)
}

# age class 3
R3 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.3[i]*3 + beta2[i]*temp + beta3[i]*dens*exp(-kR.3*temp))
  return(rec)
}

# age class 4
R4 <- function(temp, dens, i){
  rec <- exp(v[i] + beta1.4[i]*4 + beta2[i]*temp + beta3[i]*dens*exp(-kR.4*temp))
  return(rec)
}

# IMMIGRATION
# eq 6
# without year effect
#I <- function(Nlocal, temp){
#  i <- exp(41.145/(1+exp(0.101-0.738*temp)*exp(-0.037*Nlocal)))
#  return(i)
#}

#I(mean.Nlocal.scaled, mean.temp)
#I(mean.Nlocal, mean.temp)
```

## 2) Covariates

These are the input data for the vital rate functions. Ideally, it should be a time series because it allows us to calculate not only means and variances but also covariances, the latter being important to calculate scaled sensitivities.

```{r}
# density
mean.dens=mean(cov$N)
sd.dens=sd(cov$N)
min.dens=min(cov$N)
max.dens=max(cov$N)

# scaled density
#min.dens.scaled=(min.dens-mean.dens)/sd.dens
#max.dens.scaled=(max.dens-mean.dens)/sd.dens
#sd.dens.scaled=1
#mean.dens.scaled=0

# standardized mean winter temperature
#mean.temp=0
#sd.temp=1
#min.temp=-2
#max.temp=2
mean.temp=mean(cov$temp)
sd.temp=sd(cov$temp)
min.temp=min(cov$temp)
max.temp=max(cov$temp)

# covariation
dens_when_temp_max=cov$N[which(cov$temp==max(cov$temp))][1]
dens_when_temp_min=cov$N[which(cov$temp==min(cov$temp))][1]

temp_when_dens_max=cov$temp[which(cov$N==max(cov$N))][1]
temp_when_dens_min=cov$temp[which(cov$N==min(cov$N))][1]
```

## 3) Population model

Here, we use the vital rate function to construct an annual population model that can give us the population growth rate (lambda).

In the following example, the MPM is constructed with mean covariate values.
In the perturbations, the covariate values are changed.

```{r}
n.stage=4
lam=NULL
for(i in 1:100){
  mpm = matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)
  lam[i]=lambda(mpm)
  
}

hist(lam)
```


## 4) Scaled sensitivity analyses 

Here, we calculate scaled sensitivities, according to Morris et al. 2020 (DOI: https://doi.org/10.1073/pnas.1918363117)

With the information given in 1-3, we should be able to run these analyses.

```{r}
# empty vectors to store the sensitivities
delta.T=NULL
deltaT.cov=NULL

delta.D=NULL
deltaD.cov=NULL

deltaT.R1=NULL
deltaT.R2=NULL
deltaT.R3=NULL
deltaT.R4=NULL
deltaT.S1=NULL
deltaT.S2=NULL
deltaT.S3=NULL
deltaT.S4=NULL

deltaD.R1=NULL
deltaD.R2=NULL
deltaD.R3=NULL
deltaD.R4=NULL
deltaD.S1=NULL
deltaD.S2=NULL
deltaD.S3=NULL
deltaD.S4=NULL

# repeat calculation 100 times and resample parameters each time
for(i in 1:100){
  # TEMP
# 1. Sensitivity to temperature assuming mean density
mpm.max=matrix(c(R1(max.temp,mean.dens,i),S1(max.temp,mean.dens,i),0,0,
               R2(max.temp,mean.dens,i),0,S2(max.temp,mean.dens,i),0,
               R3(max.temp,mean.dens,i),0,0,S3(max.temp,mean.dens,i),
               R4(max.temp,mean.dens,i),0,0,S4(max.temp,mean.dens,i)),n.stage,n.stage)

mpm.min=matrix(c(R1(min.temp,mean.dens,i),S1(min.temp,mean.dens,i),0,0,
               R2(min.temp,mean.dens,i),0,S2(min.temp,mean.dens,i),0,
               R3(min.temp,mean.dens,i),0,0,S3(min.temp,mean.dens,i),
               R4(min.temp,mean.dens,i),0,0,S4(min.temp,mean.dens,i)),n.stage,n.stage)

delta.T[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp-min.temp)/sd.temp))


# 2. Sensitivity to temperature assuming covariation among covariates
mpm.max=matrix(c(R1(max.temp,dens_when_temp_max,i),S1(max.temp,dens_when_temp_max,i),0,0,
               R2(max.temp,dens_when_temp_max,i),0,S2(max.temp,dens_when_temp_max,i),0,
               R3(max.temp,dens_when_temp_max,i),0,0,S3(max.temp,dens_when_temp_max,i),
               R4(max.temp,dens_when_temp_max,i),0,0,S4(max.temp,dens_when_temp_max,i)),n.stage,n.stage)

mpm.min=matrix(c(R1(min.temp,dens_when_temp_min,i),S1(min.temp,dens_when_temp_min,i),0,0,
               R2(min.temp,dens_when_temp_min,i),0,S2(min.temp,dens_when_temp_min,i),0,
               R3(min.temp,dens_when_temp_min,i),0,0,S3(min.temp,dens_when_temp_min,i),
               R4(min.temp,dens_when_temp_min,i),0,0,S4(min.temp,dens_when_temp_min,i)),n.stage,n.stage)


deltaT.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.temp-min.temp)/sd.temp))
  
# DENSITY
# 1. Sensitivity to density assuming mean temperature
mpm.max = matrix(c(R1(mean.temp,max.dens,i),S1(mean.temp,max.dens,i),0,0,
               R2(mean.temp,max.dens,i),0,S2(mean.temp,max.dens,i),0,
               R3(mean.temp,max.dens,i),0,0,S3(mean.temp,max.dens,i),
               R4(mean.temp,max.dens,i),0,0,S4(mean.temp,max.dens,i)),n.stage,n.stage)

mpm.min = matrix(c(R1(mean.temp,min.dens,i),S1(mean.temp,min.dens,i),0,0,
               R2(mean.temp,min.dens,i),0,S2(mean.temp,min.dens,i),0,
               R3(mean.temp,min.dens,i),0,0,S3(mean.temp,min.dens,i),
               R4(mean.temp,min.dens,i),0,0,S4(mean.temp,min.dens,i)),n.stage,n.stage)

delta.D[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens-min.dens)/sd.dens))
    
  
# 2. Sensitivity to density assuming covariation among covariates
mpm.max = matrix(c(R1(temp_when_dens_max,max.dens,i),S1(temp_when_dens_max,max.dens,i),0,0,
               R2(temp_when_dens_max,max.dens,i),0,S2(temp_when_dens_max,max.dens,i),0,
               R3(temp_when_dens_max,max.dens,i),0,0,S3(temp_when_dens_max,max.dens,i),
               R4(temp_when_dens_max,max.dens,i),0,0,S4(temp_when_dens_max,max.dens,i)),n.stage,n.stage)

mpm.min = matrix(c(R1(temp_when_dens_max,min.dens,i),S1(temp_when_dens_max,min.dens,i),0,0,
               R2(temp_when_dens_max,min.dens,i),0,S2(temp_when_dens_max,min.dens,i),0,
               R3(temp_when_dens_max,min.dens,i),0,0,S3(temp_when_dens_max,min.dens,i),
               R4(temp_when_dens_max,min.dens,i),0,0,S4(temp_when_dens_max,min.dens,i)),n.stage,n.stage)

deltaD.cov[i]=abs((lambda(mpm.max)-lambda(mpm.min))/((max.dens-min.dens)/sd.dens))

# Finally, perturb covariates (temp, rain, dens) in each vital rate to get vital-rate specific sensitivities

# Perturb R1

# TEMP

R1.max=matrix(c(R1(max.temp,dens_when_temp_max,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R1.min=matrix(c(R1(min.temp,dens_when_temp_min,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R1[i]=abs((lambda(R1.max)-lambda(R1.min))/((max.temp-min.temp)/sd.temp))

# DENS
R1.max=matrix(c(R1(temp_when_dens_max,max.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R1.min=matrix(c(R1(temp_when_dens_min,min.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R1[i]=abs((lambda(R1.max)-lambda(R1.min))/((max.dens-min.dens)/sd.dens))


# Perturb R2

# TEMP
R2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(max.temp,dens_when_temp_max,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.dens,mean.dens,i),0,0,
               R2(min.temp,dens_when_temp_min,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R2[i]=abs((lambda(R2.max)-lambda(R2.min))/((max.temp-min.temp)/sd.temp))

# DENS
R2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(temp_when_dens_max,max.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(temp_when_dens_min,min.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R2[i]=abs((lambda(R2.max)-lambda(R2.min))/((max.dens-min.dens)/sd.dens))


# Perturb R3

# TEMP
R3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(max.temp,dens_when_temp_max,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(min.temp,dens_when_temp_min,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R3[i]=abs((lambda(R3.max)-lambda(R3.min))/((max.temp-min.temp)/sd.temp))

# DENS
R3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(temp_when_dens_max,max.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(temp_when_dens_min,min.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R3[i]=abs((lambda(R3.max)-lambda(R3.min))/((max.dens-min.dens)/sd.dens))


# Perturb R4

# TEMP
R4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(max.temp,dens_when_temp_max,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(min.temp,dens_when_temp_min,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.R4[i]=abs((lambda(R4.max)-lambda(R4.min))/((max.temp-min.temp)/sd.temp))

# DENS
R4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(temp_when_dens_max,max.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

R4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(temp_when_dens_min,min.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.R4[i]=abs((lambda(R4.max)-lambda(R4.min))/((max.dens-min.dens)/sd.dens))

# Perturb S1

# TEMP
S1.max=matrix(c(R1(mean.temp,mean.dens,i),S1(max.temp,dens_when_temp_max,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S1.min=matrix(c(R1(mean.temp,mean.dens,i),S1(min.temp,dens_when_temp_min,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.S1[i]=abs((lambda(S1.max)-lambda(S1.min))/((max.temp-min.temp)/sd.temp))

# DENS
S1.max=matrix(c(R1(mean.temp,mean.dens,i),S1(temp_when_dens_max,max.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S1.min=matrix(c(R1(mean.temp,mean.dens,i),S1(temp_when_dens_min,min.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.S1[i]=abs((lambda(S1.max)-lambda(S1.min))/((max.dens-min.dens)/sd.dens))


# Perturb S2

# TEMP
S2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(max.temp,dens_when_temp_max,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(min.temp,dens_when_temp_min,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.S2[i]=abs((lambda(S2.max)-lambda(S2.min))/((max.temp-min.temp)/sd.temp))

# DENS
S2.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(temp_when_dens_max,max.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S2.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(temp_when_dens_min,min.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.S2[i]=abs((lambda(S2.max)-lambda(S2.min))/((max.dens-min.dens)/sd.dens))


# Perturb S3

# TEMP
S3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(max.temp,dens_when_temp_max,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(min.temp,dens_when_temp_min,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaT.S3[i]=abs((lambda(S3.max)-lambda(S3.min))/((max.temp-min.temp)/sd.temp))

# DENS
S3.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(temp_when_dens_max,max.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

S3.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(temp_when_dens_min,min.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(mean.temp,mean.dens,i)),n.stage,n.stage)

deltaD.S3[i]=abs((lambda(S3.max)-lambda(S3.min))/((max.dens-min.dens)/sd.dens))


# Perturb S4

# TEMP
S4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(max.temp,dens_when_temp_max,i)),n.stage,n.stage)

S4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(min.temp,dens_when_temp_min,i)),n.stage,n.stage)

deltaT.S4[i]=abs((lambda(S4.max)-lambda(S4.min))/((max.temp-min.temp)/sd.temp))

# DENS
S4.max=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(temp_when_dens_max,max.dens,i)),n.stage,n.stage)

S4.min=matrix(c(R1(mean.temp,mean.dens,i),S1(mean.temp,mean.dens,i),0,0,
               R2(mean.temp,mean.dens,i),0,S2(mean.temp,mean.dens,i),0,
               R3(mean.temp,mean.dens,i),0,0,S3(mean.temp,mean.dens,i),
               R4(mean.temp,mean.dens,i),0,0,S4(temp_when_dens_min,min.dens,i)),n.stage,n.stage)

deltaD.S4[i]=abs((lambda(S4.max)-lambda(S4.min))/((max.dens-min.dens)/sd.dens))
}

hist(delta.T-deltaT.cov) # should be positive
hist(delta.D-deltaD.cov)

```


## 5) Save output
```{r}
# sens all VR
sens <- data.frame(study.doi="10.1126/sciadv.1602298",
                  year.of.publication="2017",
                  group="Birds",
                  species="Cinclus cinclus",
                  continent="Europe",
                  driver=c(rep(c("temperature","density"),each=200)),
                  driver.type=c(rep(c("C","D"),each=200)),
                  stage.age="all",
                  vital.rates="all",
                  sens=c(delta.T,deltaT.cov,delta.D,deltaD.cov),
                  cov=rep(c(0,1),each=100),
                  gen.time=1, # generation time
                  n.vr=8, # number of vital rates with covariates
                  n.pam=44, # number of parameters (per vital rate ca. 5.5)
                  dens=1, # density dependence in it?
                  biotic_interactions=0) # any biotic interactions?

write.csv(sens,"Sens_Dipper_Resampling.csv",row.names = F)


# sens per VR
sensVR <- data.frame(study.doi="10.1126/sciadv.1602298",
                  year.of.publication="2017",
                  group="Birds",
                  species="Cinclus cinclus",
                  continent="Europe",
                  driver=c(rep(c("temperature","density"),each=800)),
                  driver.type=c(rep(c("C","D"),each=800)),
                  stage.age=c(rep(c("Reproduction","Survival"),each=400)),
                  vital.rates=c(rep(c("1","2","3","4"),each=100)),
                  sens=c(deltaT.R1,deltaT.R2,deltaT.R3,deltaT.R4,
                         deltaT.S1,deltaT.S2,deltaT.S3,deltaT.S4,
                         deltaD.R1,deltaD.R2,deltaD.R3,deltaD.R4,
                         deltaD.S1,deltaD.S2,deltaD.S3,deltaD.S4),
                  gen.time=1, # generation time
                  n.vr=8, # number of vital rates with covariates
                  n.pam=44, # number of parameters (per vital rate ca. 5.5)
                  dens=1, # density dependence in it?
                  biotic_interactions=0) # other biotic interactions?


write.csv(sensVR,"Sens_VR_Dipper_Resampling.csv",row.names = F)

```

 
 
 
